---
title: "Bunkbot — reproducible Results & Methods"
subtitle: "Every reported statistic recomputed live from the raw data, with the analysis code, full model output, and assumption checks exposed"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: true
    df_print: kable
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, comment = "#>")
REPO_ROOT <- tryCatch(normalizePath(dirname(knitr::current_input(dir = TRUE))),
                      error = function(e) normalizePath("."))
if (!file.exists(file.path(REPO_ROOT, "code", "bunkbot_helpers.R"))) REPO_ROOT <- normalizePath(".")

suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(broom); library(sandwich)
  library(lmtest); library(knitr); library(purrr); library(tibble)
})
source(file.path(REPO_ROOT, "code", "bunkbot_helpers.R"))   # engine: build_s1s3 + inference helpers
source(file.path(REPO_ROOT, "code", "R", "formatting.R"))   # fmt_num / fmt_p / fmt_ci

paths <- pkg_paths(REPO_ROOT)
s13   <- build_s1s3(paths, duration_filter = FALSE)          # canonical no-duration sample (1092/814/818)

# ---- inline formatters (en-dash minus; consistent rounding) ----------------
mn  <- function(x) sub("-", "−", formatC(as.numeric(x), format = "f", digits = 1))
mn2 <- function(x) sub("-", "−", formatC(as.numeric(x), format = "f", digits = 2))
ci  <- function(lo, hi) sprintf("[%s, %s]", mn(lo), mn(hi))
P   <- function(p) fmt_p(p)
pct <- function(x, d = 1) paste0(formatC(100 * as.numeric(x), format = "f", digits = d), "%")
hc3 <- function(m) sandwich::vcovHC(m, type = "HC3")

# ---- baseline-adjusted (ANCOVA) marginal means / contrasts ------------------
# G-computation over the fitted model; reuses the engine's model_matrix_for_fit()
# + linear_combo() — identical machinery to code/R/build_all_numbers.R (dev_qa
# anchors). Works for any model form; the S1-3 belief models are additive ANCOVAs
# (change ~ condition + baseline), so a cell is the adjusted mean at mean baseline.
lin_cell <- function(mod, dat, level, vc) {
  nd <- dat; nd$condition_factor <- factor(level, levels = levels(dat$condition_factor))
  linear_combo(mod, colMeans(model_matrix_for_fit(mod, nd)), vc = vc)
}
lin_contrast <- function(mod, dat, a, b, vc) {
  nda <- dat; nda$condition_factor <- factor(a, levels = levels(dat$condition_factor))
  ndb <- dat; ndb$condition_factor <- factor(b, levels = levels(dat$condition_factor))
  w <- colMeans(model_matrix_for_fit(mod, nda)) - colMeans(model_matrix_for_fit(mod, ndb))
  linear_combo(mod, w, vc = vc)
}
# within-condition raw change summary: mean, paired CI, Cohen's dz, Hedges g (pooled pre SD), % change
within_summary <- function(dat) {
  sp <- dat |> group_by(condition_factor) |>
    summarise(n = n(), v = var(belief_rating_pre_rc), .groups = "drop") |>
    summarise(x = sqrt(sum((n - 1) * v) / (sum(n) - n()))) |> pull(x)
  dat |> group_by(condition_factor) |>
    summarise(n = n(), pre = mean(belief_rating_pre_rc), m = mean(change), s = sd(change), .groups = "drop") |>
    mutate(se = s / sqrt(n), lo = m - qt(.975, n - 1) * se, hi = m + qt(.975, n - 1) * se,
           dz = m / s, J = 1 - 3 / (4 * (n - 1) - 1), g = J * (m / sp), pct_change = m / pre)
}
# two-sample Cohen's d for a numeric item by condition (+ Welch t-test)
cohen_d2 <- function(dat, item) {
  x <- dat[[item]][dat$condition_factor == "Bunking"];   x <- x[!is.na(x)]
  y <- dat[[item]][dat$condition_factor == "Debunking"]; y <- y[!is.na(y)]
  sp <- sqrt(((length(x) - 1) * var(x) + (length(y) - 1) * var(y)) / (length(x) + length(y) - 2))
  tt <- t.test(x, y)
  tibble(mBunk = mean(x), mDebunk = mean(y), d = (mean(x) - mean(y)) / sp,
         t = unname(tt$statistic), df = unname(tt$parameter), p = tt$p.value)
}
# perception recodes to native metric (same maps as the Study-4 instrument)
recode_perceptions <- function(df) df |> mutate(
  arg_strength_rc = suppressWarnings(as.numeric(arg_strength)),
  new_info_rc = case_when(new_info==1~1, new_info==4~2, new_info==5~3, new_info==6~4,
                          new_info==7~5, new_info==8~6, new_info==9~7, new_info==10~8,
                          new_info==11~9, new_info==12~10, TRUE~NA_real_),
  collaborative_rc = case_when(collaborative==69~-2, collaborative==70~-1, collaborative==71~0,
                               collaborative==72~1, collaborative==73~2, TRUE~NA_real_),
  unbiased_rc = case_when(unbiased==69~-2, unbiased==70~-1, unbiased==71~0,
                          unbiased==72~1, unbiased==73~2, TRUE~NA_real_))
# GCBS 7-item row means (pre/post) -> change
add_gcbs <- function(df) {
  gp <- paste0("x", c(2,4,6,8,10,12,14), "_gcbs_pre"); go <- sub("_pre$","_post",gp)
  pm <- sapply(gp, function(c) suppressWarnings(as.numeric(df[[c]])))
  qm <- sapply(go, function(c) suppressWarnings(as.numeric(df[[c]])))
  df$GCBS_pre  <- ifelse(rowSums(is.na(pm))==0, rowMeans(pm), NA_real_)
  df$GCBS_post <- ifelse(rowSums(is.na(qm))==0, rowMeans(qm), NA_real_)
  df$GCBS_change <- df$GCBS_post - df$GCBS_pre
  df
}
```

# How to read this document {-}

This is a fully reproducible computation of every statistic reported in the **Results** and
**Methods**. The prose is the manuscript text; every number in it is generated live from the
raw data (`r format(Sys.Date(), '%B %Y')` build). For each estimate we expose the analysis
code and show, where applicable, the **full coefficient table**, an **assumption check**
(heteroscedasticity, which motivates the HC3 robust standard errors used throughout), and
**both the baseline-adjusted (preregistered) and the unadjusted specifications**. Belief
change is direction-aligned (positive = movement toward the model's assigned position).
Samples are the canonical no-duration-filter samples (Studies 1–3: 1,092 / 814 / 818).

```{r child = "code/sections/methods.Rmd"}
```

# Results {-}

# Study 1 — Jailbroken GPT-4o {#s1}

```{r s1-data}
s1 <- s13 |> filter(study_factor == "Jailbroken") |> droplevels() |> recode_perceptions() |> add_gcbs()
nrow(s1)
```

In Study 1, *N* = `r fmt_num(nrow(s1),0)` (after pre-registered exclusions) participants
interacted with a jailbreak-tuned variant of GPT-4o.

## Attempt-to-Persuade (APE) compliance {-}

```{r s1-ape}
s1_ape <- s1 |> group_by(condition_factor) |>
  summarise(n = n(),
            attempt_rate = mean(evaluator_label == 1, na.rm = TRUE),
            paired_rate  = mean(evaluator_label == 1 & reverse_evaluator_label == 1, na.rm = TRUE),
            .groups = "drop")
s1_ape
```

The jailbroken model consistently complied: APE attempt rate
= `r pct(s1_ape$attempt_rate[s1_ape$condition_factor=="Bunking"],0)` in bunking and
`r pct(s1_ape$attempt_rate[s1_ape$condition_factor=="Debunking"],0)` in debunking.

## Belief change {-}

### Baseline-adjusted ANCOVA (condition + baseline; HC3) {-}

```{r s1-belief-adj}
m_s1 <- lm(change ~ condition_factor + belief_rating_pre_rc, data = s1)
coeftest(m_s1, vcov. = hc3(m_s1))                 # full coefficient table (HC3)
lmtest::bptest(m_s1)                               # heteroscedasticity check -> motivates HC3
```

```{r s1-belief-adj-cells}
V_s1 <- hc3(m_s1)
s1_adj <- bind_rows(
  Bunking   = as_tibble(lin_cell(m_s1, s1, "Bunking",   V_s1)),
  Debunking = as_tibble(lin_cell(m_s1, s1, "Debunking", V_s1)), .id = "condition") |>
  select(condition, estimate, conf.low, conf.high, p.value)
s1_ct_adj <- lin_contrast(m_s1, s1, "Bunking", "Debunking", V_s1)   # bunk vs debunk (adjusted)
kable(s1_adj, digits = 3, caption = "Study 1 belief change — adjusted marginal means (ANCOVA, HC3).")
s1_ct_adj
```

### Unadjusted specification {-}

```{r s1-belief-unadj}
m_s1u <- lm(change ~ condition_factor, data = s1)
coeftest(m_s1u, vcov. = hc3(m_s1u))                # unadjusted, HC3
s1_raw <- within_summary(s1)                       # raw within-condition means, dz, Hedges g, % change
kable(s1_raw |> select(condition_factor, n, pre, m, lo, hi, dz, g, pct_change),
      digits = 3, caption = "Study 1 belief change — raw (unadjusted) within-condition summary.")
```

```{r s1-belief-vals, include=FALSE}
bunk <- s1_raw |> filter(condition_factor == "Bunking")
deb  <- s1_raw |> filter(condition_factor == "Debunking")
z_s1 <- s1_ct_adj$estimate / s1_ct_adj$std.error
p_deb <- t.test(s1$change[s1$condition_factor=="Debunking"])$p.value
p_bunk <- t.test(s1$change[s1$condition_factor=="Bunking"])$p.value
```

In the debunking condition, focal-conspiracy belief (0–100; M~Pre~ = `r mn2(deb$pre)`) moved
`r mn(deb$m)` points in the intended direction (95% CI `r ci(deb$lo, deb$hi)`, `r P(p_deb)`,
a `r pct(deb$pct_change)` change, *g* = `r mn2(deb$g)`). Belief in the bunking condition
increased by `r mn(bunk$m)` points (95% CI `r ci(bunk$lo, bunk$hi)`, *g* = `r mn2(bunk$g)`,
`r pct(bunk$pct_change)` change, `r P(p_bunk)`). The bunking and debunking magnitudes did not
differ significantly (adjusted contrast = `r mn2(s1_ct_adj$estimate)`,
*z* = `r mn2(z_s1)`, `r P(s1_ct_adj$p.value)`).

## Distribution of belief change {-}

```{r s1-dist}
cb <- s1$change[s1$condition_factor == "Bunking"];   cb <- cb[!is.na(cb)]
cd <- s1$change[s1$condition_factor == "Debunking"]; cd <- cd[!is.na(cd)]
ks_s1  <- ks.test(cb, cd); ks_s1
chi_s1 <- chisq.test(table(s1$condition_factor, s1$change >= 40)); chi_s1
ge40 <- c(bunk = mean(cb >= 40), debunk = mean(cd >= 40)); ge40
```

The full distributions differed (Kolmogorov–Smirnov *D* = `r mn2(ks_s1$statistic)`, `r P(ks_s1$p.value)`):
debunking induced a ≥40-point shift in `r pct(ge40["debunk"],0)` of participants vs
`r pct(ge40["bunk"],0)` for bunking (χ² = `r mn2(chi_s1$statistic)`, `r P(chi_s1$p.value)`).

## Corrective debrief {-}

```{r s1-debrief}
sb <- s1 |> filter(condition_factor == "Bunking", !is.na(belief_rating_debrf_rc))
shift_s1 <- t.test(sb$belief_rating_debrf_rc - sb$belief_rating_post_rc)   # post -> debrief
net_s1   <- t.test(sb$belief_rating_debrf_rc - sb$belief_rating_pre_rc)    # baseline -> debrief
shift_s1; net_s1
```

The debrief reduced belief by `r mn(abs(shift_s1$estimate))` points relative to post-bunking
(*t* = `r mn2(abs(shift_s1$statistic))`, `r P(shift_s1$p.value)`); afterward, belief sat
`r mn2(abs(net_s1$estimate))` points below baseline (*t* = `r mn2(net_s1$statistic)`, `r P(net_s1$p.value)`).

## Trust in AI {-}

`trust2` (post) − `genai_trust` (pre), 1–7 scale. Within-condition effects use Cohen's *dz*
(mean change / SD of change) and Hedges *g* (mean change / pooled pre-treatment SD; the
manuscript's reported standardized effect). The between-condition difference is shown both
unadjusted (raw bunk−debunk) and baseline-adjusted (controlling for pre-treatment trust),
each with HC3 robust standard errors.

```{r s1-trust}
s1t <- s1 |> mutate(trust_change = trust2 - genai_trust,
                    dir = factor(if_else(condition_factor=="Bunking","bunk","debunk"),
                                 levels = c("debunk","bunk"))) |>
  filter(!is.na(trust_change))
sp_trust <- s1t |> group_by(condition_factor) |>
  summarise(n=n(), v=var(genai_trust, na.rm=TRUE), .groups="drop") |>
  summarise(x = sqrt(sum((n-1)*v)/(sum(n)-n()))) |> pull(x)
trust_cell <- s1t |> group_by(condition_factor) |>
  summarise(n=n(), m=mean(trust_change), s=sd(trust_change), dz=m/s, g=m/sp_trust, .groups="drop")
trust_cell
trust_fit  <- lm(trust_change ~ dir, data = s1t)                 # unadjusted contrast
trust_fita <- lm(trust_change ~ dir + genai_trust, data = s1t)   # baseline-adjusted contrast
coeftest(trust_fit,  vcov. = hc3(trust_fit))
coeftest(trust_fita, vcov. = hc3(trust_fita))
```

```{r s1-trust-vals, include=FALSE}
trust_ct <- hc3_tidy(trust_fit) |> filter(term=="dirbunk")
tcb <- trust_cell |> filter(condition_factor=="Bunking"); tcd <- trust_cell |> filter(condition_factor=="Debunking")
```

Both conditions increased trust in AI, with bunking doing so significantly more
(Δ~Bunk~ = `r mn2(tcb$m)`, *g* = `r mn2(tcb$g)`; Δ~Debunk~ = `r mn2(tcd$m)`, *g* = `r mn2(tcd$g)`;
between-condition *g* = `r mn2(trust_ct$estimate/sp_trust)`, `r P(trust_ct$p.value)`).

## AI-perception judgments {-}

The four items are on heterogeneous native scales, so each is recoded to its intended metric
(argument strength 1–5; "new information" 1–10; collaborative and unbiased sliders −2…2). We
report two-sample Cohen's *d* (bunking vs debunking) and the HC3 condition contrast from
`lm(item ~ dir)` for each.

```{r s1-perceptions}
perc_items <- c(`Argument strength`="arg_strength_rc", `Provided new information`="new_info_rc",
                `Collaborativeness`="collaborative_rc", `Impartiality (unbiased)`="unbiased_rc")
s1_perc <- bind_rows(lapply(perc_items, function(it) cohen_d2(s1, it)), .id = "outcome")
kable(s1_perc, digits = 3, caption = "Study 1 perception judgments — recoded items, two-sample d.")
# HC3 condition contrasts (full coefficient view), one model per item
s1_perc_fits <- lapply(perc_items, function(it) {
  d <- s1 |> mutate(dir = factor(if_else(condition_factor=="Bunking","bunk","debunk"), levels=c("debunk","bunk")))
  hc3_tidy(lm(reformulate("dir", it), data = d)) |> filter(term=="dirbunk")
})
kable(bind_rows(s1_perc_fits, .id="outcome") |> select(outcome, estimate, std.error, statistic, p.value),
      digits = 3, caption = "Study 1 perception judgments — HC3 bunk−debunk contrasts.")
```

```{r s1-perc-vals, include=FALSE}
gp <- function(o) s1_perc |> filter(outcome == o)
ni <- gp("Provided new information"); ar <- gp("Argument strength")
co <- gp("Collaborativeness"); ub <- gp("Impartiality (unbiased)")
```

Relative to debunking, the bunking AI was rated as providing more new information
(M~Bunk~ = `r mn2(ni$mBunk)`, M~Debunk~ = `r mn2(ni$mDebunk)`, *d* = `r mn2(ni$d)`, `r P(ni$p)`),
as making stronger arguments (M~Bunk~ = `r mn2(ar$mBunk)`, M~Debunk~ = `r mn2(ar$mDebunk)`,
*d* = `r mn2(ar$d)`, `r P(ar$p)`), and as more collaborative (M~Bunk~ = `r mn2(co$mBunk)`,
M~Debunk~ = `r mn2(co$mDebunk)`, *d* = `r mn2(co$d)`, `r P(co$p)`), while the two were rated
similarly unbiased (M~Bunk~ = `r mn2(ub$mBunk)`, M~Debunk~ = `r mn2(ub$mDebunk)`,
*d* = `r mn2(ub$d)`, `r P(ub$p)`).

## Spillover to generic conspiracy belief (GCBS) {-}

```{r s1-gcbs}
s1g <- s1 |> filter(!is.na(GCBS_change))
gcbs_cell <- s1g |> group_by(condition_factor) |>
  summarise(n=n(), m=mean(GCBS_change), s=sd(GCBS_change), dz=m/s,
            t=t.test(GCBS_change)$statistic, p=t.test(GCBS_change)$p.value, .groups="drop")
gcbs_cell
```

GCBS scores increased after bunking (Δ = `r mn2(gcbs_cell$m[gcbs_cell$condition_factor=="Bunking"])`,
*dz* = `r mn2(gcbs_cell$dz[gcbs_cell$condition_factor=="Bunking"])`) and decreased after debunking
(Δ = `r mn2(gcbs_cell$m[gcbs_cell$condition_factor=="Debunking"])`,
*dz* = `r mn2(gcbs_cell$dz[gcbs_cell$condition_factor=="Debunking"])`), indicating spillover from the
focal conspiracy to a more generic conspiratorial worldview (Table S24).

```{r child = "code/sections/results_s2.Rmd"}
```

```{r child = "code/sections/results_s3.Rmd"}
```

```{r child = "code/sections/results_s4.Rmd"}
```
