# Methods (reproducible quantities)

The Methods narrative is reproduced as in the manuscript; every **quantitative**
claim is recomputed live from the raw data and the cached LLM-derived files. Each
quantity is preceded by an exposed code chunk that shows the computation and a
tidy table. Samples are the canonical no-duration-filter Studies 1–3 frame
(`s13`) and the strict Study-4 frame (`d4$s4`, *N* = 1,272). For convenience this
section builds the Study-4 object once.

```{r me-build-s4}
d4 <- build_s4_data(paths)              # strict S4 analytic frame + cached merges
s4 <- d4$s4                             # N = 1,272 (no-duration strict sample)
nrow(s4)
```

## Sample sizes {-}

```{r me-ns}
me_ns <- s13 |>
  dplyr::count(study_factor, name = "analytic_N") |>
  dplyr::mutate(study_factor = as.character(study_factor)) |>
  dplyr::bind_rows(tibble::tibble(study_factor = "Study 4 (frontier models)", analytic_N = nrow(s4)))
me_ns
total_N <- sum(me_ns$analytic_N)
total_N
```

```{r me-ns-vals, include=FALSE}
gN <- function(s) me_ns$analytic_N[me_ns$study_factor == s]
```

After the pre-registered exclusions, the analytic samples are *N* =
`r fmt_num(gN("Jailbroken"),0)` (Study 1, Jailbroken), *N* =
`r fmt_num(gN("Standard"),0)` (Study 2, Standard), *N* =
`r fmt_num(gN("Truth-Constrained"),0)` (Study 3, Truth-Constrained), and *N* =
`r fmt_num(gN("Study 4 (frontier models)"),0)` (Study 4, four frontier models),
for a total of *N* = `r fmt_num(total_N,0)` participants across the four studies.

### Screening funnel (reconstructed from the raw data) {-}

The analytic samples are the end of a deliberately stringent, identically-applied screening
protocol. We reconstruct the full funnel **from the raw exports in every study** so every
exclusion is auditable end to end: the entrant-level Qualtrics files for Studies 1–3 (the
`clean_*` analytic files begin only at the equivocality classifier, so the upstream
attention/bot screens are visible only in the raw entrants) and the raw Study-4 export.

```{r me-funnel-s13, message=FALSE}
suppressMessages({
  source(file.path(REPO_ROOT, "code", "R", "tables_dynamic.R"))   # std_row schema
  source(file.path(REPO_ROOT, "code", "R", "ext_funnel_s13.R"))   # S1-3 funnel from the raw entrants
})
fun13 <- compute_screening_funnel_s13_full(
  list(s13 = s13, paths = paths),
  file.path(REPO_ROOT, "data", "raw_qualtrics"))
funnel_s13 <- fun13 |>
  dplyr::filter(!is.na(statistic)) |>
  dplyr::transmute(order = statistic, Stage = term, study = model, n) |>
  tidyr::pivot_wider(names_from = study, values_from = n) |>
  dplyr::arrange(order) |> dplyr::select(-order)
kable(funnel_s13, caption = "Studies 1–3 screening funnel — rows retained at each stage, reconstructed from the raw entrant exports. The terminal stage equals the analytic sample; the cleaned analytic files begin at 'reached_equivocality'.")
```

```{r me-funnel-s4}
flow4 <- make_screening_flow(d4$s4_raw)
kable(flow4 |> dplyr::transmute(Stage = stage, `N remaining` = n_remaining,
                                `Excluded here` = excluded_at_step,
                                `% of raw` = round(100 * pct_of_raw, 1)),
      caption = "Study 4 screening funnel from the raw Qualtrics export. The terminal stage equals the strict analytic sample (N = 1,272).")
```

```{r me-funnel-vals, include=FALSE}
.an <- fun13 |> dplyr::filter(term == "analytic_sample") |> dplyr::arrange(model)
.eq <- fun13 |> dplyr::filter(term == "reached_equivocality") |> dplyr::arrange(model)
```

Both funnels reconcile exactly to the analytic samples (Studies 1–3:
`r paste(fmt_num(.an$n,0), collapse=" / ")`; Study 4: `r fmt_num(tail(flow4$n_remaining,1),0)`).
The cleaned Studies 1–3 files begin at the equivocality-classifier stage
(`r paste(fmt_num(.eq$n,0), collapse=" / ")` rows for Jailbroken / Standard / Truth-Constrained),
so the upstream attention and bot screens are recovered only from the raw entrant exports.

## Demographics {-}

The coded demographic items are read directly from the analytic frames: race is
the recoded `ethnicity` item (code 1 = White, non-Hispanic), gender the recoded
`gender_2`/`gender` item (code 2 = female), age the `age_2`/`age` item (clipped to
13–100 in Study 4 to drop out-of-range entries), and political ideology the 1–5
social-conservatism item. These mirror the recodes in the covariate-balance
generator (`ext_balance.R`).

```{r me-demographics}
demo_s13 <- function(d) tibble::tibble(
  pct_white  = mean(suppressWarnings(as.numeric(d$ethnicity)) == 1, na.rm = TRUE),
  pct_female = mean(suppressWarnings(as.numeric(d$gender_2)) == 2, na.rm = TRUE),
  mean_age   = mean(suppressWarnings(as.numeric(d$age_2)), na.rm = TRUE),
  mean_ideol = mean(suppressWarnings(as.numeric(d$social_conservatism)), na.rm = TRUE))
me_demo <- s13 |>
  dplyr::group_by(study = as.character(study_factor)) |>
  dplyr::group_modify(~ demo_s13(.x)) |>
  dplyr::ungroup()
s4_age <- { a <- suppressWarnings(as.numeric(s4$age)); ifelse(a >= 13 & a <= 100, a, NA_real_) }
me_demo <- dplyr::bind_rows(me_demo, tibble::tibble(
  study      = "Study 4 (frontier models)",
  pct_white  = mean(suppressWarnings(as.numeric(s4$ethnicity)) == 1, na.rm = TRUE),
  pct_female = mean(suppressWarnings(as.numeric(s4$gender)) == 2, na.rm = TRUE),
  mean_age   = mean(s4_age, na.rm = TRUE),
  mean_ideol = mean(suppressWarnings(as.numeric(s4$SocialConservatism)), na.rm = TRUE)))
kable(me_demo, digits = c(0, 3, 3, 1, 2),
      caption = "Sample demographics by study: % White (non-Hispanic), % female, mean age, mean social conservatism (1–5).")
```

```{r me-demo-vals, include=FALSE}
gd <- function(s, col) me_demo[[col]][me_demo$study == s]
```

Participants were predominantly White and slightly more than half female. The
share White ranged from `r pct(gd("Truth-Constrained","pct_white"),0)`
(Truth-Constrained) to `r pct(gd("Jailbroken","pct_white"),0)` (Jailbroken), and
the share female from `r pct(gd("Standard","pct_female"),0)` (Standard) to
`r pct(gd("Study 4 (frontier models)","pct_female"),0)` (Study 4). Samples skewed
older in the first two studies (mean age `r mn(gd("Jailbroken","mean_age"))` in
Study 1 and `r mn(gd("Standard","mean_age"))` in Study 2) and younger in the
later two (`r mn(gd("Truth-Constrained","mean_age"))` in Study 3 and
`r mn(gd("Study 4 (frontier models)","mean_age"))` in Study 4). Mean political
ideology sat near the scale midpoint in every study (social conservatism
`r mn2(gd("Jailbroken","mean_ideol"))`–`r mn2(gd("Study 4 (frontier models)","mean_ideol"))`
on the 1–5 scale).

## Attrition does not differ by condition {-}

Assignment to the bunking vs debunking arm is randomized within every study, so
post-randomization attrition should be independent of condition. For Studies 1–3
we model dropout — a missing post-conversation belief among the eligible pool
(equivocal focal description, baseline belief in the 25–75 window, a bunk/debunk
condition assigned) — as a logistic function of condition, and test it with a
likelihood-ratio chi-square (mirrors `compute_dropout_logistic_s13`).

Study 4 requires a sharper treatment, because non-completion there is **not** the
same thing as mid-conversation dropout. We classify every member of the eligible
pool (passed all strict screens through model assignment) from the continuously
saved partial-chat transcript (`__js_chatPartialData1`). This follows the
repository's substantive-versus-technical attrition-classification helper:

* **completed** — produced post-conversation outcomes (the analytic sample);
* **no successful initial callback** — no nonempty assistant response was stored;
  this can reflect participant departure at the chatbot transition or technical
  non-delivery and is not attributable at the row level;
* **initial AI, no new participant turn** — the initial response was stored but the
  participant did not type in the visible chat;
* **interactive chat, then lost** — at least one visible participant turn was
  stored before an unfinished conversation;
* **chat completed, outcomes missing** — the chat reached its end state or a full
  main transcript was saved, but post-conversation outcomes were absent.

The test that bears on direction-dependent selection is restricted to confirmed
initial delivery. Before seeing the initial AI response, participants had no way
to know whether they had been assigned to bunking or debunking.

```{r me-attrition-s13}
s13_study_map <- c("Study 1" = "Jailbroken", "Study 2" = "Standard", "Study 3" = "Truth-Constrained")
me_attr_s13 <- purrr::imap_dfr(s13_study_map, function(label, key) {
  d <- readr::read_csv(paths$s1s3_clean[[key]], show_col_types = FALSE, progress = FALSE) |>
    dplyr::mutate(
      .pre  = suppressWarnings(as.numeric(belief_rating_pre_4)),
      .post = suppressWarnings(as.numeric(belief_rating_post_4)),
      .pre_rc  = dplyr::if_else(category == "denies", 100 - .pre,  .pre),
      .post_rc = dplyr::if_else(category == "denies", 100 - .post, .post)) |>
    dplyr::filter(isEquivocal == TRUE, .pre_rc > 25, .pre_rc < 75,
                  condition %in% c("treatment_mid_bunk", "treatment_mid_debunk")) |>
    dplyr::mutate(dropout = as.integer(is.na(.post_rc)),
                  cond = factor(condition, levels = c("treatment_mid_bunk","treatment_mid_debunk")))
  f1 <- glm(dropout ~ cond, data = d, family = binomial())
  f0 <- glm(dropout ~ 1,    data = d, family = binomial())
  lr <- anova(f0, f1, test = "LRT")
  tibble::tibble(study = label, n_pool = nrow(d), dropout_rate = mean(d$dropout),
                 LRT_chisq = lr$Deviance[2], df = lr$Df[2], p = lr$`Pr(>Chi)`[2])
})
kable(me_attr_s13, digits = c(0, 0, 3, 3, 0, 3),
      caption = "Studies 1–3 differential attrition: dropout (missing post belief) ~ condition, LRT.")
```

```{r me-attrition-s4}
# Use the shared live recompute. parse_partial_chat_state() joins Qualtrics chunks
# and excludes the hidden initial `user` seed before the extension module creates
# these mutually exclusive states and confirmed-delivery tests.
me_loss_labels <- c(
  pool = "Pool", completed = "Completed outcomes",
  no_successful_callback_n = "No successful initial AI callback",
  initial_ai_no_user_n = "Initial AI; no new participant turn",
  interactive_chat_then_lost_n = "Interactive chat, then lost",
  chat_completed_outcomes_missing_n = "Chat completed; outcomes missing"
)
s4_loss_by_model <- ALL_NUMBERS |>
  dplyr::filter(block == "attrition_substantive_vs_technical",
                sample == "strict_n1272", !is.na(model),
                term %in% names(me_loss_labels)) |>
  dplyr::transmute(
    model,
    State = factor(me_loss_labels[term], levels = unname(me_loss_labels)),
    value = estimate
  ) |>
  tidyr::pivot_wider(names_from = State, values_from = value)
kable(s4_loss_by_model, digits = 0,
      caption = "Study 4 outcome status by model. No successful initial AI callback is causally ambiguous at the row level; it can represent participant departure at the chatbot transition or technical non-delivery.")

me_attr_s4 <- ALL_NUMBERS |>
  dplyr::filter(block == "attrition_substantive_vs_technical",
                sample == "strict_n1272", is.na(model),
                term %in% c("differential_attrition_by_model_LRT",
                            "postdelivery_attrition_by_direction_LRT",
                            "postdelivery_attrition_by_model_LRT")) |>
  dplyr::transmute(test = term, n, LRT_chisq = statistic, df = df_num, p = p_value)
me_attr_s4
```

```{r me-attr-s4-vals, include=FALSE}
me_a <- ALL_NUMBERS |>
  dplyr::filter(block == "attrition_substantive_vs_technical",
                sample == "strict_n1272", is.na(model))
me_get <- function(term, direction = NA_character_) {
  x <- me_a |> dplyr::filter(.data$term == term)
  if (is.na(direction)) x <- x |> dplyr::filter(is.na(.data$direction))
  else x <- x |> dplyr::filter(.data$direction == direction)
  x[1, ]
}
```

In Studies 1–3, post-randomization attrition was unrelated to condition: the
dropout-by-condition likelihood-ratio tests were all non-significant
(smallest *p* = `r P(min(me_attr_s13$p))`; overall dropout rates
`r pct(me_attr_s13$dropout_rate[1])`/`r pct(me_attr_s13$dropout_rate[2])`/`r pct(me_attr_s13$dropout_rate[3])`).

In Study 4, of the `r fmt_num(me_get("pool")$estimate,0)` participants who passed
every screen and were assigned a model, `r fmt_num(me_get("completed")$estimate,0)`
completed the primary outcomes. For
`r fmt_num(me_get("no_successful_callback_n")$estimate,0)` incomplete records, no
successful initial assistant callback was stored. This category cannot distinguish
participant departure at the chatbot transition from technical non-delivery. Its
large Gemini excess does identify an aggregate technical component: the requested
`google/gemini-3-pro-preview` endpoint was removed from OpenRouter during fielding,
producing a complete Gemini callback blackout before the survey switched to
`google/gemini-3.1-pro-preview`.

The relevant condition test begins after confirmed initial AI delivery, because
participants could not know whether they had been assigned to bunking or debunking
before that response appeared. Postdelivery outcome attrition was
`r pct(me_get("postdelivery_attrition_rate", "bunk")$estimate)` after bunking versus
`r pct(me_get("postdelivery_attrition_rate", "debunk")$estimate)` after debunking;
the difference was significant ($\chi^2$(`r me_get("postdelivery_attrition_by_direction_LRT")$df_num`) =
`r mn2(me_get("postdelivery_attrition_by_direction_LRT")$statistic)`,
`r P(me_get("postdelivery_attrition_by_direction_LRT")$p_value)`). Postdelivery
attrition did not differ by model ($\chi^2$(`r me_get("postdelivery_attrition_by_model_LRT")$df_num`) =
`r mn2(me_get("postdelivery_attrition_by_model_LRT")$statistic)`,
`r P(me_get("postdelivery_attrition_by_model_LRT")$p_value)`). Subsequent loss was
therefore uncommon and flat across models, but modestly higher following a
delivered bunking response; selective completion by assigned direction cannot be
ruled out.

## Conversation descriptives {-}

Engagement is summarized from the transcripts so that exposure can be compared
across studies. For Studies 1–3, message counts and word counts are computed from
the parsed `content_user_*` / `content_assistant_*` columns of the analytic frame
(messages = non-empty cells; words = whitespace tokens), exactly as in
`compute_engagement`. Conversation duration is the elapsed time of the last
message (`createdAt`, seconds from conversation start) parsed from the
`chathistory0*` JSON. Substantive claims per conversation is the mean number of
fact-checked claims per conversation from the claim-extraction census
(`claim_counts_full`).

```{r me-convo-s13}
ucols <- grep("^content_user_[0-9]+$",      names(s13), value = TRUE)
acols <- grep("^content_assistant_[0-9]+$", names(s13), value = TRUE)
ne  <- function(x) !is.na(x) & nzchar(trimws(as.character(x)))
wc  <- function(x) vapply(ifelse(is.na(x), "", as.character(x)),
                          function(s) { s <- trimws(s); if (!nzchar(s)) 0L else length(strsplit(s, "\\s+")[[1]]) },
                          integer(1), USE.NAMES = FALSE)
# parse one chathistory cell -> messages list (mirrors .ext_parse_s4_transcript for S1-3 single-string form)
parse_chat <- function(ch) {
  s <- ch
  if (startsWith(s, '"')) s <- substr(s, 2, nchar(s))
  if (endsWith(s,   '"')) s <- substr(s, 1, nchar(s) - 1)
  s <- gsub('\\\\"', '"', s); s <- gsub('\\\\\\\\', '\\\\', s)
  p <- tryCatch(jsonlite::fromJSON(s, simplifyVector = FALSE), error = function(e) NULL)
  if (is.character(p)) p <- tryCatch(jsonlite::fromJSON(p, simplifyVector = FALSE), error = function(e) NULL)
  if (is.list(p) && !is.null(p$messages)) p$messages else NULL
}
chcols13 <- intersect(paste0("chathistory0", 1:5), names(s13))
conv_dur_min <- function(row) {
  msgs <- parse_chat(paste0(unlist(row[chcols13]), collapse = ""))
  if (is.null(msgs)) return(NA_real_)
  ca <- suppressWarnings(as.numeric(vapply(msgs, function(m) if (is.null(m$createdAt)) NA_character_ else as.character(m$createdAt), character(1))))
  if (all(is.na(ca))) return(NA_real_)
  (max(ca, na.rm = TRUE) - min(ca, na.rm = TRUE)) / 60
}
s13d <- s13 |>
  dplyr::mutate(.n_user = rowSums(dplyr::across(dplyr::all_of(ucols), ne)),
                .w_user = rowSums(dplyr::across(dplyr::all_of(ucols), wc)),
                .w_asst = rowSums(dplyr::across(dplyr::all_of(acols), wc)))
s13d$.cdur <- vapply(seq_len(nrow(s13d)), function(i) conv_dur_min(s13d[i, ]), numeric(1))
me_convo_s13 <- s13d |>
  dplyr::group_by(study = as.character(study_factor)) |>
  dplyr::summarise(mean_user_msgs = mean(.n_user), mean_user_words = mean(.w_user),
                   mean_asst_words = mean(.w_asst), median_dur_min = median(.cdur, na.rm = TRUE),
                   .groups = "drop")
me_convo_s13
```

```{r me-convo-s4}
# robust reconstruction of an S4 chunked-JSON transcript (mirrors .ext_parse_s4_transcript):
# chathistory0* are successive fragments of one escaped JSON string literal.
parse_s4_transcript <- function(chunks) {
  chunks <- chunks[!is.na(chunks) & nzchar(chunks)]
  if (!length(chunks)) return(NULL)
  inner <- vapply(chunks, function(s) {
    if (startsWith(s, '"')) s <- substr(s, 2, nchar(s))
    if (endsWith(s, '"'))   s <- substr(s, 1, nchar(s) - 1)
    s }, character(1), USE.NAMES = FALSE)
  j <- paste0(inner, collapse = "")
  j <- gsub('\\\\"', '"', j); j <- gsub('\\\\\\\\', '\\\\', j)
  p <- tryCatch(jsonlite::fromJSON(j, simplifyVector = FALSE), error = function(e) NULL)
  if (is.character(p)) p <- tryCatch(jsonlite::fromJSON(p, simplifyVector = FALSE), error = function(e) NULL)
  if (is.list(p) && !is.null(p$messages)) p$messages else NULL
}
chat_cols <- intersect(paste0("chathistory0", 1:5), names(s4))
s4_msgs <- lapply(seq_len(nrow(s4)), function(i) parse_s4_transcript(as.character(unlist(s4[i, chat_cols]))))
s4_metrics <- purrr::map_dfr(s4_msgs, function(m) {
  if (is.null(m)) return(tibble::tibble(n_user = NA, w_asst = NA, dur = NA_real_))
  roles <- vapply(m, function(x) if (!is.null(x$role)) x$role else NA_character_, character(1))
  cont  <- vapply(m, function(x) if (!is.null(x$content)) as.character(x$content) else "", character(1))
  ca    <- suppressWarnings(as.numeric(vapply(m, function(x) if (is.null(x$createdAt)) NA_character_ else as.character(x$createdAt), character(1))))
  tibble::tibble(n_user = sum(roles == "user", na.rm = TRUE),
                 w_asst = sum(wc(cont[roles == "assistant"])),
                 dur = if (all(is.na(ca))) NA_real_ else (max(ca, na.rm = TRUE) - min(ca, na.rm = TRUE)) / 60)
})
me_convo_s4 <- tibble::tibble(
  study = "Study 4 (frontier models)",
  mean_user_msgs = mean(s4_metrics$n_user, na.rm = TRUE),
  mean_asst_words = mean(s4_metrics$w_asst, na.rm = TRUE),
  median_dur_min = median(s4_metrics$dur, na.rm = TRUE))
me_convo_s4
```

```{r me-substantive-claims}
# Load the assembled all_numbers table using the same robust pattern as the SI docs
# (supplement/Bunkbot_SI.Rmd): reuse the cached engine build if present and readable,
# otherwise build it live (API-free) and cache it. This makes the render work on a
# fresh checkout, where output/ is empty (the .rds is a gitignored build artifact).
source(file.path(REPO_ROOT, "code", "R", "build_all_numbers.R"))
nums_an <- load_or_build_all_numbers(REPO_ROOT)$all_numbers
me_subst <- nums_an |>
  dplyr::filter(block == "claim_counts_full", term == "mean_factchecked_per_conv",
                sample == "cached_api_claims",
                model %in% c("Study1", "Study2", "Study3", "Study4"), is.na(direction)) |>
  dplyr::transmute(study = dplyr::recode(model, Study1 = "Jailbroken", Study2 = "Standard",
                                         Study3 = "Truth-Constrained", Study4 = "Study 4 (frontier models)"),
                   substantive_claims_per_conv = estimate)
me_subst
```

```{r me-convo-vals, include=FALSE}
gc13 <- function(s, col) me_convo_s13[[col]][me_convo_s13$study == s]
gsub_ <- function(s) me_subst$substantive_claims_per_conv[me_subst$study == s]
```

Conversations were short multi-turn exchanges. Participants sent on average
`r mn(gc13("Jailbroken","mean_user_msgs"))`–`r mn(gc13("Standard","mean_user_msgs"))`
messages in Studies 1–3 and `r mn(me_convo_s4$mean_user_msgs)` in Study 4, writing
roughly `r fmt_num(gc13("Jailbroken","mean_user_words"),0)`–`r fmt_num(gc13("Truth-Constrained","mean_user_words"),0)`
words. The assistant produced far more text, and markedly more under the
later-study models: a mean of `r fmt_num(gc13("Jailbroken","mean_asst_words"),0)`
assistant words per conversation in Study 1, about
`r fmt_num(gc13("Standard","mean_asst_words"),0)` and
`r fmt_num(gc13("Truth-Constrained","mean_asst_words"),0)` in Studies 2 and 3, and
`r fmt_num(me_convo_s4$mean_asst_words,0)` in Study 4. Each conversation contained
`r mn(gsub_("Jailbroken"))`/`r mn(gsub_("Standard"))`/`r mn(gsub_("Truth-Constrained"))`/`r mn(gsub_("Study 4 (frontier models)"))`
fact-checked substantive claims (Studies 1–4). Median conversation duration was
`r mn(gc13("Jailbroken","median_dur_min"))`/`r mn(gc13("Standard","median_dur_min"))`/`r mn(gc13("Truth-Constrained","median_dur_min"))`
minutes in Studies 1–3 and `r mn(me_convo_s4$median_dur_min)` minutes in Study 4.

## Claim extraction and fact-checking volume {-}

The veracity pipeline extracted discrete factual claims from every assistant
turn and fact-checked the substantive ones. The reported totals are restricted to
the analytic samples. For Studies 1–3 the per-conversation extraction files
(`*_nfacts_veracity.jsonl`) are read and a claim counts as fact-checked when its
`veracity_request_status == "success"`; we join to the analytic frame by
`ResponseId`. For Study 4 the per-conversation claim census in the master dataset
(`all_claims_total`, `factcheck_scored_n`) is summed over the strict sample
(mirrors `compute_claim_counts_full`).

```{r me-claims-s13}
me_claims_s13 <- purrr::map_dfr(1:3, function(st) {
  ls   <- readLines(paths$nfacts[[paste0("Study", st)]], warn = FALSE)
  recs <- lapply(ls, jsonlite::fromJSON)
  rid  <- vapply(recs, function(x) if (is.null(x$ResponseId)) NA_character_ else as.character(x$ResponseId), character(1))
  vok  <- vapply(recs, function(x) !is.null(x$veracity_request_status) &&
                   identical(as.character(x$veracity_request_status), "success"), logical(1))
  tibble::tibble(study = c("Jailbroken","Standard","Truth-Constrained")[st],
                 fact_checked_analytic = sum(vok & rid %in% s13$response_id))
})
me_claims_s13
s13_factchecked_total <- sum(me_claims_s13$fact_checked_analytic)
s13_factchecked_total
```

```{r me-claims-s4}
ids4 <- as.character(s4$ResponseId)
idcol <- if ("conversation_id" %in% names(d4$s4_master)) "conversation_id" else "ResponseId"
m_anal <- d4$s4_master[as.character(d4$s4_master[[idcol]]) %in% ids4, ]
me_claims_s4 <- tibble::tibble(
  s4_conversations = nrow(m_anal),
  extracted  = sum(m_anal$all_claims_total,   na.rm = TRUE),
  factchecked = sum(m_anal$factcheck_scored_n, na.rm = TRUE))
me_claims_s4
```

Across Studies 1–3, `r fmt_num(s13_factchecked_total,0)` extracted claims in the
analytic-sample conversations received a fact-check score. In Study 4 the four
frontier models produced `r fmt_num(me_claims_s4$extracted,0)` extracted claims
across the `r fmt_num(me_claims_s4$s4_conversations,0)` analytic conversations, of
which `r fmt_num(me_claims_s4$factchecked,0)` were fact-checked.

## Five-rater post-stance reliability {-}

Every Study-4 pre- and post-conversation social-media post is scored by a
five-model consensus rater drawn from five developers
(`claude-sonnet-4.6`, `gpt-5.2`, `gemini-3.1-pro`, `grok-4.3`, `deepseek-v3.2`).
Reliability is computed live from the per-rater 0–100 `stance_score` columns in
the consolidated stance file: the interval Krippendorff's $\alpha$, the two-way ICC(2,k),
and the mean pairwise inter-rater correlation. Off-topic and declined post rates
are the post-conversation consensus `focal_relevance == "other_topic"` and
`response_type == "declines_to_post"` shares (cross-checked against the
`stance_reliability` block).

```{r me-stance-reliability}
sv <- readr::read_csv(paths$s4_stance_v2, show_col_types = FALSE, progress = FALSE)
raters <- c("claude_stance_score","gpt_stance_score","gemini_stance_score","grok_stance_score","deepseek_stance_score")
M <- as.matrix(sv[, raters]); storage.mode(M) <- "numeric"
alpha_interval <- irr::kripp.alpha(t(M), method = "interval")$value
icc2k <- { ic <- psych::ICC(M, missing = FALSE, lmer = FALSE); ic$results$ICC[ic$results$type == "ICC2k"] }
cm <- cor(M, use = "pairwise.complete.obs"); pairwise_r <- mean(cm[upper.tri(cm)])
me_rel <- tibble::tibble(
  n_rated_items = nrow(sv),
  krippendorff_alpha_interval = alpha_interval,
  ICC_2k = icc2k,
  mean_pairwise_r = pairwise_r)
me_rel
```

```{r me-stance-composition}
post <- sv |> dplyr::filter(timepoint == "post")
me_comp <- tibble::tibble(
  n_post = nrow(post),
  pct_off_topic = mean(post$consensus_focal_relevance == "other_topic", na.rm = TRUE),
  pct_declining = mean(post$consensus_response_type == "declines_to_post", na.rm = TRUE))
me_comp
```

Inter-rater agreement was high: across the `r fmt_num(me_rel$n_rated_items,0)`
rated post items the interval Krippendorff's $\alpha$ was
`r mn2(me_rel$krippendorff_alpha_interval)`, the two-way ICC(2,k) was
`r mn2(me_rel$ICC_2k)`, and the mean pairwise correlation among raters was
`r mn2(me_rel$mean_pairwise_r)`. Among post-conversation posts, only
`r pct(me_comp$pct_off_topic)` were scored as off-topic and
`r pct(me_comp$pct_declining)` as declining to post; these were flagged for the
composition tables and assigned the neutral midpoint in the direction-coded
continuous analyses. Annotation inter-rater agreement for the APE compliance
labels — reported as 84% agreement and a Cohen's $\kappa$ of 0.66 — is carried over from
the APE instrument's original validation (Kowal et al., 2025, arXiv:2506.02873)
and is not recomputed here.

## Topic taxonomy {-}

To ask whether persuasion effects concentrate in recognizable conspiracy
families, the pooled focal descriptions were clustered into topics with a fixed,
fully specified embedding → PCA → HDBSCAN procedure. The number of interpretable
topics and the size of the residual "Mixed / Unclassified" bucket are read from
the canonical cached assignment table (`topic_assignments.csv`; cross-checked
against the `topic_effects` block).

```{r me-topics}
ta <- readr::read_csv(paths$topic_assignments, show_col_types = FALSE, progress = FALSE)
n_topics_named <- ta |>
  dplyr::filter(topic != "Mixed / Unclassified") |>
  dplyr::distinct(topic) |> nrow()
pct_mixed <- mean(ta$topic == "Mixed / Unclassified")
me_topics <- tibble::tibble(n_corpus = nrow(ta), n_named_topics = n_topics_named, pct_mixed = pct_mixed)
me_topics
```

The selected solution yields `r fmt_num(me_topics$n_named_topics,0)` named,
interpretable density clusters (e.g. "Moon Landing Hoax", "9/11 Inside Job",
"Vaccines Cause Autism") plus a residual Mixed / Unclassified bucket that absorbs
HDBSCAN noise points and any cluster below the size floor; the Mixed bucket
accounts for `r pct(me_topics$pct_mixed)` of the
`r fmt_num(me_topics$n_corpus,0)`-description pooled corpus.
