# Methods (reproducible quantities)

The Methods narrative is reproduced as in the manuscript; every **quantitative**
claim is recomputed live from the raw data and the cached LLM-derived files. Each
quantity is preceded by an exposed code chunk that shows the computation and a
tidy table. Samples are the canonical no-duration-filter Studies 1–3 frame
(`s13`) and the strict Study-4 frame (`d4$s4`, *N* = 1,272). For convenience this
section builds the Study-4 object once.

```{r me-build-s4}
d4 <- build_s4_data(paths)              # strict S4 analytic frame + cached merges
s4 <- d4$s4                             # N = 1,272 (no-duration strict sample)
nrow(s4)
```

## Sample sizes {-}

```{r me-ns}
me_ns <- s13 |>
  dplyr::count(study_factor, name = "analytic_N") |>
  dplyr::mutate(study_factor = as.character(study_factor)) |>
  dplyr::bind_rows(tibble::tibble(study_factor = "Study 4 (frontier models)", analytic_N = nrow(s4)))
me_ns
total_N <- sum(me_ns$analytic_N)
total_N
```

```{r me-ns-vals, include=FALSE}
gN <- function(s) me_ns$analytic_N[me_ns$study_factor == s]
```

After the pre-registered exclusions, the analytic samples are *N* =
`r fmt_num(gN("Jailbroken"),0)` (Study 1, Jailbroken), *N* =
`r fmt_num(gN("Standard"),0)` (Study 2, Standard), *N* =
`r fmt_num(gN("Truth-Constrained"),0)` (Study 3, Truth-Constrained), and *N* =
`r fmt_num(gN("Study 4 (frontier models)"),0)` (Study 4, four frontier models),
for a total of *N* = `r fmt_num(total_N,0)` participants across the four studies.

### Screening funnel (reconstructed from the raw data) {-}

The analytic samples are the end of a deliberately stringent, identically-applied screening
protocol. We reconstruct the full funnel **from the raw exports in every study** so every
exclusion is auditable end to end: the entrant-level Qualtrics files for Studies 1–3 (the
`clean_*` analytic files begin only at the equivocality classifier, so the upstream
attention/bot screens are visible only in the raw entrants) and the raw Study-4 export.

```{r me-funnel-s13, message=FALSE}
suppressMessages({
  source(file.path(REPO_ROOT, "code", "R", "tables_dynamic.R"))   # std_row schema
  source(file.path(REPO_ROOT, "code", "R", "ext_funnel_s13.R"))   # S1-3 funnel from the raw entrants
})
fun13 <- compute_screening_funnel_s13_full(
  list(s13 = s13, paths = paths),
  file.path(REPO_ROOT, "data", "raw_qualtrics"))
funnel_s13 <- fun13 |>
  dplyr::filter(!is.na(statistic)) |>
  dplyr::transmute(order = statistic, Stage = term, study = model, n) |>
  tidyr::pivot_wider(names_from = study, values_from = n) |>
  dplyr::arrange(order) |> dplyr::select(-order)
kable(funnel_s13, caption = "Studies 1–3 screening funnel — rows retained at each stage, reconstructed from the raw entrant exports. The terminal stage equals the analytic sample; the cleaned analytic files begin at 'reached_equivocality'.")
```

```{r me-funnel-s4}
flow4 <- make_screening_flow(d4$s4_raw)
kable(flow4 |> dplyr::transmute(Stage = stage, `N remaining` = n_remaining,
                                `Excluded here` = excluded_at_step,
                                `% of raw` = round(100 * pct_of_raw, 1)),
      caption = "Study 4 screening funnel from the raw Qualtrics export. The terminal stage equals the strict analytic sample (N = 1,272).")
```

```{r me-funnel-vals, include=FALSE}
.an <- fun13 |> dplyr::filter(term == "analytic_sample") |> dplyr::arrange(model)
.eq <- fun13 |> dplyr::filter(term == "reached_equivocality") |> dplyr::arrange(model)
```

Both funnels reconcile exactly to the analytic samples (Studies 1–3:
`r paste(fmt_num(.an$n,0), collapse=" / ")`; Study 4: `r fmt_num(tail(flow4$n_remaining,1),0)`).
The cleaned Studies 1–3 files begin at the equivocality-classifier stage
(`r paste(fmt_num(.eq$n,0), collapse=" / ")` rows for Jailbroken / Standard / Truth-Constrained),
so the upstream attention and bot screens are recovered only from the raw entrant exports.

## Demographics {-}

The coded demographic items are read directly from the analytic frames: race is
the recoded `ethnicity` item (code 1 = White, non-Hispanic), gender the recoded
`gender_2`/`gender` item (code 2 = female), age the `age_2`/`age` item (clipped to
13–100 in Study 4 to drop out-of-range entries), and political ideology the 1–5
social-conservatism item. These mirror the recodes in the covariate-balance
generator (`ext_balance.R`).

```{r me-demographics}
demo_s13 <- function(d) tibble::tibble(
  pct_white  = mean(suppressWarnings(as.numeric(d$ethnicity)) == 1, na.rm = TRUE),
  pct_female = mean(suppressWarnings(as.numeric(d$gender_2)) == 2, na.rm = TRUE),
  mean_age   = mean(suppressWarnings(as.numeric(d$age_2)), na.rm = TRUE),
  mean_ideol = mean(suppressWarnings(as.numeric(d$social_conservatism)), na.rm = TRUE))
me_demo <- s13 |>
  dplyr::group_by(study = as.character(study_factor)) |>
  dplyr::group_modify(~ demo_s13(.x)) |>
  dplyr::ungroup()
s4_age <- { a <- suppressWarnings(as.numeric(s4$age)); ifelse(a >= 13 & a <= 100, a, NA_real_) }
me_demo <- dplyr::bind_rows(me_demo, tibble::tibble(
  study      = "Study 4 (frontier models)",
  pct_white  = mean(suppressWarnings(as.numeric(s4$ethnicity)) == 1, na.rm = TRUE),
  pct_female = mean(suppressWarnings(as.numeric(s4$gender)) == 2, na.rm = TRUE),
  mean_age   = mean(s4_age, na.rm = TRUE),
  mean_ideol = mean(suppressWarnings(as.numeric(s4$SocialConservatism)), na.rm = TRUE)))
kable(me_demo, digits = c(0, 3, 3, 1, 2),
      caption = "Sample demographics by study: % White (non-Hispanic), % female, mean age, mean social conservatism (1–5).")
```

```{r me-demo-vals, include=FALSE}
gd <- function(s, col) me_demo[[col]][me_demo$study == s]
```

Participants were predominantly White and slightly more than half female. The
share White ranged from `r pct(gd("Truth-Constrained","pct_white"),0)`
(Truth-Constrained) to `r pct(gd("Jailbroken","pct_white"),0)` (Jailbroken), and
the share female from `r pct(gd("Standard","pct_female"),0)` (Standard) to
`r pct(gd("Study 4 (frontier models)","pct_female"),0)` (Study 4). Samples skewed
older in the first two studies (mean age `r mn(gd("Jailbroken","mean_age"))` in
Study 1 and `r mn(gd("Standard","mean_age"))` in Study 2) and younger in the
later two (`r mn(gd("Truth-Constrained","mean_age"))` in Study 3 and
`r mn(gd("Study 4 (frontier models)","mean_age"))` in Study 4). Mean political
ideology sat near the scale midpoint in every study (social conservatism
`r mn2(gd("Jailbroken","mean_ideol"))`–`r mn2(gd("Study 4 (frontier models)","mean_ideol"))`
on the 1–5 scale).

## Attrition does not differ by condition {-}

Assignment to the bunking vs debunking arm is randomized within every study, so
post-randomization attrition should be independent of condition. For Studies 1–3
we model dropout — a missing post-conversation belief among the eligible pool
(equivocal focal description, baseline belief in the 25–75 window, a bunk/debunk
condition assigned) — as a logistic function of condition, and test it with a
likelihood-ratio chi-square (mirrors `compute_dropout_logistic_s13`).

Study 4 requires a sharper treatment, because non-completion there is **not** the
same thing as mid-conversation dropout. We classify every member of the eligible
pool (passed all strict screens through model assignment) from the continuously
saved partial-chat transcript (`__js_chatPartialData1`), mirroring
`compute_attrition_substantive_vs_technical`:

* **completed** — produced post-conversation outcomes (the analytic sample);
* **substantive mid-conversation dropout** — the model replied at least once and the
  participant *then* abandoned the study (a genuine, partial transcript exists);
* **no-message** — the participant reached the chat but never sent a message, so **no
  partial transcript was ever saved**. This is a technical chat-load failure (not
  disengagement), overwhelmingly concentrated in Gemini, whose endpoint Google was
  deprecating during fielding.

The test that bears on whether the analytic samples are differentially shaped is the
**substantive** one — among participants who actually began the conversation, does
mid-conversation dropout vary by assigned direction or by model? The technical
no-message loss is excluded from it.

```{r me-attrition-s13}
s13_study_map <- c("Study 1" = "Jailbroken", "Study 2" = "Standard", "Study 3" = "Truth-Constrained")
me_attr_s13 <- purrr::imap_dfr(s13_study_map, function(label, key) {
  d <- readr::read_csv(paths$s1s3_clean[[key]], show_col_types = FALSE, progress = FALSE) |>
    dplyr::mutate(
      .pre  = suppressWarnings(as.numeric(belief_rating_pre_4)),
      .post = suppressWarnings(as.numeric(belief_rating_post_4)),
      .pre_rc  = dplyr::if_else(category == "denies", 100 - .pre,  .pre),
      .post_rc = dplyr::if_else(category == "denies", 100 - .post, .post)) |>
    dplyr::filter(isEquivocal == TRUE, .pre_rc > 25, .pre_rc < 75,
                  condition %in% c("treatment_mid_bunk", "treatment_mid_debunk")) |>
    dplyr::mutate(dropout = as.integer(is.na(.post_rc)),
                  cond = factor(condition, levels = c("treatment_mid_bunk","treatment_mid_debunk")))
  f1 <- glm(dropout ~ cond, data = d, family = binomial())
  f0 <- glm(dropout ~ 1,    data = d, family = binomial())
  lr <- anova(f0, f1, test = "LRT")
  tibble::tibble(study = label, n_pool = nrow(d), dropout_rate = mean(d$dropout),
                 LRT_chisq = lr$Deviance[2], df = lr$Df[2], p = lr$`Pr(>Chi)`[2])
})
kable(me_attr_s13, digits = c(0, 0, 3, 3, 0, 3),
      caption = "Studies 1–3 differential attrition: dropout (missing post belief) ~ condition, LRT.")
```

```{r me-attrition-s4}
# S4 pool = entrants passing all strict screens through model assignment. The raw
# incomplete-outcome loss mixes a technical chat-load failure (a chat that never
# loaded, concentrated in Gemini) with genuine mid-conversation disengagement. The
# test that bears on whether the analytic samples are differentially shaped is the
# SUBSTANTIVE one: among participants who actually began the conversation (the
# model replied >= once), does mid-conversation dropout vary by condition? We
# classify each non-completer from the continuously-saved partial-chat log
# (__js_chatPartialData1), mirroring compute_attrition_substantive_vs_technical.
parse_partial_counts <- function(cell) {
  if (is.na(cell) || !nzchar(trimws(cell))) return(c(0L, 0L))
  ms <- tryCatch({
    arr <- jsonlite::fromJSON(cell, simplifyVector = TRUE); acc <- list()
    for (s in arr) { o <- tryCatch(jsonlite::fromJSON(s, simplifyDataFrame = FALSE), error = function(e) NULL)
                     if (!is.null(o$messages)) acc <- c(acc, o$messages) }
    acc }, error = function(e) NULL)
  if (is.null(ms) || !length(ms)) return(c(0L, 0L))
  role <- vapply(ms, function(m) if (is.null(m$role)) "" else m$role, character(1))
  cont <- vapply(ms, function(m) if (is.null(m$content)) "" else paste(as.character(m$content), collapse = ""), character(1))
  c(sum(role == "user" & nchar(trimws(cont)) > 0L), sum(role == "assistant" & nchar(trimws(cont)) > 0L))
}
s4_pool <- d4$s4_raw |>
  dplyr::filter(attention_pass, berry_pass, dedup_pass, condition_assigned, model_assigned,
                pre_belief_present, pre_post_present, equivocal_pass, !invalid_claim,
                nonempty(direction), belief_window_pass) |>
  dplyr::mutate(completed = post_outcomes_present,
                direction = factor(direction, levels = c("bunk","debunk")),
                model = model_pooled_label(modelName))
# Parse the partial-chat log for non-completers: count non-empty user messages and
# model replies. A conversation that "began" has >=1 model reply.
s4_pool$n_user <- 0L; s4_pool$n_ai <- 0L
inc <- which(!s4_pool$completed)
if (length(inc)) {
  pcc <- t(vapply(s4_pool[["__js_chatPartialData1"]][inc], parse_partial_counts, integer(2)))
  s4_pool$n_user[inc] <- pcc[, 1]; s4_pool$n_ai[inc] <- pcc[, 2]
}
s4_pool <- s4_pool |>
  dplyr::mutate(loss_class = dplyr::case_when(
    completed   ~ "completed",
    n_ai > 0    ~ "substantive_midchat",    # model replied >=1x, then participant abandoned
    n_user > 0  ~ "technical_no_reply",     # sent a message, model never replied (~0)
    TRUE        ~ "no_message"))            # reached chat, never sent a message -> NO partial transcript (technical)

# Loss classification BY MODEL (+ pooled), with the substantive dropout rate among
# those who engaged (completed + substantive_midchat).
s4_loss_by_model <- s4_pool |>
  dplyr::count(model, loss_class) |>
  tidyr::pivot_wider(names_from = loss_class, values_from = n, values_fill = 0)
for (c in c("completed","substantive_midchat","technical_no_reply","no_message"))
  if (!c %in% names(s4_loss_by_model)) s4_loss_by_model[[c]] <- 0L
s4_loss_by_model <- s4_loss_by_model |>
  dplyr::transmute(model, pool = completed + substantive_midchat + technical_no_reply + no_message,
                   completed, no_message_technical = no_message, substantive_midchat,
                   substantive_rate = substantive_midchat / (completed + substantive_midchat)) |>
  dplyr::bind_rows(dplyr::summarise(s4_pool |> dplyr::mutate(lc = loss_class),
    model = "ALL", pool = dplyr::n(), completed = sum(lc == "completed"),
    no_message_technical = sum(lc == "no_message"), substantive_midchat = sum(lc == "substantive_midchat"),
    substantive_rate = sum(lc == "substantive_midchat") / sum(lc %in% c("completed","substantive_midchat"))))
kable(s4_loss_by_model, digits = c(0,0,0,0,0,3),
      caption = "Study 4 conversation-pool loss classification, by model. no_message = reached the chat but never sent a message (no partial transcript saved; technical chat-load failure). substantive_midchat = a partial transcript exists (model replied) but the participant abandoned mid-conversation.")

# Tests. Raw non-completion (includes technical loss) by model + by direction; and
# the SUBSTANTIVE mid-conversation dropout among the engaged, by direction + by model.
s4_pool <- s4_pool |> dplyr::mutate(incomplete = as.integer(!completed))
raw_mod <- anova(glm(incomplete ~ 1, data = s4_pool, family = binomial()),
                 glm(incomplete ~ model, data = s4_pool, family = binomial()), test = "LRT")
engaged <- s4_pool |> dplyr::filter(loss_class %in% c("completed","substantive_midchat")) |>
  dplyr::mutate(sub_dropout = as.integer(loss_class == "substantive_midchat"))
sub_dir <- anova(glm(sub_dropout ~ 1, data = engaged, family = binomial()),
                 glm(sub_dropout ~ direction, data = engaged, family = binomial()), test = "LRT")
sub_mod <- anova(glm(sub_dropout ~ 1, data = engaged, family = binomial()),
                 glm(sub_dropout ~ model, data = engaged, family = binomial()), test = "LRT")
me_attr_s4 <- tibble::tibble(
  test = c("raw non-completion ~ model (incl. technical loss)",
           "substantive mid-conversation dropout ~ direction (engaged)",
           "substantive mid-conversation dropout ~ model (engaged)"),
  n = c(nrow(s4_pool), nrow(engaged), nrow(engaged)),
  LRT_chisq = c(raw_mod$Deviance[2], sub_dir$Deviance[2], sub_mod$Deviance[2]),
  df = c(raw_mod$Df[2], sub_dir$Df[2], sub_mod$Df[2]),
  p = c(raw_mod$`Pr(>Chi)`[2], sub_dir$`Pr(>Chi)`[2], sub_mod$`Pr(>Chi)`[2]))
me_attr_s4
```

```{r me-attr-s4-vals, include=FALSE}
A   <- s4_loss_by_model |> dplyr::filter(model == "ALL")
gem <- s4_loss_by_model |> dplyr::filter(model == "Gemini")
sdir <- me_attr_s4[2, ]; smod <- me_attr_s4[3, ]
```

In Studies 1–3, post-randomization attrition was unrelated to condition: the
dropout-by-condition likelihood-ratio tests were all non-significant
(smallest *p* = `r P(min(me_attr_s13$p))`; overall dropout rates
`r pct(me_attr_s13$dropout_rate[1])`/`r pct(me_attr_s13$dropout_rate[2])`/`r pct(me_attr_s13$dropout_rate[3])`).

In Study 4, of the `r fmt_num(A$pool,0)` participants who passed every screen and were
assigned a model, `r fmt_num(A$completed,0)` completed. The remaining loss is
overwhelmingly **technical, not disengagement**: `r fmt_num(A$no_message_technical,0)`
reached the chat but never sent a message — leaving no partial transcript — and these
are concentrated in Gemini (`r fmt_num(gem$no_message_technical,0)` of
`r fmt_num(gem$pool,0)`), whose endpoint was being deprecated during fielding (raw
non-completion thus differs sharply by model, χ²(`r me_attr_s4$df[1]`) =
`r mn2(me_attr_s4$LRT_chisq[1])`, `r P(me_attr_s4$p[1])`). Only
`r fmt_num(A$substantive_midchat,0)` participants across all four models had a partial
transcript (the model replied) and then abandoned the conversation
(`r pct(A$substantive_rate)` of the `r fmt_num(sdir$n,0)` who engaged). This genuine
mid-conversation dropout did not differ by assigned direction
(χ²(`r sdir$df`) = `r mn2(sdir$LRT_chisq)`, `r P(sdir$p)`) or by model
(χ²(`r smod$df`) = `r mn2(smod$LRT_chisq)`, `r P(smod$p)`). The analytic samples are
therefore not differentially shaped by condition- or model-dependent disengagement.

## Conversation descriptives {-}

Engagement is summarized from the transcripts so that exposure can be compared
across studies. For Studies 1–3, message counts and word counts are computed from
the parsed `content_user_*` / `content_assistant_*` columns of the analytic frame
(messages = non-empty cells; words = whitespace tokens), exactly as in
`compute_engagement`. Conversation duration is the elapsed time of the last
message (`createdAt`, seconds from conversation start) parsed from the
`chathistory0*` JSON. Substantive claims per conversation is the mean number of
fact-checked claims per conversation from the claim-extraction census
(`claim_counts_full`).

```{r me-convo-s13}
ucols <- grep("^content_user_[0-9]+$",      names(s13), value = TRUE)
acols <- grep("^content_assistant_[0-9]+$", names(s13), value = TRUE)
ne  <- function(x) !is.na(x) & nzchar(trimws(as.character(x)))
wc  <- function(x) vapply(ifelse(is.na(x), "", as.character(x)),
                          function(s) { s <- trimws(s); if (!nzchar(s)) 0L else length(strsplit(s, "\\s+")[[1]]) },
                          integer(1), USE.NAMES = FALSE)
# parse one chathistory cell -> messages list (mirrors .ext_parse_s4_transcript for S1-3 single-string form)
parse_chat <- function(ch) {
  s <- ch
  if (startsWith(s, '"')) s <- substr(s, 2, nchar(s))
  if (endsWith(s,   '"')) s <- substr(s, 1, nchar(s) - 1)
  s <- gsub('\\\\"', '"', s); s <- gsub('\\\\\\\\', '\\\\', s)
  p <- tryCatch(jsonlite::fromJSON(s, simplifyVector = FALSE), error = function(e) NULL)
  if (is.character(p)) p <- tryCatch(jsonlite::fromJSON(p, simplifyVector = FALSE), error = function(e) NULL)
  if (is.list(p) && !is.null(p$messages)) p$messages else NULL
}
chcols13 <- intersect(paste0("chathistory0", 1:5), names(s13))
conv_dur_min <- function(row) {
  msgs <- parse_chat(paste0(unlist(row[chcols13]), collapse = ""))
  if (is.null(msgs)) return(NA_real_)
  ca <- suppressWarnings(as.numeric(vapply(msgs, function(m) if (is.null(m$createdAt)) NA_character_ else as.character(m$createdAt), character(1))))
  if (all(is.na(ca))) return(NA_real_)
  (max(ca, na.rm = TRUE) - min(ca, na.rm = TRUE)) / 60
}
s13d <- s13 |>
  dplyr::mutate(.n_user = rowSums(dplyr::across(dplyr::all_of(ucols), ne)),
                .w_user = rowSums(dplyr::across(dplyr::all_of(ucols), wc)),
                .w_asst = rowSums(dplyr::across(dplyr::all_of(acols), wc)))
s13d$.cdur <- vapply(seq_len(nrow(s13d)), function(i) conv_dur_min(s13d[i, ]), numeric(1))
me_convo_s13 <- s13d |>
  dplyr::group_by(study = as.character(study_factor)) |>
  dplyr::summarise(mean_user_msgs = mean(.n_user), mean_user_words = mean(.w_user),
                   mean_asst_words = mean(.w_asst), median_dur_min = median(.cdur, na.rm = TRUE),
                   .groups = "drop")
me_convo_s13
```

```{r me-convo-s4}
# robust reconstruction of an S4 chunked-JSON transcript (mirrors .ext_parse_s4_transcript):
# chathistory0* are successive fragments of one escaped JSON string literal.
parse_s4_transcript <- function(chunks) {
  chunks <- chunks[!is.na(chunks) & nzchar(chunks)]
  if (!length(chunks)) return(NULL)
  inner <- vapply(chunks, function(s) {
    if (startsWith(s, '"')) s <- substr(s, 2, nchar(s))
    if (endsWith(s, '"'))   s <- substr(s, 1, nchar(s) - 1)
    s }, character(1), USE.NAMES = FALSE)
  j <- paste0(inner, collapse = "")
  j <- gsub('\\\\"', '"', j); j <- gsub('\\\\\\\\', '\\\\', j)
  p <- tryCatch(jsonlite::fromJSON(j, simplifyVector = FALSE), error = function(e) NULL)
  if (is.character(p)) p <- tryCatch(jsonlite::fromJSON(p, simplifyVector = FALSE), error = function(e) NULL)
  if (is.list(p) && !is.null(p$messages)) p$messages else NULL
}
chat_cols <- intersect(paste0("chathistory0", 1:5), names(s4))
s4_msgs <- lapply(seq_len(nrow(s4)), function(i) parse_s4_transcript(as.character(unlist(s4[i, chat_cols]))))
s4_metrics <- purrr::map_dfr(s4_msgs, function(m) {
  if (is.null(m)) return(tibble::tibble(n_user = NA, w_asst = NA, dur = NA_real_))
  roles <- vapply(m, function(x) if (!is.null(x$role)) x$role else NA_character_, character(1))
  cont  <- vapply(m, function(x) if (!is.null(x$content)) as.character(x$content) else "", character(1))
  ca    <- suppressWarnings(as.numeric(vapply(m, function(x) if (is.null(x$createdAt)) NA_character_ else as.character(x$createdAt), character(1))))
  tibble::tibble(n_user = sum(roles == "user", na.rm = TRUE),
                 w_asst = sum(wc(cont[roles == "assistant"])),
                 dur = if (all(is.na(ca))) NA_real_ else (max(ca, na.rm = TRUE) - min(ca, na.rm = TRUE)) / 60)
})
me_convo_s4 <- tibble::tibble(
  study = "Study 4 (frontier models)",
  mean_user_msgs = mean(s4_metrics$n_user, na.rm = TRUE),
  mean_asst_words = mean(s4_metrics$w_asst, na.rm = TRUE),
  median_dur_min = median(s4_metrics$dur, na.rm = TRUE))
me_convo_s4
```

```{r me-substantive-claims}
# Load the assembled all_numbers table using the same robust pattern as the SI docs
# (supplement/Bunkbot_SI.Rmd): reuse the cached engine build if present and readable,
# otherwise build it live (API-free) and cache it. This makes the render work on a
# fresh checkout, where output/ is empty (the .rds is a gitignored build artifact).
source(file.path(REPO_ROOT, "code", "R", "build_all_numbers.R"))
nums_an <- load_or_build_all_numbers(REPO_ROOT)$all_numbers
me_subst <- nums_an |>
  dplyr::filter(block == "claim_counts_full", term == "mean_factchecked_per_conv",
                sample == "cached_api_claims",
                model %in% c("Study1", "Study2", "Study3", "Study4"), is.na(direction)) |>
  dplyr::transmute(study = dplyr::recode(model, Study1 = "Jailbroken", Study2 = "Standard",
                                         Study3 = "Truth-Constrained", Study4 = "Study 4 (frontier models)"),
                   substantive_claims_per_conv = estimate)
me_subst
```

```{r me-convo-vals, include=FALSE}
gc13 <- function(s, col) me_convo_s13[[col]][me_convo_s13$study == s]
gsub_ <- function(s) me_subst$substantive_claims_per_conv[me_subst$study == s]
```

Conversations were short multi-turn exchanges. Participants sent on average
`r mn(gc13("Jailbroken","mean_user_msgs"))`–`r mn(gc13("Standard","mean_user_msgs"))`
messages in Studies 1–3 and `r mn(me_convo_s4$mean_user_msgs)` in Study 4, writing
roughly `r fmt_num(gc13("Jailbroken","mean_user_words"),0)`–`r fmt_num(gc13("Truth-Constrained","mean_user_words"),0)`
words. The assistant produced far more text, and markedly more under the
later-study models: a mean of `r fmt_num(gc13("Jailbroken","mean_asst_words"),0)`
assistant words per conversation in Study 1, about
`r fmt_num(gc13("Standard","mean_asst_words"),0)` and
`r fmt_num(gc13("Truth-Constrained","mean_asst_words"),0)` in Studies 2 and 3, and
`r fmt_num(me_convo_s4$mean_asst_words,0)` in Study 4. Each conversation contained
`r mn(gsub_("Jailbroken"))`/`r mn(gsub_("Standard"))`/`r mn(gsub_("Truth-Constrained"))`/`r mn(gsub_("Study 4 (frontier models)"))`
fact-checked substantive claims (Studies 1–4). Median conversation duration was
`r mn(gc13("Jailbroken","median_dur_min"))`/`r mn(gc13("Standard","median_dur_min"))`/`r mn(gc13("Truth-Constrained","median_dur_min"))`
minutes in Studies 1–3 and `r mn(me_convo_s4$median_dur_min)` minutes in Study 4.

## Claim extraction and fact-checking volume {-}

The veracity pipeline extracted discrete factual claims from every assistant
turn and fact-checked the substantive ones. The reported totals are restricted to
the analytic samples. For Studies 1–3 the per-conversation extraction files
(`*_nfacts_veracity.jsonl`) are read and a claim counts as fact-checked when its
`veracity_request_status == "success"`; we join to the analytic frame by
`ResponseId`. For Study 4 the per-conversation claim census in the master dataset
(`all_claims_total`, `factcheck_scored_n`) is summed over the strict sample
(mirrors `compute_claim_counts_full`).

```{r me-claims-s13}
me_claims_s13 <- purrr::map_dfr(1:3, function(st) {
  ls   <- readLines(paths$nfacts[[paste0("Study", st)]], warn = FALSE)
  recs <- lapply(ls, jsonlite::fromJSON)
  rid  <- vapply(recs, function(x) if (is.null(x$ResponseId)) NA_character_ else as.character(x$ResponseId), character(1))
  vok  <- vapply(recs, function(x) !is.null(x$veracity_request_status) &&
                   identical(as.character(x$veracity_request_status), "success"), logical(1))
  tibble::tibble(study = c("Jailbroken","Standard","Truth-Constrained")[st],
                 fact_checked_analytic = sum(vok & rid %in% s13$response_id))
})
me_claims_s13
s13_factchecked_total <- sum(me_claims_s13$fact_checked_analytic)
s13_factchecked_total
```

```{r me-claims-s4}
ids4 <- as.character(s4$ResponseId)
idcol <- if ("conversation_id" %in% names(d4$s4_master)) "conversation_id" else "ResponseId"
m_anal <- d4$s4_master[as.character(d4$s4_master[[idcol]]) %in% ids4, ]
me_claims_s4 <- tibble::tibble(
  s4_conversations = nrow(m_anal),
  extracted  = sum(m_anal$all_claims_total,   na.rm = TRUE),
  factchecked = sum(m_anal$factcheck_scored_n, na.rm = TRUE))
me_claims_s4
```

Across Studies 1–3, `r fmt_num(s13_factchecked_total,0)` extracted claims in the
analytic-sample conversations received a fact-check score. In Study 4 the four
frontier models produced `r fmt_num(me_claims_s4$extracted,0)` extracted claims
across the `r fmt_num(me_claims_s4$s4_conversations,0)` analytic conversations, of
which `r fmt_num(me_claims_s4$factchecked,0)` were fact-checked.

## Five-rater post-stance reliability {-}

Every Study-4 pre- and post-conversation social-media post is scored by a
five-model consensus rater drawn from five developers
(`claude-sonnet-4.6`, `gpt-5.2`, `gemini-3.1-pro`, `grok-4.3`, `deepseek-v3.2`).
Reliability is computed live from the per-rater 0–100 `stance_score` columns in
the consolidated stance file: the interval Krippendorff's α, the two-way ICC(2,k),
and the mean pairwise inter-rater correlation. Off-topic and declined post rates
are the post-conversation consensus `focal_relevance == "other_topic"` and
`response_type == "declines_to_post"` shares (cross-checked against the
`stance_reliability` block).

```{r me-stance-reliability}
sv <- readr::read_csv(paths$s4_stance_v2, show_col_types = FALSE, progress = FALSE)
raters <- c("claude_stance_score","gpt_stance_score","gemini_stance_score","grok_stance_score","deepseek_stance_score")
M <- as.matrix(sv[, raters]); storage.mode(M) <- "numeric"
alpha_interval <- irr::kripp.alpha(t(M), method = "interval")$value
icc2k <- { ic <- psych::ICC(M, missing = FALSE, lmer = FALSE); ic$results$ICC[ic$results$type == "ICC2k"] }
cm <- cor(M, use = "pairwise.complete.obs"); pairwise_r <- mean(cm[upper.tri(cm)])
me_rel <- tibble::tibble(
  n_rated_items = nrow(sv),
  krippendorff_alpha_interval = alpha_interval,
  ICC_2k = icc2k,
  mean_pairwise_r = pairwise_r)
me_rel
```

```{r me-stance-composition}
post <- sv |> dplyr::filter(timepoint == "post")
me_comp <- tibble::tibble(
  n_post = nrow(post),
  pct_off_topic = mean(post$consensus_focal_relevance == "other_topic", na.rm = TRUE),
  pct_declining = mean(post$consensus_response_type == "declines_to_post", na.rm = TRUE))
me_comp
```

Inter-rater agreement was high: across the `r fmt_num(me_rel$n_rated_items,0)`
rated post items the interval Krippendorff's α was
`r mn2(me_rel$krippendorff_alpha_interval)`, the two-way ICC(2,k) was
`r mn2(me_rel$ICC_2k)`, and the mean pairwise correlation among raters was
`r mn2(me_rel$mean_pairwise_r)`. Among post-conversation posts, only
`r pct(me_comp$pct_off_topic)` were scored as off-topic and
`r pct(me_comp$pct_declining)` as declining to post; these were flagged for the
composition tables and assigned the neutral midpoint in the direction-coded
continuous analyses. Annotation inter-rater agreement for the APE compliance
labels — reported as 84% agreement and a Cohen's κ of 0.66 — is carried over from
the APE instrument's original validation (Kowal et al., 2025, arXiv:2506.02873)
and is not recomputed here.

## Topic taxonomy {-}

To ask whether persuasion effects concentrate in recognizable conspiracy
families, the pooled focal descriptions were clustered into topics with a fixed,
fully specified embedding → PCA → HDBSCAN procedure. The number of interpretable
topics and the size of the residual "Mixed / Unclassified" bucket are read from
the canonical cached assignment table (`topic_assignments.csv`; cross-checked
against the `topic_effects` block).

```{r me-topics}
ta <- readr::read_csv(paths$topic_assignments, show_col_types = FALSE, progress = FALSE)
n_topics_named <- ta |>
  dplyr::filter(topic != "Mixed / Unclassified") |>
  dplyr::distinct(topic) |> nrow()
pct_mixed <- mean(ta$topic == "Mixed / Unclassified")
me_topics <- tibble::tibble(n_corpus = nrow(ta), n_named_topics = n_topics_named, pct_mixed = pct_mixed)
me_topics
```

The selected solution yields `r fmt_num(me_topics$n_named_topics,0)` named,
interpretable density clusters (e.g. "Moon Landing Hoax", "9/11 Inside Job",
"Vaccines Cause Autism") plus a residual Mixed / Unclassified bucket that absorbs
HDBSCAN noise points and any cluster below the size floor; the Mixed bucket
accounts for `r pct(me_topics$pct_mixed)` of the
`r fmt_num(me_topics$n_corpus,0)`-description pooled corpus.
