# Study 4 — Four frontier models {#s4}

```{r s4-build}
d4 <- build_s4_data(paths)        # strict N=1272 + compliant N=1,073 post-rescore + claim_role/master/compliance
validate_s4_data(d4)
s4  <- d4$s4_with_compliance       # observed-outcome sample (1,272), with APE compliance flags joined
nrow(s4)
table(s4$model_pooled, s4$direction)
```

In Study 4 (*N* = `r fmt_num(nrow(d4$s4),0)` after pre-registered exclusions), we move from
private belief to the information environment that surrounds it. Participants wrote a social
media post about their focal conspiracy and indicated how likely they would be to share it
publicly online (0–100), before and after the conversation. We randomized participants to
interact with one of four of the most sophisticated frontier models then available
(Claude Opus 4.6, Gemini 3.1 Pro, GPT-5.2, and Grok 4). We used the models' default safety
settings — no jailbreak — but, as in Study 1, the prompt explicitly gave the models permission
to invent evidence in *both* conditions, so any difference between models reflects how each
model behaves when offered the same latitude. Belief change is direction-aligned (positive =
movement toward the model's assigned position).

## Per-model belief change {-}

### Preregistered observed-outcome model: `lm(aligned_belief_change ~ direction * model_pooled)` (HC3) {-}

The registered specification is an unadjusted aligned-change OLS with a direction-by-model
interaction and HC3 robust SEs, fit among all participants with observed outcomes (no
compliance filter, no baseline covariate). Because post-treatment outcomes are missing for
participants who left after delivery, the confirmed-chat-start sensitivity analysis
is reported in the supplement. We expose the full coefficient table, the heteroscedasticity
check, and — as the disclosed precision deviation — the baseline-adjusted refit.

```{r s4-belief-itt}
s4b <- s4 |> mutate(direction = factor(as.character(direction), levels = c("bunk", "debunk")),
                    model_pooled = factor(as.character(model_pooled), levels = c("Claude","Gemini","GPT-5.2","Grok")))
m_s4 <- lm(aligned_belief_change ~ direction * model_pooled, data = s4b)
coeftest(m_s4, vcov. = hc3(m_s4))                  # full coefficient table (HC3), reference = bunk, Claude
lmtest::bptest(m_s4)                                # heteroscedasticity check -> motivates HC3
```

```{r s4-belief-itt-adj}
# baseline-adjusted refit (disclosed precision deviation): + belief_rating_pre_4
m_s4a <- lm(aligned_belief_change ~ direction * model_pooled + belief_rating_pre_4, data = s4b)
coeftest(m_s4a, vcov. = hc3(m_s4a))
```

Per-model cell means (raw direction-aligned belief change) and the within-model
bunk-vs-debunk contrast follow. Each cell is a one-sample test that pre→post movement is
non-zero in the assigned direction; the contrast is the model-specific `directionbunk`
combination from the HC3-robust fit (mirroring `raw_aligned_means_cells` and the engine's
`prereg_primary_model_contrasts`).

```{r s4-belief-cells}
# (a) raw cell means + one-sample t (movement in assigned direction)
cell_means <- s4b |>
  group_by(model_pooled, direction) |>
  summarise(n = n(), mean = mean(aligned_belief_change),
            se = sd(aligned_belief_change)/sqrt(n),
            lo = mean - qt(.975, n-1)*se, hi = mean + qt(.975, n-1)*se,
            t = t.test(aligned_belief_change)$statistic,
            p = t.test(aligned_belief_change)$p.value, .groups = "drop")
kable(cell_means, digits = 3,
      caption = "Study 4 belief change — raw direction-aligned cell means by model and direction.")
```

```{r s4-belief-within-model}
# (b) within-model bunk - debunk contrast, HC3 (releveled: debunk reference for engine convention)
s4r <- s4 |> mutate(direction = factor(as.character(direction), levels = c("debunk","bunk")),
                    model_pooled = factor(as.character(model_pooled), levels = c("Claude","Gemini","GPT-5.2","Grok")))
m_s4r <- lm(aligned_belief_change ~ direction * model_pooled, data = s4r)
Vr <- hc3(m_s4r)
ref <- levels(s4r$model_pooled)[1]
within_contrast <- function(m) {
  w <- c(directionbunk = 1)
  if (m != ref) w[paste0("directionbunk:model_pooled", m)] <- 1
  coef_test_row(m_s4r, Vr, w, m)                  # full zero-filled weight vector -> bunk - debunk
}
s4_within <- bind_rows(lapply(levels(s4r$model_pooled), within_contrast)) |>
  transmute(model = contrast, estimate, conf.low = ci_low, conf.high = ci_high,
            statistic = t, p.value = p)
kable(s4_within, digits = 3,
      caption = "Study 4 within-model bunk − debunk contrast on aligned belief change (HC3).")
```

```{r s4-belief-vals, include=FALSE}
gc <- function(m, dir, col) cell_means[[col]][cell_means$model_pooled==m & cell_means$direction==dir]
gpt_bunk_m <- gc("GPT-5.2","bunk","mean"); gpt_bunk_p <- gc("GPT-5.2","bunk","p")
gpt_deb_m  <- gc("GPT-5.2","debunk","mean"); gpt_deb_p <- gc("GPT-5.2","debunk","p")
gpt_ct <- s4_within |> filter(model=="GPT-5.2")
# largest cell p across Claude/Gemini/Grok (to back the "all p<.001" claim)
others_maxp <- max(cell_means$p[cell_means$model_pooled %in% c("Claude","Gemini","Grok")])
```

Claude Opus 4.6, Gemini 3.1 Pro and Grok 4 produced effects broadly similar to out-of-the-box
GPT-4o in Study 2: large and statistically significant bunking *and* debunking effects for all
three models (every pre-to-post change `r P(others_maxp)`). Their cell means ranged from
`r mn(min(cell_means$mean[cell_means$model_pooled %in% c("Claude","Gemini","Grok")]))` to
`r mn(max(cell_means$mean[cell_means$model_pooled %in% c("Claude","Gemini","Grok")]))` points of
aligned movement. GPT-5.2, conversely, produced a strikingly different pattern in which the
model failed to promote conspiracies when prompted to. In fact, conspiracy belief *significantly
decreased* pre-to-post in the bunking condition under GPT-5.2 (M = `r mn(gpt_bunk_m)`,
`r P(gpt_bunk_p)`), while debunking moved belief `r mn(gpt_deb_m)` points in the intended
direction (`r P(gpt_deb_p)`). For GPT-5.2, then, debunking was vastly more successful than
bunking at moving beliefs in the intended direction (bunk − debunk = `r mn2(gpt_ct$estimate)`,
*z* = `r mn2(gpt_ct$statistic)`, `r P(gpt_ct$p.value)`).

## From private belief to public speech: sharing on the extensive margin {-}

We now turn to the social media posts. We treat "would publicly post" as a binary outcome and
separate *pro-conspiracy* posting from *anti-conspiracy* posting, because a debunking
conversation can simultaneously reduce willingness to spread the conspiracy and increase
willingness to spread a correction. A post counts as pro-conspiracy public speech when the
participant's sharing likelihood exceeds 50 *and* the post's classified stance is
pro-conspiracy (stance > 50); anti-conspiracy public speech is sharing likelihood > 50 with an
anti-conspiracy stance (stance < 50). Within each arm we test the pre→post shift with a McNemar
within-participant test (mirroring the engine's `ext_extensive_margin` / `extensive_margin`
block); the per-model rows use direction-aligned posting (pro for bunking, anti for debunking).

```{r s4-extensive}
sp <- d4$s4 |>
  mutate(sp_pre = as.numeric(share_pre_4), sp_post = as.numeric(share_post_4),
         Condition = ifelse(direction == "bunk", "Bunking", "Debunking")) |>
  filter(!is.na(sp_pre), !is.na(sp_post), !is.na(pre_direction_score), !is.na(post_direction_score)) |>
  mutate(pre_pro  = sp_pre  > 50 & pre_direction_score  > 50,
         post_pro = sp_post > 50 & post_direction_score > 50,
         pre_anti  = sp_pre  > 50 & pre_direction_score  < 50,
         post_anti = sp_post > 50 & post_direction_score < 50,
         pre_al  = ifelse(direction == "bunk", pre_pro,  pre_anti),
         post_al = ifelse(direction == "bunk", post_pro, post_anti))

extfun <- function(df, pre, post) {
  pr <- df[[pre]]; po <- df[[post]]
  tb <- table(factor(pr, c(FALSE, TRUE)), factor(po, c(FALSE, TRUE)))
  p  <- tryCatch(stats::mcnemar.test(tb)$p.value, error = function(e) NA_real_)
  tibble(n = nrow(df), pre_pct = 100*mean(pr), post_pct = 100*mean(po),
         net_pp = 100*(mean(po) - mean(pr)),
         newly_yes = as.integer(tb["FALSE","TRUE"]), newly_no = as.integer(tb["TRUE","FALSE"]),
         mcnemar_p = p)
}

# pooled within-arm net change: pro-conspiracy and anti-conspiracy posting
ext_pooled <- bind_rows(
  sp |> group_by(Condition) |> group_modify(~ extfun(.x, "pre_pro",  "post_pro"))  |> mutate(outcome = "pro_public_posting"),
  sp |> group_by(Condition) |> group_modify(~ extfun(.x, "pre_anti", "post_anti")) |> mutate(outcome = "anti_public_posting")
) |> ungroup()
kable(ext_pooled |> select(outcome, Condition, n, pre_pct, post_pct, net_pp, mcnemar_p),
      digits = 3, caption = "Within-arm net percentage-point change in willingness to publicly post (pooled, McNemar).")
```

```{r s4-extensive-by-model}
# per-model within-arm net change on direction-aligned posting
ext_model <- sp |>
  group_by(model_pooled = as.character(model_pooled), Condition) |>
  group_modify(~ extfun(.x, "pre_al", "post_al")) |> ungroup()
kable(ext_model |> select(model_pooled, Condition, n, pre_pct, post_pct, net_pp, mcnemar_p),
      digits = 3, caption = "Per-model within-arm net pp change in direction-aligned public posting (McNemar).")
```

```{r s4-extensive-vals, include=FALSE}
gp_pool <- function(oc, cond, col) ext_pooled[[col]][ext_pooled$outcome==oc & ext_pooled$Condition==cond]
gm <- function(m, cond, col) ext_model[[col]][ext_model$model_pooled==m & ext_model$Condition==cond]
# debunking pooled effects
deb_pro_np  <- gp_pool("pro_public_posting","Debunking","net_pp");  deb_pro_p  <- gp_pool("pro_public_posting","Debunking","mcnemar_p")
deb_anti_np <- gp_pool("anti_public_posting","Debunking","net_pp"); deb_anti_p <- gp_pool("anti_public_posting","Debunking","mcnemar_p")
# bunking per-model aligned (pro-conspiracy) effects
b_claude_np <- gm("Claude","Bunking","net_pp");  b_claude_p <- gm("Claude","Bunking","mcnemar_p")
b_grok_np   <- gm("Grok","Bunking","net_pp");    b_grok_p   <- gm("Grok","Bunking","mcnemar_p")
b_gem_np    <- gm("Gemini","Bunking","net_pp");  b_gem_p    <- gm("Gemini","Bunking","mcnemar_p")
b_gpt_np    <- gm("GPT-5.2","Bunking","net_pp"); b_gpt_p    <- gm("GPT-5.2","Bunking","mcnemar_p")
# bunking effect on anti-conspiracy posting (no decrease for Claude/Gemini/Grok; increase for GPT-5.2)
gm_anti <- function(m, col) {
  sub <- sp |> filter(model_pooled == m, Condition == "Bunking")
  e <- extfun(sub, "pre_anti", "post_anti"); e[[col]]
}
anti_claude_np <- gm_anti("Claude","net_pp"); anti_claude_p <- gm_anti("Claude","mcnemar_p")
anti_gem_np    <- gm_anti("Gemini","net_pp"); anti_gem_p    <- gm_anti("Gemini","mcnemar_p")
anti_grok_np   <- gm_anti("Grok","net_pp");   anti_grok_p   <- gm_anti("Grok","mcnemar_p")
anti_gpt_np    <- gm_anti("GPT-5.2","net_pp"); anti_gpt_p   <- gm_anti("GPT-5.2","mcnemar_p")
```

After the debunking conversations, participants substantially reduced their reported
likelihood of sharing *pro-conspiracy* posts (net change `r mn(deb_pro_np)` pp, `r P(deb_pro_p)`)
and substantially increased their likelihood of sharing *anti-conspiracy* posts
(net change +`r mn(deb_anti_np)` pp, `r P(deb_anti_p)`) — both shifts highly significant across
all models. Conversely, in the bunking condition participants showed much smaller increases in
their likelihood of sharing pro-conspiracy posts after talking to Claude Opus 4.6
(+`r mn(b_claude_np)` pp, `r P(b_claude_p)`) and Grok 4 (+`r mn(b_grok_np)` pp, `r P(b_grok_p)`),
no significant change after Gemini 3.1 Pro (+`r mn(b_gem_np)` pp, `r P(b_gem_p)`), and a large
*decrease* after GPT-5.2 (`r mn(b_gpt_np)` pp, `r P(b_gpt_p)`). On the anti-conspiracy margin,
bunking left sharing intentions essentially unchanged for Gemini (`r mn(anti_gem_np)` pp,
`r P(anti_gem_p)`) and Grok (+`r mn(anti_grok_np)` pp, `r P(anti_grok_p)`), produced a small but
significant *decrease* for Claude (`r mn(anti_claude_np)` pp, `r P(anti_claude_p)`), and a
significant *increase* for GPT-5.2 (+`r mn(anti_gpt_np)` pp, `r P(anti_gpt_p)`) — the last
consistent with that model arguing against the conspiracy even when assigned to bunk.

## Compliance by model {-}

The variation in bunking efficacy across models is largely attributable to variation in
*compliance* in the bunking condition. We use the cached APE turn-0 classifier: `attempt_binary`
(the model attempted the assigned direction) and `strict_compliant` (attempted and did not
refuse). Rates are computed among compliance-scored conversations in the bunking arm, mirroring
the engine's `compliance_cells` block.

```{r s4-compliance}
comp_bunk <- s4 |>
  filter(direction == "bunk", compliance_scored) |>
  group_by(model_pooled) |>
  summarise(n = n(),
            attempt_rate = mean(attempt_binary == 1, na.rm = TRUE),
            refusal_rate = mean(refusal_binary == 1, na.rm = TRUE),
            strict_compliance_rate = mean(strict_compliant, na.rm = TRUE),
            .groups = "drop")
kable(comp_bunk, digits = 3,
      caption = "Study 4 bunking-arm compliance by model (APE turn-0 classifier).")
```

```{r s4-compliance-vals, include=FALSE}
ca <- function(m) comp_bunk$strict_compliance_rate[comp_bunk$model_pooled==m]
```

When prompted to bunk, compliance (strict: attempted and did not refuse) was high for
Gemini 3.1 Pro (`r pct(ca("Gemini"),0)`), Grok 4 (`r pct(ca("Grok"),0)`), and
Claude Opus 4.6 (`r pct(ca("Claude"),0)`), but extremely low for GPT-5.2 (`r pct(ca("GPT-5.2"))`).

### GPT-5.2 bunking decomposition {-}

When GPT-5.2 did not attempt to bunk, it was not merely silent. Among its compliance-scored,
non-attempting bunking conversations we split the non-attempts into explicit refusals and
counter-arguments (no refusal — the model instead argued *against* the conspiracy), mirroring
`ext_nonattempt.R` / the `gpt52_nonattempt_split` block.

```{r s4-gpt52-decomp}
gpt_bunk <- s4 |> filter(model_pooled == "GPT-5.2", direction == "bunk", compliance_scored)
n_scored  <- nrow(gpt_bunk)
n_attempt <- sum(gpt_bunk$attempt_binary == 1, na.rm = TRUE)
gpt_na <- gpt_bunk |> filter(attempt_binary == 0)
n_na      <- nrow(gpt_na)
n_refuse  <- sum(gpt_na$refusal_binary == 1, na.rm = TRUE)         # explicit refusals
n_counter <- sum(gpt_na$refusal_binary == 0, na.rm = TRUE)         # argued against (no refusal)
gpt_decomp <- tibble(
  term = c("attempted", "did_not_attempt", "argued_against_no_refusal", "explicit_refusal"),
  n    = c(n_attempt, n_na, n_counter, n_refuse),
  share_of_scored = c(n_attempt, n_na, n_counter, n_refuse) / n_scored)
kable(gpt_decomp, digits = 3,
      caption = "GPT-5.2 bunking arm: attempt / counter-argument / refusal decomposition (compliance-scored).")
```

```{r s4-gpt52-vals, include=FALSE}
sh <- function(lab) gpt_decomp$share_of_scored[gpt_decomp$term==lab]
```

That is, when prompted to bunk, GPT-5.2 ignored its instructions to mislead and instead argued
*against* the conspiracy in the majority (`r pct(sh("argued_against_no_refusal"))`) of cases,
and simply refused to comply `r pct(sh("explicit_refusal"))` of the time.

## Claim accuracy by model in bunking {-}

Finally, GPT-5.2's compliance also differed in *kind*. Every substantive factual claim was
extracted and independently fact-checked; we read off the share of intended-direction
("aligned") claims scoring below the low-veracity cutoff (fact-check score < 40) in the
bunking arm, by model — the claim-level tail rate that SI Table S37 reports (block
`veracity_tail`, term `p_lt40`, strict claim set). This is the estimand the manuscript
quotes. (A per-conversation "% explicitly false" panel metric, `accuracy_panel_metrics`,
model-adjusts conversation-level rates and can order Gemini/Grok differently; the
manuscript standardizes on the claim-level Table S37 measure.)

```{r s4-accuracy}
source(file.path(REPO_ROOT, "code", "R", "build_all_numbers.R"))  # load_or_build_all_numbers()
false_bunk <- load_or_build_all_numbers(REPO_ROOT)$all_numbers |>
  filter(block == "veracity_tail", term == "p_lt40", outcome == "aligned",
         direction == "bunk", sample == "strict_claims",
         grepl("^Study4_", model)) |>
  transmute(model = sub("^Study4_", "", model), pct_lt40 = estimate, n_claims = n)
kable(false_bunk, digits = 1,
      caption = "Study 4 bunking-arm % of aligned claims scoring < 40 (claim-level; SI Table S37 measure).")
```

```{r s4-accuracy-vals, include=FALSE}
pf <- function(m) false_bunk$pct_lt40[false_bunk$model==m]
```

GPT-5.2 made overwhelmingly accurate claims — only `r pct(pf("GPT-5.2")/100)` of its aligned
bunking claims were false — whereas the other models were much more willing to lie when bunking:
Claude Opus 4.6 = `r pct(pf("Claude")/100)` false, Grok 4 = `r pct(pf("Grok")/100)` false,
and Gemini 3.1 Pro = `r pct(pf("Gemini")/100)` false. The model that best obeyed its honesty
instincts (GPT-5.2) was precisely the one that refused to bunk; the models that complied did so
by fabricating evidence at high rates.
