# Study 3 — Truth-Constrained GPT-4o {#s3}

```{r s3-data}
s3 <- s13 |> filter(study_factor == "Truth-Constrained") |> droplevels() |> recode_perceptions() |> add_gcbs()
nrow(s3)
```

In Study 3 (*N* = `r fmt_num(nrow(s3),0)`) we test the effect of a very basic safeguard —
simply instructing the model to argue only with accurate, truthful evidence. This
"truth-constrained" GPT-4o was identical to the Study 2 standard model except for a system
prompt forbidding fabricated or misleading claims. The constraint lets us ask two questions
at once: does it actually make the model's arguments more accurate, and does honest persuasion
remain effective?

## Claim veracity across Studies 1–3 {-}

Every substantive factual claim the model made was extracted and independently fact-checked
(0–100 veracity). We adopt the *direction-aligned* measure: the mean veracity of the
direct-and-indirect claims that point in the model's assigned direction (supporting the
conspiracy when bunking, opposing it when debunking), averaged within conversation and then
across conversations. This mirrors `compute_veracity_numbers()` exactly — cached per-claim
LLM labels aggregated in R, no API calls.

```{r s3-veracity}
labels <- read_claim_labels(paths)
# Restrict to the analytic sample so excluded-participant claims cannot bleed in
# (the cached labels are already analytic-only; this enforces it explicitly).
labels <- labels |> filter(conversation_id %in% s13$response_id)
conv   <- conv_aligned_veracity(labels) |>
  filter(study %in% c("Study1","Study2","Study3"), !is.na(aligned_veracity))

# study x direction mean aligned veracity (conversation-level), with CI
vtab <- conv |>
  group_by(study, direction) |>
  summarise(n_conv = n(), mean_veracity = mean(aligned_veracity),
            se = sd(aligned_veracity)/sqrt(n_conv),
            lo = mean_veracity - qt(.975, n_conv-1)*se,
            hi = mean_veracity + qt(.975, n_conv-1)*se, .groups = "drop")
kable(vtab, digits = 2, caption = "Studies 1–3 aligned claim veracity by direction (conversation-level).")
```

```{r s3-veracity-standardized}
# two-sample d (debunk vs bunk) + Welch t, per study
vstat <- function(st) {
  cv <- conv |> filter(study == st)
  b <- cv$aligned_veracity[cv$direction == "bunk"]
  d <- cv$aligned_veracity[cv$direction == "debunk"]
  sp <- sqrt(((length(b)-1)*var(b) + (length(d)-1)*var(d)) / (length(b)+length(d)-2))
  tt <- t.test(d, b)                                   # debunk − bunk
  tibble(study = st, mBunk = mean(b), mDebunk = mean(d),
         d = (mean(d)-mean(b))/sp, t = unname(tt$statistic), p = tt$p.value)
}
vd <- bind_rows(lapply(c("Study1","Study2","Study3"), vstat))
kable(vd, digits = 3, caption = "Aligned veracity, debunking − bunking two-sample d (per study).")
```

```{r s3-veracity-false}
# Percent of aligned, fact-checked claims below 40 ("percent false"), per-claim level
pct_false <- labels |>
  mutate(is_aligned = aligned_flag(stance_to_focal, directness_to_focal, direction)) |>
  filter(is_aligned, !is.na(veracity_score), study_source %in% c("Study1","Study2","Study3")) |>
  group_by(study_source, direction) |>
  summarise(n = n(), pct_false = mean(veracity_score < 40), .groups = "drop")
kable(pct_false, digits = 3, caption = "Share of aligned claims below 40 veracity ('percent false').")
```

```{r s3-veracity-vals, include=FALSE}
gv <- function(st, dir) vtab$mean_veracity[vtab$study==st & vtab$direction==dir]
v3 <- vd |> filter(study == "Study3")
pf <- function(st, dir) pct_false$pct_false[pct_false$study_source==st & pct_false$direction==dir]
```

The constraint worked. In Study 1, the jailbroken model's aligned claims averaged
`r mn(gv("Study1","bunk"))` veracity when bunking and `r mn(gv("Study1","debunk"))` when
debunking; in Study 2 the standard model averaged `r mn(gv("Study2","bunk"))` (bunk) and
`r mn(gv("Study2","debunk"))` (debunk). Under the truth constraint these rose to
`r mn(gv("Study3","bunk"))` (bunk) and `r mn(gv("Study3","debunk"))` (debunk), and the
bunking–debunking veracity gap collapsed (debunk − bunk *d* = `r mn2(v3$d)`,
`r P(v3$p)`, versus *d* = `r mn2(vd$d[vd$study=="Study1"])` and
*d* = `r mn2(vd$d[vd$study=="Study2"])` in the two unconstrained studies). The share of
explicitly false aligned claims (veracity < 40) fell accordingly: in Study 1,
`r pct(pf("Study1","bunk"))` of bunking claims and `r pct(pf("Study1","debunk"))` of
debunking claims were false, dropping to `r pct(pf("Study3","bunk"))` (bunk) and
`r pct(pf("Study3","debunk"))` (debunk) in the truth-constrained study.

## Belief change {-}

### Baseline-adjusted ANCOVA (condition + baseline; HC3) {-}

```{r s3-belief-adj}
m_s3 <- lm(change ~ condition_factor + belief_rating_pre_rc, data = s3)
coeftest(m_s3, vcov. = hc3(m_s3))                  # full coefficient table (HC3)
lmtest::bptest(m_s3)                                # heteroscedasticity check -> motivates HC3
```

```{r s3-belief-adj-cells}
V_s3 <- hc3(m_s3)
s3_adj <- bind_rows(
  Bunking   = as_tibble(lin_cell(m_s3, s3, "Bunking",   V_s3)),
  Debunking = as_tibble(lin_cell(m_s3, s3, "Debunking", V_s3)), .id = "condition") |>
  select(condition, estimate, conf.low, conf.high, p.value)
s3_ct_adj <- lin_contrast(m_s3, s3, "Bunking", "Debunking", V_s3)   # bunk vs debunk (adjusted)
kable(s3_adj, digits = 3, caption = "Study 3 belief change — adjusted marginal means (ANCOVA, HC3).")
s3_ct_adj
```

### Unadjusted specification {-}

```{r s3-belief-unadj}
m_s3u <- lm(change ~ condition_factor, data = s3)
coeftest(m_s3u, vcov. = hc3(m_s3u))                 # unadjusted, HC3
s3_raw <- within_summary(s3)                        # raw within-condition means, dz, Hedges g, % change
kable(s3_raw |> select(condition_factor, n, pre, m, lo, hi, dz, g, pct_change),
      digits = 3, caption = "Study 3 belief change — raw (unadjusted) within-condition summary.")
```

```{r s3-belief-vals, include=FALSE}
bunk3 <- s3_raw |> filter(condition_factor == "Bunking")
deb3  <- s3_raw |> filter(condition_factor == "Debunking")
z_s3  <- s3_ct_adj$estimate / s3_ct_adj$std.error
p_deb3  <- t.test(s3$change[s3$condition_factor=="Debunking"])$p.value
p_bunk3 <- t.test(s3$change[s3$condition_factor=="Bunking"])$p.value
```

Honest debunking remained strong: focal-conspiracy belief fell by `r mn(deb3$m)` points in
the intended direction (95% CI `r ci(deb3$lo, deb3$hi)`, `r P(p_deb3)`, *g* = `r mn2(deb3$g)`).
Truth-constrained bunking, by contrast, was greatly reduced: belief rose only `r mn(bunk3$m)`
points (95% CI `r ci(bunk3$lo, bunk3$hi)`, `r P(p_bunk3)`, *d* = `r mn2(bunk3$dz)`), roughly a
third the size of the standard-model bunking effect. The two magnitudes now differed sharply
(adjusted bunk − debunk contrast = `r mn2(s3_ct_adj$estimate)`, *z* = `r mn2(abs(z_s3))`,
`r P(s3_ct_adj$p.value)`) — the truth constraint hobbled bunking far more than debunking.

## Attempt-to-Persuade (APE) compliance {-}

```{r s3-ape}
s3_ape <- s3 |> group_by(condition_factor) |>
  summarise(n = n(),
            attempt_rate = mean(evaluator_label == 1, na.rm = TRUE),
            paired_rate  = mean(evaluator_label == 1 & reverse_evaluator_label == 1, na.rm = TRUE),
            .groups = "drop")
s3_ape
```

```{r s3-ape-vals, include=FALSE}
ape_b <- s3_ape$attempt_rate[s3_ape$condition_factor == "Bunking"]
ape_d <- s3_ape$attempt_rate[s3_ape$condition_factor == "Debunking"]
```

Part of the bunking shortfall is refusal: under the truth constraint the model attempted to
bunk in only `r pct(ape_b)` of bunking conversations, versus `r pct(ape_d)` of debunking
conversations — when asked to argue *for* a conspiracy using only true evidence, it sometimes
declined.

## Non-compliant bunking {-}

```{r s3-noncompliant}
nc3 <- s3 |> filter(condition_factor == "Bunking", evaluator_label == 0)
nc3_t <- t.test(nc3$change)
tibble(n = nrow(nc3), mean = unname(nc3_t$estimate),
       lo = nc3_t$conf.int[1], hi = nc3_t$conf.int[2],
       t = unname(nc3_t$statistic), p = nc3_t$p.value)
```

```{r s3-noncompliant-vals, include=FALSE}
nc3_m <- unname(nc3_t$estimate)
```

The refusals were not innocuous. In the `r fmt_num(nrow(nc3),0)` bunking conversations the
model did not attempt (first-turn `evaluator_label == 0`), belief actually moved *against* the
assigned bunking goal: `r mn(nc3_m)` points (95% CI
`r ci(nc3_t$conf.int[1], nc3_t$conf.int[2])`, *t* = `r mn2(nc3_t$statistic)`,
`r P(nc3_t$p.value)`) — when the truth-constrained model balked at promoting a conspiracy, the
exchange tended to leave participants *less* conspiratorial, a small backfire on the bunker.

## Matched-compliance counterfactual {-}

The cleanest test isolates the truth constraint from refusal. We restrict *both* arms to the
conversations the model would have engaged under either prompt — the dual-compliance rule
`evaluator_label == 1 & reverse_evaluator_label == 1`, identical to the pooled-symmetry rule —
and contrast direction-aligned belief change. This mirrors the engine's `ext_s3_counterfactual`
(canonical block `s3_counterfactual`).

```{r s3-counterfactual}
A <- s3 |> filter(condition == "treatment_mid_bunk",   evaluator_label == 1, reverse_evaluator_label == 1)
B <- s3 |> filter(condition == "treatment_mid_debunk", evaluator_label == 1, reverse_evaluator_label == 1)
matched <- bind_rows(A, B)
matched$direction <- factor(matched$direction, levels = c("debunk", "bunk"))
m_cf <- lm(change ~ direction + belief_rating_pre_rc, data = matched)
coeftest(m_cf, vcov. = hc3(m_cf))                   # full coefficient table (HC3)
lmtest::bptest(m_cf)                                 # assumption check
```

```{r s3-counterfactual-vals, include=FALSE}
cf <- coeftest(m_cf, vcov. = hc3(m_cf))["directionbunk", ]
cf_est <- cf[["Estimate"]]; cf_se <- cf[["Std. Error"]]
cf_lo  <- cf_est - 1.96 * cf_se; cf_hi <- cf_est + 1.96 * cf_se
cf_p   <- cf[["Pr(>|t|)"]]
```

Even among these matched, fully compliant conversations, truth-constrained bunking moved
belief `r mn2(abs(cf_est))` points *less* than debunking did (bunk − debunk =
`r mn2(cf_est)`, 95% CI `r ci(cf_lo, cf_hi)`, `r P(cf_p)`). The gap is not driven by refusals
or topic mix: holding the model's willingness to engage constant, honesty alone costs bunking
several points of persuasive force.

```{r s3-relefficacy}
# Baseline-adjusted (ANCOVA) compliant bunking by study (mirrors block "belief_change_compliant")
comp_lin_bunk <- function(sf) {
  dat <- s13 |> filter(study_factor == sf, evaluator_label == 1, reverse_evaluator_label == 1) |> droplevels()
  mod <- lm(change ~ condition_factor + belief_rating_pre_rc, data = dat)
  lin_cell(mod, dat, "Bunking", hc3(mod))$estimate
}
rel <- tibble(study = c("Jailbroken","Standard","Truth-Constrained"),
              compliant_bunk = vapply(c("Jailbroken","Standard","Truth-Constrained"), comp_lin_bunk, numeric(1)))
rel <- rel |> mutate(efficacy_vs_S3 = compliant_bunk[study=="Truth-Constrained"] / compliant_bunk)
kable(rel, digits = 3, caption = "Compliant bunking (baseline-adjusted ANCOVA) and S3 relative efficacy.")
```

```{r s3-relefficacy-vals, include=FALSE}
eff_vs_s2 <- rel$efficacy_vs_S3[rel$study == "Standard"]
eff_vs_s1 <- rel$efficacy_vs_S3[rel$study == "Jailbroken"]
```

Cast as relative efficacy among compliant conversations, truth-constrained bunking retained
only `r pct(eff_vs_s2,0)` of the standard-model bunking effect and `r pct(eff_vs_s1,0)` of the
jailbroken bunking effect — the honesty constraint roughly halves how far a compliant bunking
conversation can move belief.

## Truthful "ammunition" and the top veracity quartile {-}

Why does honest bunking work at all? Where genuinely suspicious true evidence exists, the
truth-constrained model can still build a persuasive case. We rank each bunking conversation by
its conversation-average aligned veracity and read off the mean belief change in the top
quartile, by study.

```{r s3-topquartile}
sjoin <- s13 |> filter(condition_factor == "Bunking") |>
  inner_join(conv |> transmute(response_id = conversation_id, aligned_veracity), by = "response_id")
topq <- function(sf) {
  sub <- sjoin |> filter(study_factor == sf, !is.na(aligned_veracity))
  cut <- quantile(sub$aligned_veracity, .75, na.rm = TRUE)
  tq  <- sub |> filter(aligned_veracity >= cut)
  tibble(study = sf, n_top = nrow(tq), cutoff = cut, mean_change = mean(tq$change))
}
tq_tab <- bind_rows(lapply(c("Jailbroken","Standard","Truth-Constrained"), topq))
kable(tq_tab, digits = 2, caption = "Top-veracity-quartile bunking belief change, by study.")
```

```{r s3-topquartile-vals, include=FALSE}
tq_s1 <- tq_tab$mean_change[tq_tab$study=="Jailbroken"]
tq_s2 <- tq_tab$mean_change[tq_tab$study=="Standard"]
tq_s3 <- tq_tab$mean_change[tq_tab$study=="Truth-Constrained"]
```

In the top quartile of aligned veracity, bunking moved belief `r mn(tq_s1)` points in Study 1
and `r mn(tq_s2)` in Study 2; under the truth constraint the top-quartile effect was
`r mn(tq_s3)` points. Honest bunking can still shift belief substantially, but only on the
slice of topics where truthful, on-direction evidence is abundant.

## Paltering and the texture of honest bunking {-}

Reading the truth-constrained transcripts, the model rarely lies; instead it *palters* —
stringing together individually true but selectively chosen facts, leaning on rhetorical
framing, appeals to unanswered questions, and motivated emphasis to nudge belief toward the
conspiracy without stating anything false. The constraint changes the *texture* of persuasion
more than it removes it, which is why a real (if attenuated) bunking effect survives.

## Exploratory moderation {-}

Finally we ask, exploratorily, for whom truth-constrained persuasion is strongest. Following
the repository's moderator-number routine, we regress direction-aligned belief change on each z-scored
pre-treatment moderator (baseline conspiracist ideation via the 7-item GCBS, baseline trust in
AI, and political ideology), within condition and baseline-belief-adjusted, plus the
condition-by-moderator interaction.

```{r s3-moderators}
.z <- function(x) { x <- suppressWarnings(as.numeric(x)); s <- sd(x, na.rm=TRUE)
  if (!is.finite(s) || s == 0) return(rep(NA_real_, length(x))); (x - mean(x, na.rm=TRUE))/s }
.gcbs_pre <- function(df) {
  gp <- paste0("x", c(2,4,6,8,10,12,14), "_gcbs_pre")
  m <- sapply(gp, function(c) suppressWarnings(as.numeric(df[[c]])))
  ifelse(rowSums(is.na(m)) == 0, rowMeans(m), NA_real_) }
.s13_ideology <- function(x) { v <- suppressWarnings(as.numeric(as.character(x))); ifelse(v %in% 1:3, v, NA_real_) }

mods <- list(gcbs = .gcbs_pre(s3),
             ai_trust = suppressWarnings(as.numeric(s3$genai_trust)),
             ideology = .s13_ideology(s3$party_affil))
z_base <- .z(s3$belief_rating_pre_rc)

# per-condition baseline-adjusted standardized slopes
slope_rows <- list(); inter_rows <- list()
for (nm in names(mods)) {
  base_dat <- tibble(change = s3$change, direction = as.character(s3$direction),
                     z_mod = .z(mods[[nm]]), z_base = z_base)
  for (cond in c("bunk","debunk")) {
    d <- base_dat |> filter(direction == cond, !is.na(change), !is.na(z_mod), !is.na(z_base))
    fit <- lm(change ~ z_mod + z_base, data = d)
    r <- hc3_tidy(fit) |> filter(term == "z_mod")
    slope_rows[[length(slope_rows)+1]] <- r |>
      transmute(moderator = nm, condition = cond, n = nrow(d),
                estimate, std.error, statistic, p.value)
  }
  di <- base_dat |> filter(!is.na(change), !is.na(z_mod), !is.na(z_base))
  di$direction <- factor(di$direction, levels = c("bunk","debunk"))
  fit_i <- lm(change ~ direction * z_mod + z_base, data = di)
  ri <- hc3_tidy(fit_i) |> filter(term == "directiondebunk:z_mod")
  inter_rows[[length(inter_rows)+1]] <- ri |>
    transmute(moderator = nm, n = nrow(di), estimate, std.error, statistic, p.value)
}
kable(bind_rows(slope_rows), digits = 3,
      caption = "Study 3 — per-condition standardized moderator slopes (HC3, baseline-adjusted).")
```

```{r s3-moderator-interactions}
kable(bind_rows(inter_rows), digits = 3,
      caption = "Study 3 — condition-by-moderator interactions (debunk − bunk × moderator, HC3).")
```

```{r s3-moderators-vals, include=FALSE}
sl <- bind_rows(slope_rows); it <- bind_rows(inter_rows)
g_gcbs_deb <- sl |> filter(moderator=="gcbs", condition=="debunk")
i_gcbs <- it |> filter(moderator=="gcbs")
i_base <- {
  bd <- tibble(change=s3$change, direction=factor(as.character(s3$direction), levels=c("bunk","debunk")), z_mod=z_base)
  bd <- bd |> filter(!is.na(change), !is.na(z_mod))
  hc3_tidy(lm(change ~ direction * z_mod, data = bd)) |> filter(term=="directiondebunk:z_mod")
}
```

Two patterns stand out. Baseline conspiracist ideation moderated debunking strongly: a 1-SD
higher GCBS predicted `r mn2(g_gcbs_deb$estimate)` points *less* debunking-aligned change
(`r P(g_gcbs_deb$p.value)`), and the condition-by-GCBS interaction was reliable
(`r mn2(i_gcbs$estimate)`, `r P(i_gcbs$p.value)`): the most conspiracy-minded participants were
the hardest to debunk. Baseline belief itself moderated the bunk–debunk gap the same way
(interaction `r mn2(i_base$estimate)`, `r P(i_base$p.value)`). Baseline trust in AI and
political ideology showed no reliable moderation of the truth-constrained effect.
