# Materials

This section gives the verbatim wording of the key measures reported in the
paper. The complete instruments are released as Qualtrics `.qsf` files, and the
verbatim model and annotation prompts (the two persuader prompts and the
fabrication license inside them, the Study-3 truth constraint, the equivocality
classifier, the corrective-debrief prompt, the multi-stage fact-check prompts, the
five-rater stance prompt, and the attention/bot screeners) are reproduced in
Appendix A.

`typewriter` tokens are the variable names under which a measure is released in
the data. Qualtrics pipes such as `${e://Field/conRestatement}` are placeholders
resolved at run time, marking where participant-specific content was injected.
Studies 1--3 (Jailbroken, Standard, and Truth-Constrained GPT-4o) share one
instrument; Study 4 re-uses those measures and adds the pre-treatment
social-media items and the public social-posting battery.

```{r ma-setup, include=FALSE}
# Verbatim key-item wording is parsed from the QSF source files (key_item_wording).
# The parent Rmd setup chunk has already loaded these.
```

## Key reported measures

Table \@ref(tab:ma-key-items) gives the verbatim wording of every measure the
paper reports: the pre- and post-conversation focal belief item, the
Generic Conspiracist Beliefs Scale (GCBS), Trust in AI, the four perception
measures, the uncertain-topic elicitation, and the Study-4 social-posting
battery.

```{r ma-key-items}
key_tags <- c(
  "mid_topic", "mid_evidence",
  "belief_rating_pre", "belief_rating_post", "belief_rating_debrf",
  "GCBS_pre", "GCBS_post",
  "genai_trust", "trust2", "trust_debrf",
  "ArgStrength", "new_info", "Unbiased", "Collaborative",
  "social_post", "final_post",
  "share_pre", "share_post", "share_original_post_now"
)
ki_label <- c(
  mid_topic = "Uncertain-topic elicitation",
  mid_evidence = "Reasons for partial belief",
  belief_rating_pre = "Focal belief (pre-conversation, 0-100)",
  belief_rating_post = "Focal belief (post-conversation, 0-100)",
  belief_rating_debrf = "Focal belief (post-debrief, 0-100; Study 4)",
  GCBS_pre = "GCBS (pre)",
  GCBS_post = "GCBS (post)",
  genai_trust = "Trust in AI (pre)",
  trust2 = "Trust in AI (post)",
  trust_debrf = "Trust in AI (post-debrief; Study 4)",
  ArgStrength = "Perceived argument strength",
  new_info = "Perceived new information",
  Unbiased = "Perceived impartiality",
  Collaborative = "Perceived collaboration",
  social_post = "Public post (pre-conversation; Study 4)",
  final_post = "Public post (post-conversation; Study 4)",
  share_pre = "Willingness to post pre-conversation message (Study 4)",
  share_post = "Willingness to post revised message (Study 4)",
  share_original_post_now = "Willingness to re-post original message (Study 4)"
)
ki <- key_item_wording |>
  dplyr::filter(data_export_tag %in% key_tags) |>
  dplyr::distinct(data_export_tag, .keep_all = TRUE) |>
  dplyr::transmute(
    Measure = ki_label[data_export_tag],
    Variable = data_export_tag,
    Wording = stringr::str_trunc(stringr::str_squish(question_text), 240)
  )
ki <- ki[match(key_tags, ki$Variable), ]
ki <- ki[!is.na(ki$Measure), ]
si_kable_styled(
  ki,
  caption = "Verbatim wording of the key reported measures (one row per measure). Wording is truncated for display; full instruments are in the Qualtrics .qsf files."
)
```

```{r ma-response-scales}
# Verbatim response anchors for the discrete-choice key items, pulled from the QSF
# (one study's version per item; identical across studies). Slider measures (focal
# belief, sharing likelihood) have no discrete choices and are 0-100 continuous.
rs <- survey_materials$choices |>
  dplyr::filter(data_export_tag %in% key_tags, section == "Choices",
                !is.na(choice_text), nzchar(trimws(choice_text))) |>
  dplyr::group_by(data_export_tag) |>
  dplyr::filter(study == dplyr::first(study)) |>
  dplyr::arrange(choice_order, .by_group = TRUE) |>
  dplyr::summarise(`Points` = dplyr::n(),
                   `Response scale (verbatim anchors)` =
                     paste0(choice_order, " = ", stringr::str_squish(choice_text), collapse = "; "),
                   .groups = "drop") |>
  dplyr::mutate(Measure = unname(ki_label[data_export_tag])) |>
  dplyr::filter(!is.na(Measure)) |>
  dplyr::arrange(match(data_export_tag, key_tags)) |>
  dplyr::transmute(Measure, Points, `Response scale (verbatim anchors)`)
si_kable_styled(
  rs,
  caption = "Response scales for the key discrete-choice measures: verbatim anchor labels and number of scale points, extracted from the Qualtrics survey definitions. Slider measures -- the 0-100 focal-belief rating and the 0-100 sharing-likelihood items -- are continuous and labelled only at their endpoints (focal belief: 0 = definitely false, 50 = uncertain, 100 = definitely true), as described below."
)
```

## Pre-conversation measures

After consent and the screening items (Section
\@ref(sample-construction-and-screening)), participants completed a baseline
battery. **Trust in AI** (`genai_trust`; re-administered post-conversation as
`trust2` and, in Study 4, post-debrief as `trust_debrf`) was a single 7-point
Likert item, "I generally trust AI technologies like ChatGPT." **Generic
conspiracist beliefs** (`GCBS_pre`, re-administered as `GCBS_post`) used the
seven-item short form of the Generic Conspiracist Beliefs Scale on a 9-point
scale, averaged into a composite. Two 0--100 sliders measured **familiarity**
(`genai_fam`) and **use frequency** (`genai_use`) of generative AI; both enter the
causal-forest moderation analysis alongside trust in AI, baseline belief, GCBS,
and the demographic covariates.

Participants then selected a single focal conspiracy theory. The analyzed cells
used the "uncertain" elicitation (`mid_topic`), which asked participants to
describe, in one to two sentences, a conspiracy theory proposed before 2024 that
they regarded as genuinely uncertain --- one with "possible evidence for both
sides." Participants then gave their reasons for partial belief (`mid_evidence`).
Both responses were passed to a GPT-4-class **equivocality classifier** that
returned exactly one of `TRUE`, `FALSE`, or `INVALID` (full prompt in Appendix
A); only participants reaching `TRUE` within one correction were eligible. The
survey then displayed an AI-generated one-sentence restatement of the chosen
theory (`${e://Field/conRestatement}`) and collected the baseline belief rating
(`belief_rating_pre`) on a 0--100 slider anchored at `0 = Definitely false`,
`50 = Uncertain`, `100 = Definitely true`. As preregistered, the analytic sample
was restricted to baseline ratings inside the equivocal window
$25 < \text{pre-belief} < 75$.

## Post-conversation measures

Immediately after the AI conversation, participants re-rated belief on the same
0--100 slider. The released variable is `belief_rating_post`. Participants then re-completed the GCBS and Trust-in-AI items
and rated four perceptions of the AI interlocutor (secondary perception measures):
**argument strength** (`arg_strength`, 5-point weak--strong), **new information**
(`new_info`, 10-point), **perceived bias** (`unbiased`, 5-point
biased--unbiased), and **perceived collaboration** (`collaborative`, 5-point
adversarial--collaborative).

In Study 4, after a corrective debrief, participants rated belief a third time
(`belief_rating_debrf`) and re-rated Trust in AI (`trust_debrf`).

## Social-posting battery (Study 4)

Study 4 bracketed the AI conversation with a public-post composition battery. Two
free-text **post composition** tasks --- one before the conversation
(`social_post`) and one after (`final_post`) --- asked the participant to write a
public post as they naturally would online; these are the texts later scored for
stance. Three **share-likelihood** sliders (0--100) captured stated willingness to
post the pre-conversation message (`share_pre`), the revised message
(`share_post`), and --- after the original post was re-shown verbatim --- the
*original* message again (`share_original_post_now`). The preregistered
weighted-sharing outcome combines post stance with share willingness; its
construction is given with the public-speech results.

## Annotation prompts and claim measures

The measurement-layer prompts --- claim extraction, fact-check eligibility triage,
focal-claim veracity, claim role relative to the focal belief, attempt-to-persuade
compliance, and social-post stance --- are reproduced verbatim in Appendix A. The
descriptive prevalence of each claim type is reported with the veracity results
(Section \@ref(claim-veracity-and-claim-counts)).
