## Claim Veracity and Claim Counts

Veracity measures *what the model said*, as distinct from the belief outcomes
that measure *what participants concluded*. Every assistant message is decomposed
into atomic factual claims; each claim is assigned a substantive/non-substantive
role and a directional alignment relative to the assigned persuasive goal; and the
substantive claims are fact-checked on a 0 (false) to 100 (true) scale. Four counts
are distinct: *all* extracted claims (every atomic proposition, including
procedural, rhetorical, and conversational filler), the *fact-checked* subset, the
*substantive* checkable subset, and the *direction-aligned* subset whose stance
supports the position the model was assigned to argue. Claim extraction was run on
the full recorded conversations of **all four studies**, so the raw pre-substantive
extracted total is reported for every study: across the four studies the AI
produced
`r fmt_n(est("claim_counts_full", sample = "cached_api_claims", model = "Study1", direction = NA, term = "n_total_claims") + est("claim_counts_full", sample = "cached_api_claims", model = "Study2", direction = NA, term = "n_total_claims") + est("claim_counts_full", sample = "cached_api_claims", model = "Study3", direction = NA, term = "n_total_claims") + est("claim_counts_full", sample = "s4_master_codebook", model = "Study4", direction = NA, term = "n_total_claims"))`
extracted atomic propositions, of which a far smaller share are *substantive*
position-taking claims (the basis for the reported veracity).

### What kinds of claims were checked

Table \@ref(tab:rb-claim-counts) reports the full claim census --- all extracted,
fact-checked, substantive, and direction-aligned counts for all four studies,
broken down by study, by Study-4 model, and by assigned condition. Table
\@ref(tab:rb-claim-prevalence) reports the corresponding role and category shares.
Across the four studies,
`r fmt_pct(est("claim_prevalence", sample="cached_api_claims", model="(all)", term="share_substantive"), 1)`
of fact-checked claims are substantive and
`r fmt_pct(est("claim_prevalence", sample="cached_api_claims", model="(all)", term="share_aligned"), 1)`
are aligned with the assigned side; among substantive claims,
`r fmt_pct(est("claim_prevalence", sample="cached_api_claims", model="(all)", term="share_substantive_indirect"), 1)`
supply *indirect* supporting evidence rather than directly asserting the focal
proposition. The substantive share is highest in Study 4
(`r fmt_pct(est("claim_prevalence", sample="cached_api_claims", model="Study4", term="share_substantive"), 1)`)
and lowest in the Truth-Constrained study
(`r fmt_pct(est("claim_prevalence", sample="cached_api_claims", model="Study3", term="share_substantive"), 1)`).

```{r rb-claim-counts}
ccf_lab <- all_numbers |>
  dplyr::filter(section == "Veracity", block == "claim_counts_full",
                sample == "cached_api_claims",
                term %in% c("n_factchecked", "n_substantive", "n_aligned")) |>
  dplyr::transmute(model, direction, term, estimate)
ccf_tot <- all_numbers |>
  dplyr::filter(section == "Veracity", block == "claim_counts_full",
                term == "n_total_claims",
                # S1-3 all-extracted from the per-study extraction files; S4 + models from the master census
                (sample == "cached_api_claims" & model %in% c("Study1","Study2","Study3")) |
                (sample == "s4_master_codebook" & model %in% c("Study4","Claude","Gemini","GPT-5.2","Grok"))) |>
  dplyr::transmute(model, direction, term = "n_total", estimate)
# all-studies total = sum of the four study-level extracted counts
ccf_tot <- dplyr::bind_rows(
  ccf_tot,
  ccf_tot |> dplyr::filter(is.na(direction), model %in% c("Study1","Study2","Study3","Study4")) |>
    dplyr::summarise(model = "(all)", direction = NA_character_, term = "n_total",
                     estimate = sum(estimate)))
claim_counts <- dplyr::bind_rows(ccf_tot, ccf_lab) |>
  tidyr::pivot_wider(names_from = term, values_from = estimate) |>
  dplyr::mutate(
    Group = dplyr::recode(model,
              Study1 = "S1 (Jailbroken)", Study2 = "S2 (Standard)",
              Study3 = "S3 (Truth-Constrained)", Study4 = "S4 (all models)",
              `(all)` = "All studies"),
    Condition = dplyr::recode(dplyr::coalesce(direction, "all"),
              bunk = "Bunk", debunk = "Debunk", all = "All"),
    .grp = dplyr::case_when(
      model == "(all)" ~ 0L,
      model %in% c("Study1","Study2","Study3","Study4") & Condition == "All" ~ 1L,
      model %in% c("Claude","Gemini","GPT-5.2","Grok") ~ 2L,
      TRUE ~ 3L),
    .ord = match(model, c("(all)","Study1","Study2","Study3","Study4",
                          "Claude","Gemini","GPT-5.2","Grok")),
    `All extracted` = ifelse(is.na(n_total), "—", fmt_n(n_total)),
    `Fact-checked`  = fmt_n(n_factchecked),
    `Substantive`   = fmt_n(n_substantive),
    `Aligned`       = fmt_n(n_aligned)) |>
  dplyr::arrange(.grp, .ord, Condition) |>
  dplyr::select(Group, Condition, `All extracted`, `Fact-checked`,
                `Substantive`, `Aligned`)
si_kable_styled(claim_counts, caption = "Full claim census, by study, by Study-4 model, and by assigned condition: all extracted atomic claims, the fact-checked subset, the substantive (position-taking, checkable) subset, and the direction-aligned subset used for the reported veracity average. All four counts are computed on the analytic conversation universe for each study. 'All extracted' is the raw pre-substantive census from the claim-extraction stage (Studies 1-3 from the per-study extraction files; Study 4 from the master codebook restricted to the analytic sample); the fact-checked, substantive, and aligned columns are the harmonized fact-check label set on the same conversations. The two stages are parallel measures over a shared set of conversations rather than a strictly nested subset, so 'All extracted' bounds but does not exactly enclose the downstream columns.")
```

```{r rb-claim-prevalence}
cp <- all_numbers |>
  dplyr::filter(block == "claim_prevalence", sample == "cached_api_claims",
                term %in% c("n_factchecked_claims", "share_substantive", "share_aligned",
                            "share_substantive_supports", "share_substantive_direct")) |>
  dplyr::select(model, term, estimate) |>
  tidyr::pivot_wider(names_from = term, values_from = estimate) |>
  dplyr::mutate(
    .ord = match(model, c("Study1", "Study2", "Study3", "Study4", "(all)")),
    Study = dplyr::recode(model,
              Study1 = "S1 (Jailbroken)", Study2 = "S2 (Standard)",
              Study3 = "S3 (Truth-Constrained)", Study4 = "S4 (frontier models)",
              `(all)` = "All studies"),
    `Fact-checked claims` = fmt_n(n_factchecked_claims),
    `% substantive` = fmt_pct(share_substantive, 1),
    `% aligned` = fmt_pct(share_aligned, 1),
    `% supports (of subst.)` = fmt_pct(share_substantive_supports, 1),
    `% direct (of subst.)` = fmt_pct(share_substantive_direct, 1)) |>
  dplyr::arrange(.ord) |>
  dplyr::select(Study, `Fact-checked claims`, `% substantive`, `% aligned`,
                `% supports (of subst.)`, `% direct (of subst.)`)
si_kable_styled(cp, caption = "Claim-role prevalence among fact-checked claims, by study and overall: the share that is substantive (takes a checkable position on the focal proposition), the share aligned with the assigned side (the veracity denominator), and, among substantive claims, the supports and direct shares.")
```

Veracity is aggregated to the conversation level before any study/model summary
is formed, so that a long conversation with many extracted claims cannot
dominate the average. The default aligned-veracity measure pools direct claims
about the focal proposition with indirect supporting evidence (institutional
incentives, alleged anomalies, historical precedents) whenever the claim's
stance supports the assigned direction. Table \@ref(tab:rb-veracity) reports the
mean aligned veracity by study/model and direction.

```{r rb-veracity}
vera <- all_numbers |>
  dplyr::filter(section == "Veracity", block == "table_aligned_veracity",
                term == "aligned_veracity") |>
  dplyr::mutate(Direction = dplyr::recode(direction, bunk = "Bunk", debunk = "Debunk"),
                Conversations = n,
                `Mean aligned veracity` = fmt_num(estimate, 1)) |>
  dplyr::select(`Study / model` = model, Direction, Conversations,
                `Mean aligned veracity`)
si_kable_styled(vera, caption = "Mean veracity (0 = false, 100 = true) of substantive direction-aligned claims, by study/model and assigned direction. Conversations is the number of conversations contributing at least one aligned claim; veracity is aggregated to the conversation level before averaging.")
```

```{r rb-fig-veracity, fig.cap="Mean veracity of substantive claims aligned with the assigned persuasive direction, by study, model, and direction."}
fig_veracity()
```

The mean conceals the tail: even where average aligned veracity is high, a
non-trivial fraction of individual claims are scored as predominantly false. Table
\@ref(tab:rb-veracity-tail) reports, for each study and assigned direction, the
share of fact-checkable claims falling into successively lower veracity bands
(below 10, below 20, and below a more-false-than-true cutoff of 40) for both the
direction-aligned (headline) and all fact-checked claim sets for Studies 1-3. The
low-veracity tail is heavier under bunking than debunking. In Study 4 it is also
strongly model-dependent, so rather than collapse the four frontier models into a
single Study-4 cell we report the Study-4 tail rates per model in Table
\@ref(tab:rb-veracity-tail-s4models), where the high-volume Grok conversations and
the near-fully veridical GPT-5.2 conversations are kept distinct.

```{r rb-veracity-tail}
vt <- all_numbers |>
  dplyr::filter(section == "Veracity", block == "veracity_tail",
                sample == "strict_claims",
                model %in% c("Study1", "Study2", "Study3")) |>
  dplyr::select(outcome, model, direction, term, estimate) |>
  tidyr::pivot_wider(names_from = term, values_from = estimate) |>
  dplyr::mutate(
    `Claim set` = dplyr::recode(outcome, aligned = "Aligned (headline)",
                                substantive = "All fact-checked"),
    Study = dplyr::recode(model, Study1 = "S1 (Jailbroken)", Study2 = "S2 (Standard)",
                          Study3 = "S3 (Truth-Constrained)"),
    Direction = dplyr::recode(direction, bunk = "Bunk", debunk = "Debunk"),
    .csord = match(outcome, c("aligned", "substantive")),
    .stord = match(model, c("Study1", "Study2", "Study3")),
    `n claims` = scales::comma(n_claims),
    `Mean veracity` = fmt_num(mean_veracity, 1),
    `% < 10` = fmt_num(p_lt10, 1), `% < 20` = fmt_num(p_lt20, 1),
    `% < 40` = fmt_num(p_lt40, 1)) |>
  dplyr::arrange(.csord, .stord, Direction) |>
  dplyr::select(`Claim set`, Study, Direction, `n claims`,
                `Mean veracity`, `% < 10`, `% < 20`, `% < 40`)
si_kable_styled(vt, caption = "Low-veracity tail rates of the persuader's fact-checkable claims, for Studies 1-3, by assigned direction, for the direction-aligned (headline) and all fact-checked claim sets. Veracity is the per-claim fact-check score (0 = false, 100 = true), clipped to 0-100. '% < k' is the share of claims scoring below k; the low-veracity cutoff is 40 (a generous more-false-than-true threshold). Study 4 is reported per frontier model in the next table rather than as a single pooled cell, because the tail rates differ sharply across the four models.")
```

```{r rb-veracity-tail-s4models}
vt4 <- all_numbers |>
  dplyr::filter(section == "Veracity", block == "veracity_tail",
                sample == "strict_claims", grepl("^Study4_", model)) |>
  dplyr::select(outcome, model, direction, term, estimate) |>
  tidyr::pivot_wider(names_from = term, values_from = estimate) |>
  dplyr::mutate(
    `Claim set` = dplyr::recode(outcome, aligned = "Aligned (headline)",
                                substantive = "All fact-checked"),
    Model = sub("^Study4_", "", model),
    Direction = dplyr::recode(direction, bunk = "Bunk", debunk = "Debunk"),
    .csord = match(outcome, c("aligned", "substantive")),
    .mord = match(Model, c("Claude", "Gemini", "GPT-5.2", "Grok")),
    `n claims` = scales::comma(n_claims),
    `Mean veracity` = fmt_num(mean_veracity, 1),
    `% < 10` = fmt_num(p_lt10, 1), `% < 20` = fmt_num(p_lt20, 1),
    `% < 40` = fmt_num(p_lt40, 1)) |>
  dplyr::arrange(.csord, .mord, Direction) |>
  dplyr::select(`Claim set`, Model, Direction, `n claims`,
                `Mean veracity`, `% < 10`, `% < 20`, `% < 40`)
si_kable_styled(vt4, caption = "Study-4 low-veracity tail rates by frontier model and assigned direction (claim-pooled within model), for the aligned (headline) and all fact-checked claim sets. Same metrics and cutoff as the previous table. The wide spread across models (GPT-5.2 near-fully veridical versus Grok largely low-veracity) is why Study 4 is reported per model here rather than collapsed into a single pooled cell.")
```

## Veracity Versus Persuasion

A central interpretive question is whether more truthful arguments are more
persuasive. We answer it at two levels of aggregation, because the two levels
can disagree. At the *cross-cell* (ecological) level, each study/model-by-direction
cell contributes one point: its mean aligned veracity and its mean belief change.
At the *within-model* level, the unit is the individual conversation, and we ask
whether a participant exposed to more truthful aligned claims moved more.

```{r rb-veracity-corr}
vbc <- all_numbers |>
  dplyr::filter(section == "S4 + pooled", block == "veracity_belief_correlations",
                sample %in% c("strict_n1272", "compliant_n1073"), term == "pearson_r") |>
  dplyr::mutate(Sample = dplyr::recode(sample,
                  strict_n1272 = "Observed outcomes", compliant_n1073 = "Compliant"),
                `r [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                p = fmt_p(p_value)) |>
  dplyr::arrange(model, direction, dplyr::desc(Sample)) |>
  dplyr::select(Model = model, Direction = direction, Sample, n, `r [95% CI]`, p)
si_kable_styled(vbc, caption = "Within-cell Pearson correlation between conversation-level aligned-claim veracity and direction-aligned belief change (Study 4), in the strict and compliant samples.", digits = 2)
```

```{r rb-fig-veracity-persuasion, fig.cap="Cross-cell (ecological) relationship between mean aligned-claim veracity and mean direction-aligned belief change. Each point is one Study-4 model-by-direction cell. Both the observed-outcome and compliant samples are shown, faceted side by side (Observed outcomes | Compliant)."}
fig_veracity_persuasion()
```

### Paltering: bunking persuades even with accurate claims

A sharper version of the same question asks whether bunking still moves belief when
the conversation's claims are accurate. Restricting to *compliant bunking*
conversations pooled across all four studies, Figure \@ref(fig:rb-fig-paltering) plots
each conversation's belief change against its average aligned-claim veracity. Belief
change declines only gently with veracity and remains clearly positive through the
high-veracity range (the veracity-octile means stay near +8 points even above 75/100)
-- i.e., bunking persuades in part by *paltering*, marshalling accurate claims toward
a misleading conclusion, rather than depending on outright falsehoods.

```{r rb-fig-paltering, fig.width=7.2, fig.height=4.4, fig.cap="Belief change versus conversation-average aligned-claim veracity for compliant bunking conversations, pooled across Studies 1-4. Grey points are conversations; white points are veracity-octile means with 95% CIs; the band is a loess smooth. Belief change stays positive even where the conversation's claims are highly accurate."}
fig_paltering()
```

## Topic Heterogeneity

Because each participant selects an individualized focal conspiracy, persuasion
effects can concentrate in recognizable conspiracy families. Table \@ref(tab:rb-topic) reports, for
each adequately covered topic, an HC3 regression contrast of bunking-minus-debunking
aligned belief change with a 95% confidence interval, alongside the model-implied
per-condition means. The clearest pattern is that topics with abundant genuine
real-world ambiguity (e.g., 9/11, COVID lab origin) sustain large bunking gains,
whereas a near-physically-settled topic such as the moon-landing hoax is close to
unbunkable: its bunking mean is
`r fmt_est_ci(num("topic_contrasts", sample="strict", model="Moon Landing Hoax", direction="Bunking", term="topic_mean_aligned_change"), 1)`,
versus
`r fmt_est_ci(num("topic_contrasts", sample="strict", model="9/11 Inside Job Conspiracy", direction="Bunking", term="topic_mean_aligned_change"), 1)`
for 9/11.

```{r rb-topic}
tc_contrast <- all_numbers |>
  dplyr::filter(section == "Topic", block == "topic_contrasts", sample == "strict",
                term == "bunk_vs_debunk_contrast") |>
  dplyr::transmute(Topic = model, n,
                   `Bunk − debunk [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 1),
                   p = fmt_p(p_value), .ord = estimate)
tc_means <- all_numbers |>
  dplyr::filter(section == "Topic", block == "topic_contrasts", sample == "strict",
                term == "topic_mean_aligned_change") |>
  dplyr::mutate(Direction = dplyr::recode(direction, Bunking = "bunk_m", Debunking = "debunk_m"),
                cell = fmt_ci(estimate, conf_low, conf_high, 1)) |>
  dplyr::select(Topic = model, Direction, cell) |>
  tidyr::pivot_wider(names_from = Direction, values_from = cell)
topic_tab <- tc_contrast |>
  dplyr::left_join(tc_means, by = "Topic") |>
  dplyr::arrange(.ord) |>
  dplyr::transmute(Topic, n,
                   `Bunk mean [95% CI]` = bunk_m,
                   `Debunk mean [95% CI]` = debunk_m,
                   `Bunk − debunk [95% CI]`, p)
si_kable_styled(topic_tab, caption = "Topic-specific HC3 bunk-minus-debunk regression contrast of direction-aligned belief change, with model-implied per-condition means, pooled across all four studies (embedding-based topic grouping; exploratory; adequately covered topics only).")
```

```{r rb-fig-topic, fig.height=6.4, fig.cap="Topic-specific bunk-minus-debunk regression contrast of direction-aligned belief change with HC3 95% confidence intervals, ordered by the strict-sample contrast (exploratory). Both the observed-outcome and compliant samples are shown (dodged); topics retained only where both conditions have adequate coverage in that sample."}
fig_topic_contrasts()
```

Topics differ not only in how persuadable they are but in how *truthful* the
persuader's claims are. Figure \@ref(fig:rb-fig-topic-veracity) reports the mean
direction-aligned claim veracity within each conspiracy topic cluster, separately
by arm and for both samples. Debunking claims are uniformly high-veracity, whereas
bunking veracity varies sharply by topic --- reflecting how much genuinely
suspicious real-world material a given conspiracy affords --- which is the
claim-level signature behind the topic heterogeneity in persuasion above.

```{r rb-fig-topic-veracity, fig.width=8, fig.height=6.4, fig.cap="Mean direction-aligned claim veracity (0-100, 95% CI) by conspiracy topic cluster and assigned direction, pooled across studies, on the all-conversations and compliant samples (faceted). Topics are ordered by the debunk-minus-bunk veracity gap; only clusters with sufficient conversations are shown."}
fig_topic_veracity()
```

## Public Speech

Study 4 added a public-speech battery so that persuasion could be measured not
only as private belief but as expressed, shareable content.

Table \@ref(tab:rb-prereg) reports the weighted-sharing test alongside the other
public-speech contrasts (post stance, sharing likelihood, and private belief for
reference), equal-weighted across the four models with HC3 standard errors, from
the **baseline-adjusted** models (a disclosed precision deviation from the
registration). The registered specification itself — unadjusted OLS on the
direction-aligned change scores — is reported in Table
\@ref(tab:rb-prereg-aligned). The equal-model-weighted pooled contrast is the one
place a single pooled Study-4 number is the headline; everywhere else Study 4 is
reported per model. The effect on weighted sharing is large and directional: the
conversation moves not just belief but the content people would broadcast.

```{r rb-prereg}
prereg <- all_numbers |>
  dplyr::filter(section == "S4 + pooled", block == "prereg_primary",
                sample %in% c("strict_n1272", "compliant_n1073"),
                term == "avg_bunk_minus_debunk") |>
  dplyr::mutate(Sample = dplyr::recode(sample,
                  strict_n1272 = "Observed outcomes", compliant_n1073 = "Compliant"),
                Outcome = dplyr::recode(outcome,
                  weighted_share = "Weighted sharing (registered primary)",
                  post_stance = "Post stance",
                  share_likelihood = "Sharing likelihood",
                  belief = "Private belief"),
                .ord = match(outcome,
                  c("weighted_share", "post_stance", "share_likelihood", "belief")),
                `Bunk - debunk [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                p = fmt_p(p_value)) |>
  dplyr::arrange(.ord, dplyr::desc(Sample)) |>
  dplyr::select(Outcome, Sample, n, `Bunk - debunk [95% CI]`, p)
si_kable_styled(prereg, caption = "Study 4 public-speech and belief contrasts, equal-weighted across the four models (HC3), in the strict and compliant samples. Baseline-adjusted specification (a disclosed precision deviation from the registered unadjusted model; the registered specification is reported in the direction-aligned table below). p is the HC3 t test of the equal-weighted contrast. Per-model breakdowns are reported in the adjacent tables.", digits = 2)
```

Table \@ref(tab:rb-prereg-model) disaggregates the registered family by model
family, reporting the bunk-minus-debunk contrast on each registered outcome for
Claude, Gemini, GPT-5.2, and Grok (strict and compliant samples). The registered
weighted-sharing effect is large and reliable for Claude, Gemini, and Grok but
small and---in the strict sample---non-significant for GPT-5.2, mirroring that
model's bunking defection; under the compliant sample, where only delivered
GPT-5.2 bunking conversations remain, the GPT-5.2 weighted-sharing contrast turns
positive and significant.

```{r rb-prereg-model}
pregm <- all_numbers |>
  dplyr::filter(section == "S4 + pooled", block == "prereg_primary_model_contrasts",
                sample %in% c("strict_n1272", "compliant_n1073")) |>
  dplyr::mutate(Sample = dplyr::recode(sample,
                  strict_n1272 = "Observed outcomes", compliant_n1073 = "Compliant"),
                Outcome = dplyr::recode(outcome,
                  weighted_share = "Weighted sharing",
                  post_stance = "Post stance",
                  share_likelihood = "Sharing likelihood",
                  belief = "Private belief"),
                .ord = match(outcome,
                  c("weighted_share", "post_stance", "share_likelihood", "belief")),
                `Bunk - debunk [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                p = fmt_p(p_value)) |>
  dplyr::arrange(.ord, dplyr::desc(Sample), model) |>
  dplyr::select(Outcome, Sample, Model = model, n, `Bunk - debunk [95% CI]`, p)
si_kable_styled(pregm, caption = "Study 4 public-speech and belief contrasts by model family (baseline-adjusted bunk-minus-debunk; strict and compliant samples). n is the per-model estimation sample (both arms).", digits = 2)
```

Table \@ref(tab:rb-prereg-aligned) reports the registered (AsPredicted #282517)
specification verbatim: unadjusted OLS of each direction-aligned outcome on
direction x model, HC3, among all observed-outcome participants regardless of
compliance, with positive values meaning movement
toward the assigned direction. It includes the joint and direction-by-model
interaction tests recorded in the registration, making the registered
model-difference test—the omnibus HC3 F of the direction-by-model term—explicit
alongside the average contrast.

```{r rb-prereg-aligned}
prereg2 <- all_numbers |>
  dplyr::filter(section == "S4 + pooled", block == "prereg_registered_tests",
                sample == "strict_n1272") |>
  dplyr::mutate(`Estimate [95% CI]` = ifelse(is.na(estimate), "—",
                  fmt_ci(estimate, conf_low, conf_high, 2)),
                p = fmt_p(p_value),
                Term = dplyr::recode(term,
                  avg_debunk_minus_bunk = "Avg debunk − bunk",
                  direction_joint_F = "Joint direction F",
                  direction_x_model_interaction = "Direction × model F")) |>
  dplyr::select(Outcome = outcome, Term, n, `Estimate [95% CI]`, p)
si_kable_styled(prereg2, caption = "Study 4 registered direction-aligned tests (equal-model-weighted pooled estimand, the pre-registered analysis) with joint and direction-by-model omnibus F-tests.", digits = 2)
```

### Missing-outcome analyses among confirmed chat starters {#chat-started-speech}

The registered speech analyses above use the 1,272 records with the complete
outcome battery. The missing-outcome analysis instead conditions only on confirmed
chat start ($N=1{,}353$). The 81 added baseline posts were passed through the same
three-rater orientation audit and five-rater stance pipeline as the original
analysis. Post stance, revised-post sharing, and original-post sharing are jointly
imputed with post belief in the MAR analysis; no compliant-subset, mechanism,
trust, perception, debrief, compliance, or claim-veracity outcome is imputed.

Table \@ref(tab:rb-chat-started-continuous-cells) reports model-by-direction
effects for weighted sharing, post stance, and willingness to re-share the
original post. Feasible-value bounds use $-50$ to $50$ for post weighted sharing,
0--100 for post stance, and each participant's baseline stance combined with a
0--100 original-post sharing rating.

```{r rb-chat-started-continuous-cells}
s4s_cells <- all_numbers |>
  dplyr::filter(block == "s4_chat_started_cell_effects",
                sample == "chat_started_n1353",
                outcome %in% c("aligned_new_minus_old_weighted",
                               "aligned_stance_change",
                               "aligned_original_post_share_change")) |>
  dplyr::mutate(
    Outcome = dplyr::recode(outcome,
      aligned_new_minus_old_weighted = "Weighted sharing",
      aligned_stance_change = "Post stance",
      aligned_original_post_share_change = "Original-post sharing"),
    Approach = dplyr::recode(term,
      available_outcomes = "Available outcomes",
      zero_change = "Zero change",
      mi_mar = "MAR imputation",
      support_bounds = "Feasible bounds"),
    value = dplyr::if_else(term == "support_bounds",
      paste0("[", fmt_num(conf_low, 1), ", ", fmt_num(conf_high, 1), "]"),
      fmt_ci(estimate, conf_low, conf_high, 1)),
    .o = match(outcome, c("aligned_new_minus_old_weighted",
                          "aligned_stance_change",
                          "aligned_original_post_share_change")),
    .m = match(model, c("Claude", "Gemini", "GPT-5.2", "Grok")),
    Direction = dplyr::recode(direction, bunk = "Bunk", debunk = "Debunk")
  ) |>
  dplyr::arrange(.o, .m, Direction) |>
  dplyr::select(Outcome, Model = model, Direction, Approach, value) |>
  tidyr::pivot_wider(names_from = Approach, values_from = value)
si_kable_styled(
  s4s_cells,
  caption = "Registered continuous public-speech effects by Study 4 model and direction among 1,353 confirmed chat starters. Point estimates are reported with 95% CIs; feasible-value bounds are identification intervals. Available-outcome N = 1,272."
)
```

```{r rb-chat-started-continuous-contrasts}
s4s_contrasts <- all_numbers |>
  dplyr::filter(block == "s4_chat_started_model_contrasts",
                sample == "chat_started_n1353",
                outcome %in% c("aligned_new_minus_old_weighted",
                               "aligned_stance_change",
                               "aligned_original_post_share_change")) |>
  dplyr::mutate(
    Outcome = dplyr::recode(outcome,
      aligned_new_minus_old_weighted = "Weighted sharing",
      aligned_stance_change = "Post stance",
      aligned_original_post_share_change = "Original-post sharing"),
    Approach = dplyr::recode(term,
      available_outcomes = "Available outcomes",
      zero_change = "Zero change",
      mi_mar = "MAR imputation",
      support_bounds = "Feasible bounds"),
    value = dplyr::if_else(term == "support_bounds",
      paste0("[", fmt_num(conf_low, 1), ", ", fmt_num(conf_high, 1), "]"),
      paste0(fmt_ci(estimate, conf_low, conf_high, 1), "; p ", fmt_p(p_value))),
    .o = match(outcome, c("aligned_new_minus_old_weighted",
                          "aligned_stance_change",
                          "aligned_original_post_share_change")),
    .m = match(model, c("Claude", "Gemini", "GPT-5.2", "Grok"))
  ) |>
  dplyr::arrange(.o, .m) |>
  dplyr::select(Outcome, Model = model, Approach, value) |>
  tidyr::pivot_wider(names_from = Approach, values_from = value)
si_kable_styled(
  s4s_contrasts,
  caption = "Within-model debunk-minus-bunk asymmetries for the registered continuous public-speech outcomes among confirmed chat starters. Positive values favor debunking; point rows give estimate [95% CI] and p."
)
```

```{r rb-chat-started-registered-tests}
s4s_reg <- all_numbers |>
  dplyr::filter(block == "s4_chat_started_registered_tests",
                sample == "chat_started_n1353") |>
  dplyr::mutate(
    Outcome = dplyr::recode(outcome,
      aligned_belief_change = "Focal belief",
      aligned_new_minus_old_weighted = "Weighted sharing",
      aligned_stance_change = "Post stance",
      aligned_original_post_share_change = "Original-post sharing"),
    Approach = dplyr::case_when(
      grepl("^available_outcomes", term) ~ "Available outcomes",
      grepl("^zero_change", term) ~ "Zero change",
      grepl("^mi_mar", term) ~ "MAR imputation",
      TRUE ~ "Feasible bounds"),
    component = ifelse(direction == "direction-x-model", "Direction x model", "Equal-model debunk - bunk"),
    value = dplyr::case_when(
      direction == "direction-x-model" ~ paste0("p ", fmt_p(p_value)),
      grepl("^support_bounds", term) ~ paste0("[", fmt_num(conf_low, 1), ", ", fmt_num(conf_high, 1), "]"),
      TRUE ~ paste0(fmt_ci(estimate, conf_low, conf_high, 1), "; p ", fmt_p(p_value))
    ),
    .o = match(outcome, c("aligned_belief_change",
                          "aligned_new_minus_old_weighted",
                          "aligned_stance_change",
                          "aligned_original_post_share_change"))
  ) |>
  dplyr::arrange(.o) |>
  dplyr::select(Outcome, Approach, component, value) |>
  tidyr::pivot_wider(names_from = component, values_from = value, values_fill = "---")
si_kable_styled(
  s4s_reg,
  caption = "Registered equal-model asymmetry and direction-by-model omnibus tests for all four outcomes among confirmed chat starters. MAR interaction p-values use a Rubin-pooled HC3 multivariate Wald test. Feasible bounds are identification intervals and do not define an omnibus p-value."
)
```

For belief, weighted sharing, and stance, the equal-model debunk-minus-bunk
estimates were positive under all three point-estimation approaches and their
feasible-value intervals excluded zero. Direction-by-model omnibus tests had *p* <
.001. For original-post sharing, the point-estimate intervals and feasible-value
interval included zero; the MAR interaction test had *p* = .264.

```{r rb-fig-public-speech, fig.cap="Study 4 private belief change and weighted public-sharing change by model and assigned direction. Both the observed-outcome and compliant samples are shown in each panel (solid points/lines = Observed outcomes; hollow points/dashed lines = Compliant)."}
fig_public_speech()
```

The extensive-margin analysis treats "would publicly post" as a binary outcome
and reports pro-conspiracy and anti-conspiracy posting separately. Within an arm, the
net change is the McNemar within-participant shift in posting willingness; the
between-arm comparison is a linear probability model of the same binary outcome.
Table \@ref(tab:rb-extensive) reports the within-arm net percentage-point changes
with McNemar p-values, broken out by Study-4 model, and Table
\@ref(tab:rb-extensive-contrast) reports the between-arm bunk-minus-debunk gaps by
model.

```{r rb-extensive}
ext <- all_numbers |>
  dplyr::filter(section == "Public-posting", block == "extensive_margin",
                term == "net_pp_change", !is.na(model)) |>
  dplyr::mutate(Outcome = dplyr::recode(outcome,
                  aligned_public_posting = "Direction-aligned posting"),
                .m = match(model, c("Claude","Gemini","GPT-5.2","Grok")),
                `Net pp change` = fmt_num(estimate, 1),
                p = fmt_p(p_value)) |>
  dplyr::arrange(.m, direction) |>
  dplyr::select(Model = model, Outcome, Direction = direction, n, `Net pp change`, p)
si_kable_styled(ext, caption = "Within-arm net percentage-point change in willingness to publicly post (McNemar tests), by Study-4 model (direction-aligned posting).", digits = 1)
```

```{r rb-extensive-contrast}
extc <- all_numbers |>
  dplyr::filter(section == "Public-posting", block == "extensive_margin_contrast",
                !is.na(model)) |>
  dplyr::mutate(Model = model,
                .m = match(model, c("Claude","Gemini","GPT-5.2","Grok")),
                `Bunk − debunk pp [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                p = fmt_p(p_value)) |>
  dplyr::arrange(.m) |>
  dplyr::select(Model, n, `Bunk − debunk pp [95% CI]`, p)
si_kable_styled(extc, caption = "Between-arm linear-probability contrasts on the direction-aligned extensive margin of public posting (bunk-minus-debunk), by Study-4 model.", digits = 2)
```

The confirmed-chat-start approaches were also applied to the preregistered
extensive-margin comparisons. Table \@ref(tab:rb-chat-started-extensive-margins)
reports pooled debunking changes in pro- and anti-conspiracy posting and per-model
bunking changes on both margins. Table
\@ref(tab:rb-chat-started-extensive-aligned) reports every per-model aligned change
and the bunk-minus-debunk contrasts. Binary feasible-value intervals allow each
missing post-treatment indicator to be either 0 or 1 while holding its observed
baseline fixed.

```{r rb-chat-started-extensive-margins}
s4e_margin <- all_numbers |>
  dplyr::filter(block == "s4_chat_started_extensive_margin",
                sample == "chat_started_n1353",
                outcome %in% c("pro_public_posting_change", "anti_public_posting_change"),
                (direction == "debunk" & is.na(model)) |
                (direction == "bunk" & !is.na(model))) |>
  dplyr::mutate(
    dir_order = match(as.character(.data$direction), c("debunk", "bunk")),
    .m = match(ifelse(is.na(.data$model), "All models", .data$model),
               c("All models", "Claude", "Gemini", "GPT-5.2", "Grok")),
    .o = match(as.character(.data$outcome),
               c("pro_public_posting_change", "anti_public_posting_change"))
  ) |>
  dplyr::mutate(
    Outcome = dplyr::recode(outcome,
      pro_public_posting_change = "Pro-conspiracy posting",
      anti_public_posting_change = "Anti-conspiracy posting"),
    Group = ifelse(is.na(model), "All models", model),
    Direction = dplyr::recode(direction, bunk = "Bunk", debunk = "Debunk"),
    Approach = dplyr::recode(term,
      available_outcomes = "Available outcomes",
      zero_change = "Zero change",
      mi_mar = "MAR imputation",
      support_bounds = "Feasible bounds"),
    value = dplyr::if_else(term == "support_bounds",
      paste0("[", fmt_num(conf_low, 1), ", ", fmt_num(conf_high, 1), "]"),
      paste0(fmt_ci(estimate, conf_low, conf_high, 1), "; p ", fmt_p(p_value)))
  ) |>
  dplyr::arrange(dir_order, .m, .o) |>
  dplyr::select(Direction, Group, Outcome, Approach, value) |>
  tidyr::pivot_wider(names_from = Approach, values_from = value)
si_kable_styled(
  s4e_margin,
  caption = "Pooled debunking and per-model bunking changes in pro- and anti-conspiracy public posting among confirmed chat starters. Estimates are percentage-point changes; available/zero-change p-values are McNemar tests and MAR p-values are Rubin pooled."
)
```

```{r rb-chat-started-extensive-aligned}
s4e_aligned <- all_numbers |>
  dplyr::filter(block == "s4_chat_started_extensive_margin",
                sample == "chat_started_n1353",
                outcome == "aligned_public_posting_change") |>
  dplyr::mutate(
    .m = match(as.character(.data$model), c("Claude", "Gemini", "GPT-5.2", "Grok")),
    .c = match(as.character(.data$direction),
               c("bunk", "debunk", "bunk-minus-debunk"))
  ) |>
  dplyr::mutate(
    Comparison = dplyr::recode(direction,
      bunk = "Bunk aligned change",
      debunk = "Debunk aligned change",
      `bunk-minus-debunk` = "Bunk - debunk contrast"),
    Approach = dplyr::recode(term,
      available_outcomes = "Available outcomes",
      zero_change = "Zero change",
      mi_mar = "MAR imputation",
      support_bounds = "Feasible bounds"),
    value = dplyr::if_else(term == "support_bounds",
      paste0("[", fmt_num(conf_low, 1), ", ", fmt_num(conf_high, 1), "]"),
      paste0(fmt_ci(estimate, conf_low, conf_high, 1), "; p ", fmt_p(p_value)))
  ) |>
  dplyr::arrange(.m, .c) |>
  dplyr::select(Model = model, Comparison, Approach, value) |>
  tidyr::pivot_wider(names_from = Approach, values_from = value)
si_kable_styled(
  s4e_aligned,
  caption = "Per-model direction-aligned public-posting changes and bunk-minus-debunk contrasts among confirmed chat starters. Estimates are percentage points; feasible-value bounds are identification intervals."
)
```

Pooled debunking estimates were negative for pro-conspiracy posting and positive
for anti-conspiracy posting under each approach; both feasible-value intervals
excluded zero. GPT-5.2 bunking estimates for aligned posting were negative. The
bunk-minus-debunk contrasts were negative for all four models under each approach,
and all four feasible-value intervals excluded zero.

```{r rb-fig-extensive, fig.cap="Net percentage-point change in direction-aligned willingness to post publicly, by Study-4 frontier model and assigned direction (per-model; no pooling)."}
fig_extensive_margin()
```

Post-stance composition records the categorical distribution of post types
(endorses, leans for, neutral/mixed, no stance, leans against, rejects) before
and after the conversation, for debunking across all models, bunking among
complying models, and bunking under GPT-5.2 separately.

```{r rb-post-composition}
pc <- all_numbers |>
  dplyr::filter(section == "S4 + pooled", block == "post_stance_composition",
                outcome == "post_stance_category") |>
  dplyr::mutate(Phase = ifelse(grepl("^pre_", term), "Pre", "Post"),
                Stance = sub("^(pre|post)_", "", term),
                Cell = dplyr::recode(direction,
                  debunk_all_models = "Debunk (all)",
                  bunk_complying_models = "Bunk (complying)",
                  bunk_gpt52 = "Bunk (GPT-5.2)"),
                pct = fmt_num(estimate, 1)) |>
  dplyr::select(Cell, Phase, Stance, n, `%` = pct)
si_kable_styled(pc, caption = "Public-post stance composition before and after the conversation, by cell.", digits = 1)
```

```{r rb-fig-post-composition, fig.cap="Public-post stance composition before versus after the Study 4 conversation, by cell."}
fig_post_composition()
```

```{r rb-post-form}
pf <- all_numbers |>
  dplyr::filter(section == "S4 + pooled", block == "post_form_composition",
                outcome == "post_response_type") |>
  dplyr::mutate(Phase = ifelse(grepl("^pre_", term), "Pre", "Post"),
                Mode = dplyr::recode(sub("^(pre|post)_", "", term),
                  assertion = "Assertion",
                  question_raising = "Question-raising",
                  mixed_assertion_question = "Mixed assertion / question",
                  uncertainty_statement = "Uncertainty statement",
                  declines_to_post = "Declines to post",
                  meta_task = "Meta / off-task",
                  unclassifiable = "Unclassifiable"),
                Direction = dplyr::recode(direction,
                  pooled = "Pooled", bunk = "Bunk", debunk = "Debunk"),
                `%` = fmt_num(estimate, 1)) |>
  dplyr::select(Direction, Phase, Mode, n, `%`)
si_kable_styled(pf, caption = "Communicative-mode composition of the Study 4 public post before versus after the conversation, by assigned direction (five-rater consensus; shares within cell).", digits = 1)
```

```{r rb-fig-post-form, fig.cap="Communicative mode of the Study 4 public post before versus after the conversation, by assigned direction."}
fig_post_form_composition()
```

### Translating private belief into public expression

The belief and public-sharing effects diverge sharply, but this could in principle
reflect nothing more than the average belief effects. To test this directly, we
related each participant's direction-aligned change in private belief to their
direction-aligned change in weighted public sharing (compliant subset, both signed
toward the AI's assigned position).

```{r rb-dissoc-stats, include=FALSE}
.ds <- core_objects$s4$s4_compliant
.ds <- data.frame(cond = .ds$direction,
                  dB = .ds$aligned_belief_change,
                  dS = .ds$aligned_new_minus_old_weighted)
.ds <- .ds[is.finite(.ds$dB) & is.finite(.ds$dS), ]
.slope <- function(cc) unname(coef(lm(dS ~ dB, data = subset(.ds, cond == cc)))[2])
.sl_b <- .slope("bunk"); .sl_d <- .slope("debunk")
.rcor  <- function(cc) cor(subset(.ds, cond == cc)$dB, subset(.ds, cond == cc)$dS)
.r_b <- .rcor("bunk"); .r_d <- .rcor("debunk")
.rlab <- function(r) sub("^0", "", sprintf("%.2f", r))
.sub  <- subset(.ds, dB <= 0)                       # belief did not move toward the AI
.m0   <- lm(dS ~ cond, data = .sub)                 # cond: bunk (ref) vs debunk
.ci0  <- confint(.m0)["conddebunk", ]
.est0 <- unname(coef(.m0)["conddebunk"]); .p0 <- summary(.m0)$coefficients["conddebunk", 4]
.n0b  <- sum(.sub$cond == "bunk"); .n0d <- sum(.sub$cond == "debunk")
.p0lab <- if (.p0 < .001) "< .001" else paste0("= ", formatC(.p0, format = "f", digits = 3))
```

Among compliant conversations the two arms moved private belief toward their
assigned positions by similar amounts, and the *within-person* translation of
belief into public sharing was likewise similar (OLS slope `r round(.sl_b, 2)`
for bunking and `r round(.sl_d, 2)` for debunking; Pearson *r* = `r .rlab(.r_b)`
and `r .rlab(.r_d)`, respectively; Fig.
\@ref(fig:rb-fig-dissociation)A). The asymmetry instead took the form of a
belief-independent shift in public expression. Among participants whose private
belief did **not** move toward the AI's position (direction-aligned belief change
$\leq 0$; *n* = `r .n0b` bunking, `r .n0d` debunking), debunking still shifted weighted
sharing toward its (anti-conspiracy) side by `r round(.est0, 1)` points more than
bunking shifted it toward its (pro-conspiracy) side (95% CI
[`r round(.ci0[1], 1)`, `r round(.ci0[2], 1)`], *p* `r .p0lab`; Fig.
\@ref(fig:rb-fig-dissociation)B) — a gap indistinguishable from the full compliant
sample. Indeed, participants whose belief backfired under debunking nonetheless
distanced themselves publicly from the conspiracy. Debunking thus suppresses the
public endorsement of conspiracies even when it does not change, or even reverses,
private belief — a pattern consistent with awareness that conspiracies are widely
regarded as dubious, such that the prospect of public expression invokes
reputational or normative considerations that operate independently of private
conviction.

```{r rb-fig-dissociation, fig.width=9, fig.height=4.4, fig.cap="Study 4 (compliant subset): translation of private belief change into public (weighted) sharing change, both signed toward the AI's assigned side. (A) Each point is a conversation; lines are per-condition OLS fits with 95% CIs -- the within-person slope is similar in the two arms. (B) Mean change in weighted sharing by belief-change stratum (backfire, no change, moved toward the AI) and condition, with 95% CIs; debunking shifts public sharing toward its side more than bunking does even among participants whose private belief did not move toward the AI (left two strata)."}
fig_belief_sharing_dissociation()
```

### Robustness: the neutral-50 imputation

```{r rb-imputation, include=FALSE}
.s4i <- core_objects$s4$s4
.v2i <- readr::read_csv(core_objects$paths$s4_stance_v2, show_col_types = FALSE, progress = FALSE) |>
  dplyr::filter(timepoint == "post") |>
  dplyr::transmute(ResponseId, post_applicable = consensus_applicable %in% c("True", "TRUE"))
.si <- .s4i |>
  dplyr::left_join(.v2i, by = "ResponseId") |>
  dplyr::filter(!is.na(aligned_new_minus_old_weighted), !is.na(direction))
.asym <- function(d) {
  m <- d |> dplyr::group_by(direction) |>
    dplyr::summarise(mw = mean(aligned_new_minus_old_weighted), .groups = "drop")
  m$mw[m$direction == "debunk"] - m$mw[m$direction == "bunk"]
}
.rate <- .si |> dplyr::group_by(direction) |>
  dplyr::summarise(pct = 100 * mean(!post_applicable, na.rm = TRUE), .groups = "drop")
imp_bunk <- fmt_num(.rate$pct[.rate$direction == "bunk"], 1)
imp_deb  <- fmt_num(.rate$pct[.rate$direction == "debunk"], 1)
imp_all  <- fmt_num(.asym(.si), 1)
imp_excl <- fmt_num(.asym(dplyr::filter(.si, post_applicable)), 1)
```

The registered weighted-sharing outcome assigns the neutral midpoint (stance 50,
and hence *zero* direction weight) to posts the classifier judges off-topic or
declining (Methods). Because bunking-arm models more often decline or go off-topic
in some conditions, one could worry that this imputation mechanically dampens the
bunking sharing effect and inflates the apparent truth advantage in public speech.
It does not. The imputation rate is, if anything, slightly *lower* under bunking
(`r imp_bunk`% of bunking posts versus `r imp_deb`% of debunking posts are imputed
at post), so it cannot differentially suppress the bunking arm. And *excluding* the
imputed posts entirely --- rather than down-weighting them to zero --- slightly
*increases* the debunk-minus-bunk asymmetry (from `r imp_all` to `r imp_excl`
points). The imputation is therefore mildly conservative against the truth-advantage
conclusion, not responsible for it. (Setting the imputed posts' share likelihood to
zero is equivalent to the registered rule, since a zero direction weight already
contributes nothing to the weighted share.)

## Stance-Classifier Validation

The five-model stance ensemble that underlies the weighted-sharing and post-stance
outcomes is the primary stance measure; its inter-rater reliability is reported in
the Methods (Section "Five-Rater Post-Stance Scoring," Table
\@ref(tab:me-stance-reliability)). Here we validate it against an independent human
gold-coding of 150 Study-4 social-media posts, coded by an author blind to the model
ratings on the same 0--100 stance scale and six-level stance category.

The ensemble consensus tracks the human coder closely: a Spearman rank correlation of
`r fmt_num(est("gold_coding", model="consensus", term="spearman_r"), 2)`
(Pearson `r fmt_num(est("gold_coding", model="consensus", term="pearson_r"), 2)`) on the
0--100 stance score, a mean absolute error of
`r fmt_num(est("gold_coding", model="consensus", term="mae"), 1)` points on that 100-point
scale, exact agreement on the six-level stance category in
`r scales::percent(est("gold_coding", model="consensus", term="pct_exact_category"), accuracy=1)`
of posts, and for/against directional agreement in
`r scales::percent(est("gold_coding", model="consensus", term="pct_directional"), accuracy=1)`
of the posts that both the human and the model call directional. Every individual
panel model reaches a comparable level (Table \@ref(tab:rb-gold-validation)): Gemini
agrees most closely with the human coder and DeepSeek least, but all five sit well
within the range expected from the ensemble's own high inter-rater reliability, and
the disagreement that remains is concentrated in the boundary between adjacent stance
categories (e.g. *leans-for* vs *argues-for*) rather than in the for/against direction
that the sharing analyses depend on.

```{r rb-gold-validation}
gv <- all_numbers |>
  dplyr::filter(block == "gold_coding", section == "Stance validation",
                !is.na(estimate)) |>
  dplyr::select(model, term, estimate) |>
  tidyr::pivot_wider(names_from = term, values_from = estimate) |>
  dplyr::mutate(
    Model = dplyr::recode(model, consensus = "Ensemble consensus", claude = "Claude",
                          gpt = "GPT", gemini = "Gemini", grok = "Grok", deepseek = "DeepSeek"),
    Spearman = fmt_num(spearman_r, 2), Pearson = fmt_num(pearson_r, 2),
    MAE = fmt_num(mae, 1),
    `Exact category` = scales::percent(pct_exact_category, accuracy = 1),
    `Directional (for/against)` = scales::percent(pct_directional, accuracy = 1),
    .so = match(model, c("consensus", "claude", "gpt", "gemini", "grok", "deepseek"))) |>
  dplyr::arrange(.so) |>
  dplyr::select(Model, Spearman, Pearson, MAE, `Exact category`, `Directional (for/against)`)
si_kable_styled(gv, caption = "Stance-classifier validation against an independent human gold-coding of 150 Study-4 posts: the five-rater ensemble consensus and each panel model versus the human coder. Spearman and Pearson are correlations on the 0-100 stance score; MAE is the mean absolute error on that scale; Exact category is exact agreement on the six-level stance category; Directional is for/against agreement among posts both the human and model call directional.")
```

## Veracity of the Focal Conspiracies

```{r rb-focal-flag, include=FALSE}
fv_has <- any(all_numbers$block == "focal_veracity" & all_numbers$term == "pct_false" &
                !is.na(all_numbers$estimate))
```

```{r rb-focal-intro, results='asis', echo=FALSE}
if (fv_has) cat("
The analyses above fact-check the *persuader's* claims. A complementary question
is how epistemically suspect the focal conspiracies themselves were: what share
of the propositions our participants were uncertain about are, on the best
available public evidence, false? Each analytic participant's focal restatement
--- the declarative statement that anchored their belief ratings --- was scored
0--100 for veracity by a live-web fact-checking judge (perplexity/sonar-pro,
temperature 0), after first classifying whether the restatement asserts a
conspiracy claim at all (denial-phrased and no-claim restatements are excluded
from the headline shares; a sensitivity folding denials back in with flipped
orientation is reported alongside). Scores are frozen measurement data under
`data/api_cached/focal_veracity/`; the generating pipeline is
`code/provenance/python_api/focal_veracity/`.
")
```

```{r rb-focal-veracity, eval=fv_has}
fvt <- all_numbers |>
  dplyr::filter(block == "focal_veracity",
                term %in% c("share_conspiracy_claim", "mean_veracity", "pct_false",
                            "pct_uncertain", "pct_true", "pct_false_sens_flipdenials")) |>
  dplyr::select(model, term, n, estimate) |>
  tidyr::pivot_wider(names_from = term, values_from = c(estimate, n)) |>
  dplyr::mutate(
    Study = model,
    `Conspiracy claims (% of scored)` = fmt_num(estimate_share_conspiracy_claim, 1),
    `n claims` = n_mean_veracity,
    `Mean veracity` = fmt_num(estimate_mean_veracity, 1),
    `% false (<40)` = fmt_num(estimate_pct_false, 1),
    `% uncertain [40,60)` = fmt_num(estimate_pct_uncertain, 1),
    `% true (>=60)` = fmt_num(estimate_pct_true, 1),
    `% false (sens., denials flipped)` = fmt_num(estimate_pct_false_sens_flipdenials, 1),
    .so = match(model, c("Jailbroken", "Standard", "Truth-Constrained", "Study 4", "All studies"))) |>
  dplyr::arrange(.so) |>
  dplyr::select(Study, `Conspiracy claims (% of scored)`, `n claims`, `Mean veracity`,
                `% false (<40)`, `% uncertain [40,60)`, `% true (>=60)`,
                `% false (sens., denials flipped)`)
si_kable_styled(fvt, caption = "Veracity of the focal conspiracy statements themselves, by study. Each analytic participant's focal restatement was classified (conspiracy claim / official account / no claim) and, when it asserts a claim, fact-checked 0-100 by perplexity/sonar-pro (temperature 0). Shares are computed among genuine conspiracy-claim restatements; the sensitivity column folds denial-phrased restatements back in with flipped orientation (100 - v).")
```

```{r rb-focal-subset, eval=fv_has}
fvs <- all_numbers |>
  dplyr::filter(block == "focal_veracity_subset",
                term %in% c("bunk_minus_debunk_ancova", "debunk_minus_bunk_equal_weighted")) |>
  dplyr::mutate(
    Subset = dplyr::recode(outcome, false_conspiracies = "Clearly false (<40)",
                           nonfalse_conspiracies = "Not clearly false (>=40)"),
    Contrast = dplyr::recode(term,
                             bunk_minus_debunk_ancova = "Bunk - debunk (ANCOVA)",
                             debunk_minus_bunk_equal_weighted = "Debunk - bunk (equal-weighted)"),
    `Estimate [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
    p = fmt_p(p_value)) |>
  dplyr::arrange(match(model, c("Jailbroken", "Standard", "Truth-Constrained", "Study 4")), Subset) |>
  dplyr::select(Study = model, Subset, Contrast, n, `Estimate [95% CI]`, p)
si_kable_styled(fvs, caption = "Bunking and debunking within veracity subsets of the focal conspiracies. Studies 1-3: the primary-spec additive ANCOVA contrast (aligned change ~ condition + centered baseline, HC3) fit within the clearly-false (<40) and not-clearly-false subsets of genuine conspiracy claims. Study 4: the equal-model-weighted direction contrast, HC3, within the same subsets. Exploratory; subset membership is defined by the focal-statement veracity scores.")
```
