# Pooled treatment-effect heterogeneity: causal forest

As a pre-registered exploratory analysis (AsPredicted #218585, #224184; S3
exploratory), we estimated treatment-effect heterogeneity with a causal forest
(`grf`), here extended to **pool the compliant samples of all four studies**
(*N* = `r fmt_n(est("cf_sample", term="n_total"))`; Studies 1--3
*n* = `r fmt_n(est("cf_sample", term="n_s13"))`, Study 4
*n* = `r fmt_n(est("cf_sample", term="n_s4"))`). Pooling Study 4 and the four
frontier models into a forest was not itself pre-registered and is exploratory.

Because no study includes a no-persuasion control arm, the causal contrast is
**bunking versus debunking**: the forest treats assignment to bunk
(*W* = 1; *n* = `r fmt_n(est("cf_sample", term="n_bunk"))`) versus debunk
(*W* = 0; *n* = `r fmt_n(est("cf_sample", term="n_debunk"))`) as the treatment and
the raw, orientation-corrected belief change *toward* the conspiracy (post minus
pre) as the outcome. The resulting conditional average treatment effect (CATE) is
the **bunk-minus-debunk swing** --- the total distance the AI moves a person's
belief in whichever direction it is assigned, equal to the sum of that person's
expected aligned movement under bunking and under debunking. Direction-aligned
change cannot serve as the outcome here, because it already encodes the assigned
arm in its sign. We entered every pre-treatment moderator measured in all four
studies --- baseline belief; trust in, familiarity with, and frequency of using
generative AI; party identification and the social and economic ideology items
(coded 1 = very conservative to 5 = very liberal);
religiosity; education; age; gender; and race --- plus generic conspiracist
mentality (GCBS; Studies 1--3 only, handled by the forest's native
missing-attribute splitting) and one-hot indicators for the seven study/model
strata and the participant's conspiracy topic (the canonical 18-topic
assignment; topics with fewer than `r fmt_n(40)` conversations and the
"Mixed / Unclassified" residual are lumped). Continuous moderators were
standardized within study/model; assignment
propensities were fixed to the known within-cell randomization probabilities; the
forest used honest splitting and `r fmt_n(4000)` trees.

The forest recovered the expected average swing (AIPW ATE =
`r fmt_num(est("cf_ate", term="ATE"), 1)` belief points, SE
`r fmt_num(num("cf_ate", term="ATE")$se, 2)`, matching the unadjusted bunk-debunk
gap of `r fmt_num(est("cf_ate", term="simple_gap"), 1)`) and detected substantial,
statistically reliable heterogeneity around it: the calibration test's
differential-forest-prediction coefficient was
`r fmt_num(est("cf_calibration", term="differential.forest.prediction"), 2)`
(SE `r fmt_num(num("cf_calibration", term="differential.forest.prediction")$se, 2)`,
*p* `r fmt_p(pv("cf_calibration", term="differential.forest.prediction"))`),
rejecting CATE homogeneity, and the honest-split AUTOC was
`r fmt_num(est("cf_autoc", term="AUTOC"), 2)`
(SE `r fmt_num(num("cf_autoc", term="AUTOC")$se, 2)`). Sorting participants by
estimated CATE, the AIPW swing rose monotonically from
`r fmt_num(est("cf_cate_quartile", term="Q1"), 1)` belief points in the
least-movable quartile to `r fmt_num(est("cf_cate_quartile", term="Q4"), 1)` in the
most-movable.

```{r cf-blp-tab}
cf_person_tab <- ALL_NUMBERS |>
  dplyr::filter(.data$block == "cf_blp_person") |>
  dplyr::mutate(
    Moderator = unname(cf_mod_labels[.data$term]),
    `Slope [95% CI]` = fmt_ci(.data$estimate, .data$conf_low, .data$conf_high, 2),
    p = fmt_p(.data$p_value),
    .abs = abs(.data$estimate)
  ) |>
  dplyr::arrange(dplyr::desc(.data$.abs)) |>
  dplyr::select(Moderator, `Slope [95% CI]`, p)
si_kable_styled(cf_person_tab, caption = "Best linear projection of the bunk-debunk swing on participant-level pre-treatment moderators. Slopes are in belief-scale points: per within-study SD for continuous moderators, and per category for binary moderators (female vs male, White vs other). Positive = a larger swing (more persuadable in whichever direction the AI argues). Generic conspiracist mentality (GCBS) is measured only in Studies 1-3 and projected on that subset.")
```

Among person-level traits (Table \@ref(tab:cf-blp-tab); Figure
\@ref(fig:cf-blp-fig)), **trust in AI** was the strongest moderator
(`r fmt_full(num("cf_blp_person", term="ai_trust"), 1, signed=TRUE)`): people who
trust AI more are more movable in whichever direction it argues. Heavier **AI
users** were more resistant
(`r fmt_full(num("cf_blp_person", term="ai_use"), 1, signed=TRUE)`), as were the
more **educated** (`r fmt_full(num("cf_blp_person", term="education"), 1, signed=TRUE)`);
higher **baseline belief** attenuated the swing
(`r fmt_full(num("cf_blp_person", term="baseline_belief"), 1, signed=TRUE)`).
**Women** showed a larger swing than men
(`r fmt_full(num("cf_blp_person", term="female"), 1, signed=TRUE)`). Generic
conspiracist mentality (GCBS) was the most important splitting variable, with more
conspiracy-minded participants somewhat harder to move overall
(`r fmt_full(num("cf_blp_person", term="gcbs"), 1, signed=TRUE)`). Political
identification, social and economic ideology, AI familiarity, race, religiosity,
and age were not reliable moderators of the pooled swing.

```{r cf-blp-fig, fig.width=7.2, fig.height=6, fig.cap="Best linear projection of the bunk-debunk swing (causal-forest CATE) on participant traits (top) and study/model context (bottom; reference = standard GPT-4o). Points are projection slopes with 95% confidence intervals; continuous traits are scaled per within-study SD. Coloured points are reliable at p < .05."}
fig_cf_blp()
```

```{r cf-importance-fig, fig.width=7.2, fig.height=5, fig.cap="Causal-forest variable importance (split frequency) for the bunk-debunk swing. Person-level traits in blue, study/model context in amber, conspiracy-topic indicators in green. Conspiracy mentality (GCBS) and AI trust dominate; note that GCBS is missing for Study 4 and so partly proxies study membership (see the robustness forest)."}
fig_cf_importance()
```

Heterogeneity by study and model reproduces the paper's headline contrasts in one
framework (Figure \@ref(fig:cf-swing-fig)). Relative to standard GPT-4o, the swing
was `r fmt_full(num("cf_blp_context", term="Truth-Constrained"), 1, signed=TRUE)`
points under the truth constraint and indistinguishable when jailbroken
(`r fmt_full(num("cf_blp_context", term="Jailbroken"), 1, signed=TRUE)`). Among
frontier models the swing was substantially larger for Claude
(`r fmt_full(num("cf_blp_context", term="Claude"), 1, signed=TRUE)`), Gemini
(`r fmt_full(num("cf_blp_context", term="Gemini"), 1, signed=TRUE)`), and Grok
(`r fmt_full(num("cf_blp_context", term="Grok"), 1, signed=TRUE)`), but smaller for
GPT-5.2 (`r fmt_full(num("cf_blp_context", term="GPT-5.2"), 1, signed=TRUE)`), which
essentially declined to bunk.

```{r cf-swing-fig, fig.width=7.2, fig.height=4, fig.cap="Bunk-debunk swing by study and model (AIPW within-stratum treatment effect, 95% CI); the dashed line marks the pooled average. The three GPT-4o variants (Studies 1-3) and the four Study-4 frontier models are shown separately."}
fig_cf_swing_by_model()
```

**Arm-specific structure.** Per-arm regression forests of aligned change clarify
effects the swing can mask (descriptive, not causal). Baseline belief acts
oppositely by arm --- a ceiling on bunking
(`r fmt_full(num("cf_perarm", term="baseline_belief", direction="Bunking"), 1, signed=TRUE)`
per SD) and a floor on debunking
(`r fmt_full(num("cf_perarm", term="baseline_belief", direction="Debunking"), 1, signed=TRUE)`),
so its modest net swing reflects partial cancellation. The female effect is a
debunking effect
(`r fmt_full(num("cf_perarm", term="female", direction="Debunking"), 1, signed=TRUE)`;
bunking `r fmt_full(num("cf_perarm", term="female", direction="Bunking"), 1, signed=TRUE)`):
women were more *debunkable*, not differentially bunkable.

```{r cf-cate-fig, fig.width=8.5, fig.height=5, fig.cap="Distribution of the estimated bunk-debunk swing (CATE) across participants, by study and model. Mass to the right of zero indicates participants the AI can move toward its assigned position; the spread within each panel is the heterogeneity the forest detects."}
fig_cf_cate_distribution()
```

## Which direction does the AI favor?

The swing CATE is a magnitude --- how far a person can be moved in whichever
direction the AI is assigned --- and does not by itself say whether *bunking* or
*debunking* is the more effective lever for a given person. To recover direction we
predict each participant's aligned change under each arm from a separate per-arm
regression forest, and ask which is larger (descriptive, not a within-person causal
contrast). Across the pooled sample the two levers are close to balanced:
bunking is the stronger predicted lever for
`r fmt_pct(est("cf_direction", model="Pooled", term="pct_favor_bunk"), 0)` of
participants and debunking for the rest, mirroring the near-equal average aligned
movement under the two arms
(`r fmt_num(est("cf_direction", model="Pooled", term="bunk_aligned"),1)` vs
`r fmt_num(est("cf_direction", model="Pooled", term="debunk_aligned"),1)` belief
points). The balance shifts sharply by model (Figure \@ref(fig:cf-direction-fig)):
bunking is favored for
`r fmt_pct(est("cf_direction", model="Claude", term="pct_favor_bunk"), 0)` of Claude
participants but only
`r fmt_pct(est("cf_direction", model="GPT-5.2", term="pct_favor_bunk"), 0)` of
GPT-5.2 participants --- GPT-5.2 is a much stronger debunker than bunker (its bunking
backfires; observed aligned change
`r fmt_num(est("cf_direction", model="GPT-5.2", term="bunk_aligned"),1)` bunking vs
`r fmt_num(est("cf_direction", model="GPT-5.2", term="debunk_aligned"),1)`
debunking).

```{r cf-direction-fig, fig.width=9, fig.height=5.5, fig.cap="For each participant, the forest-predicted aligned effect of bunking (x) versus debunking (y), by study and model. Points below the dashed parity line are people for whom debunking is the stronger lever (blue); points above, bunking (pink). The Claude cloud sits largely below the line (bunkable), GPT-5.2 largely above (debunkable); the GPT-4o variants are mixed."}
fig_cf_direction()
```

## Heterogeneity by conspiracy topic

Entering the participant's conspiracy topic lets the swing vary by subject matter
(Figure \@ref(fig:cf-topic-fig)). The swing is largest for the
9/11 Inside Job conspiracy
(`r fmt_num(est("cf_swing_by_topic", model="9/11 Inside Job Conspiracy", term="swing"),1)`
belief points) and the deliberate-COVID-lab-release theory
(`r fmt_num(est("cf_swing_by_topic", model="Deliberate COVID-19 Lab Release", term="swing"),1)`),
both above the pooled average, and smallest for the Area 51 alien cover-up
(`r fmt_num(est("cf_swing_by_topic", model="Area 51 Alien Cover-up", term="swing"),1)`)
and election-fraud beliefs
(`r fmt_num(est("cf_swing_by_topic", model="Election fraud conspiracy beliefs", term="swing"),1)`).
The direction favored also tracks the topic: the Moon Landing Hoax is the most
bunking-favored topic
(`r fmt_pct(est("cf_swing_by_topic", model="Moon Landing Hoax", term="pct_favor_bunk"),0)`
of its participants), consistent with the paper's truthful-ammunition account ---
topics with little genuine real-world evidence are the ones an AI can most readily
push *toward* the conspiracy.

```{r cf-topic-fig, fig.width=7.2, fig.height=4, fig.cap="AIPW bunk-debunk swing by conspiracy topic (95% CI), for topics with at least 60 conversations; the dashed line is the pooled average. Topics with more genuine real-world ammunition (9/11, COVID-lab) show the largest swings."}
fig_cf_swing_by_topic()
```

**Robustness.** Refitting on only the fully observed common covariates (dropping
GCBS and the study/model indicators) left conclusions unchanged: ATE =
`r fmt_num(est("cf_robustness", term="ATE"), 1)` belief points, and a
differential-forest-prediction coefficient of
`r fmt_num(est("cf_robustness", term="differential.forest.prediction"), 2)`
(SE `r fmt_num(num("cf_robustness", term="differential.forest.prediction")$se, 2)`,
*p* `r fmt_p(pv("cf_robustness", term="differential.forest.prediction"))`), so the
person-level moderation is not an artifact of the forest's missing-data handling or
of differences between the study/model strata.

## Separate forests within each study group

To read GCBS free of the pooled forest's missingness/study confound (it is absent
in Study 4, so "GCBS missing" otherwise proxies study membership), and to check
that the person-level pattern holds inside each design, we refit the forest
*within* Studies 1--3 (GPT-4o variants; GCBS fully observed) and *within* Study 4
(frontier models). Both groups retain clear heterogeneity (differential-forest-
prediction = `r fmt_num(est("cf_sub_calibration", sample="compliant_s1s3", term="differential.forest.prediction"), 2)`,
*p* `r fmt_p(pv("cf_sub_calibration", sample="compliant_s1s3", term="differential.forest.prediction"))`
in Studies 1--3; `r fmt_num(est("cf_sub_calibration", sample="compliant_s4", term="differential.forest.prediction"), 2)`,
*p* `r fmt_p(pv("cf_sub_calibration", sample="compliant_s4", term="differential.forest.prediction"))`
in Study 4), with a larger average swing in Study 4
(`r fmt_num(est("cf_sub_ate", sample="compliant_s4", term="ATE"), 1)` vs
`r fmt_num(est("cf_sub_ate", sample="compliant_s1s3", term="ATE"), 1)` belief points).

The split confirms the GCBS caveat directly: with GCBS observed for every
participant, its variable importance falls to
`r fmt_num(est("cf_sub_importance", sample="compliant_s1s3", term="gcbs"), 3)`
within Studies 1--3 (third behind AI trust and economic ideology), far below its
inflated pooled value of
`r fmt_num(est("cf_importance", term="gcbs"), 3)`, and its linear association with
the swing is modest and only marginal
(`r fmt_full(num("cf_sub_blp_person", sample="compliant_s1s3", term="gcbs"), 1, signed=TRUE)`).
The headline AI-receptivity axis, by contrast, replicates *independently* in both
designs: higher AI trust predicts a larger swing
(`r fmt_full(num("cf_sub_blp_person", sample="compliant_s1s3", term="ai_trust"), 1, signed=TRUE)`
in Studies 1--3;
`r fmt_full(num("cf_sub_blp_person", sample="compliant_s4", term="ai_trust"), 1, signed=TRUE)`
in Study 4) and heavier AI use a smaller one
(`r fmt_full(num("cf_sub_blp_person", sample="compliant_s4", term="ai_use"), 1, signed=TRUE)`
in Study 4). The female effect is carried by Studies 1--3
(`r fmt_full(num("cf_sub_blp_person", sample="compliant_s1s3", term="female"), 1, signed=TRUE)`)
and is not reliable within Study 4's smaller sample.

```{r cf-subgroup-fig, fig.width=7.2, fig.height=5, fig.cap="Person-trait moderators of the bunk-debunk swing, from forests fit separately within Studies 1-3 (blue) and within Study 4 (pink), with 95% confidence intervals. GCBS (conspiracy mentality) appears only for Studies 1-3, where it is measured. The AI-trust (positive) and AI-use (negative) effects replicate across both designs."}
fig_cf_subgroup_blp()
```
