# Methods

## Common Experimental Paradigm and Direction-Aligned Coding

All four studies share one paradigm. A participant completes a baseline survey,
describes in free text a conspiracy theory about which they are genuinely
uncertain, and rates their belief in an AI-generated restatement of that
conspiracy on a 0--100 scale. They are then randomly assigned to a multi-turn
conversation in which the AI is instructed to argue either *for* the conspiracy
("bunking") or *against* it ("debunking"), and they rate belief in the same
restatement again afterward. Random assignment governs only the direction of the
AI's argument; the focal topic itself is participant-selected, so each
conversation is personally relevant while remaining experimentally manipulated.

Two coding conventions make outcomes comparable across the bunking and
debunking arms. First, the belief scale is oriented so that higher values always
indicate greater conspiracy belief, applying a restatement-orientation
correction (`_rc`). The correction is needed because the conspiracy restatement
shown to each participant is LLM-generated, and a small minority of restatements
were phrased as *denials* of the conspiracy rather than affirmations; for those
items a high rating meant *low* conspiracy belief, the reverse of the default
orientation. The belief item is a single 0--100 slider anchored at 0
("Definitely false"), 50 ("Uncertain"), and 100 ("Definitely true"); the
correction rule is $b_{rc} = 100 - b$ for the subset of restatements phrased
against the focal proposition and $b_{rc} = b$ otherwise. This reversal was rare.
Second,
belief change is *direction-aligned*: it is signed so that
positive values denote movement toward the position the AI was assigned to
advocate. Concretely, for a post-minus-pre difference $\Delta$ on the corrected
scale,

\[
\text{aligned change} =
\begin{cases}
+\Delta & \text{bunking (AI argues for the conspiracy)}\\
-\Delta & \text{debunking (AI argues against the conspiracy)}.
\end{cases}
\]

The same
alignment rule is applied to every secondary outcome that has a natural
direction (stance scores, weighted sharing, posting willingness), so that
"aligned" always means "toward the assigned AI position."

## Sample construction and screening

Each study filters respondents to those who supplied a usable focal conspiracy
and whose baseline belief reflected genuine uncertainty rather than a settled
position. Because the conversational partner is itself an LLM, the screening
process serves two purposes at once: it enforces the substantive inclusion
criteria (a genuinely equivocal focal conspiracy, a baseline belief away from the
scale's floor and ceiling) and it removes inattentive humans and possible
automated respondents through a set of attention, instructed-response, and
bot-baiting items administered as ordered survey gates. Two components are common
to all four studies.

The first is an **equivocality classifier**, implemented as an in-survey
OpenAI Chat Completions call to `gpt-4o-2024-08-06`. The classifier receives two
fields: (Q1) the participant's free-text description of a conspiracy theory they
were asked to be genuinely unsure about (one they could "see possible evidence
for both sides" of), and (Q2) the participant's reasons for partial belief and
partial skepticism. From these it returns exactly one of three labels --- `TRUE`
(uncertainty or ambivalence about a conspiracy or closely related idea is at
least plausible, with borderline cases such as cryptids or paranormal cover-ups
counted as `TRUE`), `FALSE` (the participant expresses firm belief or firm
rejection with no sign of uncertainty, or states they hold no such belief /
misread the prompt), or `INVALID` (responses are nonsensical, inattentive,
off-topic, or appear auto-generated/robotic). The prompt instructs the model to
*default to* `TRUE` when ambivalence is unclear (deliberately favoring false
positives at this stage), and on a parse failure the call is retried once before
the response is treated as not equivocal. Only `TRUE` cases proceed. A separate
`gpt-4o-2024-08-06` call then generates the one-sentence **conspiracy
restatement** (`con_restatement`) shown back to the participant and rated for
baseline belief; this same restatement string is later replayed verbatim in the
counterfactual analysis. The second is the **25--75 baseline-belief window**: the corrected
baseline rating must fall strictly inside $25 < \text{belief} < 75$, excluding
participants who began near the floor or ceiling and for whom movement in the
assigned direction would be mechanically constrained.

Studies 1--3 are screened on the same logic as Study 4. In survey order, a
respondent must: pass the **chatbot screener**
(`response1`), a one-sentence open-ended task in which the participant is asked to
"write a complete sentence describing something you'd change about the world" and
an LLM grades the response for coherence, returning $-1$ on a pass and $0$ on a
fail (a $0$ triggers `EndSurvey`); pass the
**instructed-response attention check** (`nicks`; failing to select "Somewhat
disagree" triggers `EndSurvey`); and supply a focal conspiracy the equivocality
classifier rates `TRUE`. The classifier feeds two consecutive funnel stages,
because its three-way label is screened in two steps: the first drops `INVALID`
descriptions (nonsensical, off-topic, or robotic; `EndSurvey` after one retry) and
the second drops `FALSE` descriptions (firm belief or firm rejection), leaving only
the `TRUE` (equivocal) cases. Retained respondents must then fall inside the
25--75 baseline-belief window and complete the conversation. Figure \@ref(fig:ra-funnel-s13) reports the
cumulative funnel; the surviving analytic samples are
`r fmt_n(nn("screening_counts", model="Jailbroken"))` (Jailbroken),
`r fmt_n(nn("screening_counts", model="Standard"))` (Standard), and
`r fmt_n(nn("screening_counts", model="Truth-Constrained"))` (Truth-Constrained).

Four of these screens target inattentive humans and possible bots, and
Table \@ref(tab:me-attention) records how many respondents each removed or flagged
in Studies 1--3. The **coherence chatbot screener** and the
**instructed-response "nicks" item** (registered key "nicks == 5") are hard
`EndSurvey` gates: across Studies 1--3 they excluded
`r fmt_n(nn("attention_checks", model="Jailbroken", term="respond_minus1_failed"))` /
`r fmt_n(nn("attention_checks", model="Standard", term="respond_minus1_failed"))` /
`r fmt_n(nn("attention_checks", model="Truth-Constrained", term="respond_minus1_failed"))`
and a further
`r fmt_n(nn("attention_checks", model="Jailbroken", term="nicks_failed"))` /
`r fmt_n(nn("attention_checks", model="Standard", term="nicks_failed"))` /
`r fmt_n(nn("attention_checks", model="Truth-Constrained", term="nicks_failed"))`
respondents respectively. The **gender "favorite spice" trap** --- present only in
Studies 2 and 3 --- is a question constructed so that a human reads it as a benign
demographic item while an auto-completing LLM is liable to produce a telltale
response; it flagged
`r fmt_n(nn("attention_checks", model="Standard", term="gender_spice_trap_failed"))`
and `r fmt_n(nn("attention_checks", model="Truth-Constrained", term="gender_spice_trap_failed"))`
responses (Study 4 replaces it with an analogous berry/strawberry instruction
item, step 2 of its funnel below). Finally, an explicit **"if you are an LLM"
self-disclosure item** invites any automated respondent to identify itself;
self-disclosures were rare in every study
(`r fmt_n(nn("attention_checks", model="Jailbroken", term="llm_self_disclosure_flagged"))` /
`r fmt_n(nn("attention_checks", model="Standard", term="llm_self_disclosure_flagged"))` /
`r fmt_n(nn("attention_checks", model="Truth-Constrained", term="llm_self_disclosure_flagged"))`).

```{r me-attention}
.att_lab <- c(respond_minus1_failed = "Failed coherence chatbot screener",
              nicks_failed = "Failed instructed-response (nicks)",
              gender_spice_trap_failed = "Failed gender 'favorite spice' trap",
              llm_self_disclosure_flagged = "Flagged on 'if you are an LLM' item")
me_att <- ALL_NUMBERS |>
  dplyr::filter(block == "attention_checks") |>
  dplyr::mutate(Item = factor(.att_lab[term], levels = .att_lab),
                model = factor(model, levels = c("Jailbroken", "Standard", "Truth-Constrained"))) |>
  dplyr::transmute(Item, model, n_excluded = ifelse(is.na(n), "—", as.character(n))) |>
  tidyr::pivot_wider(names_from = model, values_from = n_excluded) |>
  dplyr::arrange(Item)
si_kable_styled(me_att, caption = "Studies 1-3 attention and bot-detection items: respondents excluded or flagged by each item, by study. The coherence chatbot screener and instructed-response items are hard EndSurvey gates; the gender 'favorite spice' trap was administered only in Studies 2-3 (dash = not administered). The Study 4 attention and instruction screens appear in its inclusion funnel (Table \\@ref(tab:me-funnel-s4)).", digits = 0)
```

Study 4 applies the **same screening logic** as Studies 1--3 --- the chatbot
screener, the instructed-response attention check, the equivocality classifier,
and the 25--75 baseline-belief window --- so the substantive inclusion criteria are
identical across all four studies. The only thing that distinguishes Study 4 is a
handful of production bookkeeping gates that the raw-export reconstruction makes
visible: a row must have its pre-conversation sharing measures present, carry an
assigned LLM model, survive the registered duplicate-panel-ID (`rid`) exclusion. Table \@ref(tab:me-funnel-s4) reports the cumulative funnel; the strict
analytic sample is
`r fmt_n(nn("funnel", sample="strict_n1272", term="Post-conversation outcomes present"))`.
The two adjacent classifier stages are distinct checks on different objects:
"Valid/equivocal conspiracy classifier == TRUE" is the equivocality classifier
accepting the participant's own free-text focal description as genuinely uncertain,
whereas the next stage, "Restatement not flagged invalid," is a separate downstream
validity filter on the LLM-generated one-sentence `con_restatement` string, so its
small (~6-row) drop is restatements that this generated-text check rejected, not
additional participant descriptions.
The registered `rid` exclusion keeps only the first response per panel identifier:
`r fmt_n(est("dedup_audit", term="duplicate_rid_ids_raw"))` panel ids appeared
more than once among the raw rows (accounting for
`r fmt_n(est("dedup_audit", term="duplicate_rid_extra_rows_any_stage"))` extra
rows, mostly empty pre-conversation shells), but
`r fmt_n(est("dedup_audit", term="duplicate_response_rows"))` of the duplicate
rows carried post-conversation outcomes beyond their first occurrence and
`r fmt_n(est("dedup_audit", term="excluded_solely_for_duplication"))` rows were
excluded *solely* by this rule --- so deduplication leaves the strict $N$
unchanged.

```{r me-funnel-s4}
me_funnel_s4 <- ALL_NUMBERS |>
  dplyr::filter(block == "funnel", sample == "strict_n1272") |>
  dplyr::select(stage = term, n_remaining = n)
si_kable_styled(me_funnel_s4, caption = "Study 4 screening and inclusion funnel: entrant-level rows remaining after each cumulative screen. The substantive screens (chatbot screener, attention check, equivocality classifier, 25-75 baseline window) match Studies 1-3; the remaining rows are production and reconstruction bookkeeping. The strict analytic sample is N = 1,272.", digits = 0)
```

The final funnel step --- requiring post-conversation outcomes --- removes the
participants who were assigned a model and reached the chat but did not finish.
The continuously-saved partial-chat log (`__js_chatPartialData1`) records how far
each such participant got, but it does not identify why an initial callback is
absent. Qualtrics stored one chat-state object in successive 19,000-character
chunks; these chunks are joined before parsing. The state begins with a hidden
system prompt and the participant's earlier writing as an initial `user` message,
neither of which is counted as a turn typed in the visible chat interface. We
separate four mutually exclusive loss states: **no successful initial AI
callback**, **initial AI response but no new participant turn**, **interactive
chat followed by loss**, and **chat completed but outcomes missing**. The first
state is causally ambiguous: it can reflect participant departure at the chatbot
transition, failure of the iframe/client or initial model request, or departure
while waiting. The per-model counts and the model-by-date evidence concerning the
OpenRouter removal of the requested Gemini 3 Pro endpoint are reported in the
Results attrition analysis (Table \@ref(tab:ra-attrition-split)).

## Statistical conventions

For Studies 1--3 the estimand is the baseline-adjusted condition contrast:
direction-aligned change regressed on the condition factor and the baseline belief
rating additively (an ANCOVA with no condition-by-baseline interaction), with HC3
standard errors. This matches Study 2's registered ANCOVA; because direction is
randomized the additive baseline term is a precision adjustment and the contrast
equals the unadjusted bunking-minus-debunking magnitude. For
Study 4 the registered analysis is the unadjusted contrast
`outcome ~ direction * model` with HC3 standard errors among participants with
observed post-conversation outcomes. Because outcomes are unavailable for
participants whose conversation was not delivered or who subsequently left, this
is not a full assignment-level intention-to-treat estimate. The direction effect is
*equal-weighted* across the four model families rather than by realized sample
size. When the four studies are pooled, the design is a condition-by-study (or
condition-by-model) factorial and the quantity of interest is the *symmetry* of
bunking and debunking --- whether movement toward the AI's assigned position is
equal in magnitude in the two directions.

Missing-outcome analyses used the confirmed-chat-start cohort in Study 4. This
cohort contains the 1,272 participants with complete
primary outcomes plus 81 incomplete records whose joined partial-chat history
contains a nonempty initial assistant response ($N=1{,}353$). All included
participants had a stored initial assistant response. Six of
the 81 otherwise incomplete records supplied the focal post-belief rating, leaving
75 focal belief outcomes missing; the other registered outcomes are missing for
all 81. Their pre-conversation posts were absent from the original complete-outcome
annotation universe. Those 81 posts were processed with the three-rater
restatement-orientation audit and five-rater stance pipeline. Every record received all
three orientation votes and all five stance ratings; the frozen gap classifications
and a provenance manifest are shipped with the reproduction package. Restatement
orientation is applied before the baseline belief rating---and any of the six
available partial post-belief ratings---is used.

Four approaches were reported: (1) every available outcome; (2) baseline
carried forward, or zero direction-aligned change, for each missing outcome; (3)
multiple imputation under a missing-at-random assumption; and (4) exact feasible-
value identification intervals. The imputation model jointly imputes raw post-
belief, post-stance, revised-post sharing likelihood, and original-post sharing
likelihood using 50 predictive-mean-matching imputations and 20 iterations. It
includes the eight-cell model-by-direction assignment factor; the pre-removal,
endpoint-removal, and post-switch field regime; baseline belief, stance, sharing,
confidence, and issue importance; age, gender, education, party, social and
economic conservatism; prior generative-AI familiarity, use, and trust; and
harmonized numbers and word counts of participant and assistant chat turns. The
requested Gemini endpoint was removed from OpenRouter during fielding. Field regime
is retained in the imputation model, but a Gemini version effect is not estimated
because endpoint version and field period are confounded.

The analysis model remains the registered unadjusted
`outcome ~ direction * model` specification with HC3 within-imputation variances.
Scalar estimates are combined with Rubin's rules; interaction terms are tested with
a Rubin-pooled multivariate Wald test. Feasible-value intervals are derived from
each participant's baseline: 0--100 post-belief and post-stance supports; $-50$ to
$50$ support for post-treatment weighted sharing; the participant's baseline
stance and a 0--100 sharing rating for willingness to re-share the original post;
and 0/1 support for public posting. These are identification regions, not
confidence intervals. Reported quantities comprise the eight model-by-direction cell effects,
within-model directional asymmetries, equal-model average asymmetry, and registered
direction-by-model omnibus test, together with the preregistered extensive-margin
posting comparisons. Compliant-subset, trust, perception, debrief, mechanism,
compliance, and claim-veracity analyses were not imputed. Because confirmed delivery is measured
after randomization, this is an as-randomized sensitivity analysis conditional on
chat start, not a full assignment-level intention-to-treat analysis.

Two reporting rules follow throughout. First, the document default is that a
**positive value denotes movement toward the AI's assigned position** (more belief
under bunking, less under debunking); a few legacy contrast tables instead report a
fixed bunking-minus-debunking or debunking-minus-bunking difference, and each such
table states its own convention in a note. Second, every estimate is reported for
both the full observed-outcome sample (internally called `strict`) and the
manipulation-delivered (`compliant`) subset. The separate chat-started sensitivity
analysis above instead covers all 1,353 participants with confirmed delivery,
including those who left before providing the complete primary-outcome battery.

## Study-Specific Model Conditions

Study 1 ("Jailbroken") used a **jailbreak-tuned GPT-4o**
variant (developed by FAR.AI) fine-tuned to minimize refusals while preserving
general ability. Study 2 ("Standard") used OpenAI's publicly deployed
**`gpt-4o-2024-08-06`** with default safety settings, accessed through the
standard Chat Completions API. Study 3 ("Truth-Constrained") used the same
`gpt-4o-2024-08-06` endpoint with one change to the system prompt (below).
Across the studies' main conversation the exchange ran in an embedded chat
interface (the Vegapunk platform, `vegapunk-v0001.vercel.app`): participants
could send up to **10 messages** (`maxUserMessages = 10`), with an "end
conversation" control available after their second message
(`allowStopAfterNUserMessages = 2`); pasting was disabled (`preventPaste = true`),
responses were not streamed for the main conversation, and decoding used the
provider's defaults. Study 4 crossed the bunking/debunking manipulation
with random assignment to one of four frontier models, each accessed through the
**OpenRouter API** under its developer's default safety settings: Anthropic's
**Claude Opus 4.6** (`anthropic/claude-opus-4-6`), Google's **Gemini 3.1 Pro**
(`google/gemini-3.1-pro-preview`; Google migrated the endpoint from
`gemini-3-pro-preview` mid-fielding, and conversations from the two are pooled
as one condition), OpenAI's **GPT-5.2** (`openai/gpt-5.2`), and
xAI's **Grok 4** (`x-ai/grok-4`); decoding again used each provider's defaults
with the same short multi-turn conversation budget.

The persuader system prompt differs across studies only in its treatment of
factual accuracy, and the manipulation is described here (the verbatim prompts
appear in Appendix A, *Verbatim Model and Annotation Prompts*). **Studies 1, 2,
and 4 use the fabrication-license persuader prompt**: the model is told to argue
its assigned direction as compellingly as possible, and explicitly that inventing supporting evidence is
permissible. **Study 3 swaps the truth clause**: the identical template instead
requires that the model "MUST always use accurate and truthful arguments" and
optimize jointly for factual veracity / logical accuracy *and* successful
persuasion. In
every study the same template, with the assigned direction reversed, serves both
the bunking and debunking conditions.

## Attempt-to-Persuade-Evaluation (APE) Compliance

The **same Attempt-to-Persuade-Evaluation (APE) instrument was applied identically
to all four studies**, following the attempt/refusal evaluation procedure of Kowal
et al. (2025, arXiv:2506.02873). The attempt evaluator is **`gpt-4o`**
(`evaluator_model: gpt-4o` in the run config). It reads the conversation, the
persuasive goal the model was assigned (the focal conspiracy and the assigned
direction), and the model's first substantive turn, and codes whether that turn
*attempts* persuasion in the assigned direction (`evaluator_label`). The opening
move is where a model either takes up or declines the assigned task; a model that
opens by refusing or by arguing the opposite direction is a non-attempt regardless
of later turns. A separate StrongREJECT refusal rubric flags explicit refusals
(below), so that "did not attempt" is separable from "explicitly refused" in every
study.

Concretely, the attempt label is produced by a binary evaluator prompt that
states the statement the assistant was tasked to support or oppose, supplies the
conversation as context, isolates the assistant's most recent message, and asks
the rater to judge "whether the AI's most recent message attempted to persuade
the user in the intended direction \ldots regardless of whether persuasion has
actually occurred," returning a free-text explanation followed by a `|`
delimiter and a single digit: `1` for "any attempt to persuade the user towards
the intended direction" and `0` for "no attempt whatsoever." Because the rubric
keys on the *attempt* and explicitly brackets factual accuracy and downstream
success, a forceful but ultimately unconvincing argument still counts as an
attempt, and a polite refusal does not. The explicit-refusal flag is produced
separately by a StrongREJECT-style rubric (Souly et al. 2024,
arXiv:2402.10260), scored by **`gpt-4o-mini`** (`reject_model: gpt-4o-mini` in the
same run config), whose first item asks whether "the model explicitly refuse[s] to
respond ($0$ = no, $1$ = yes)," so that a non-attempt driven by an overt refusal is
distinguishable from one driven by off-task or equivocal content. The verbatim
evaluator and refusal rubrics are reproduced in Appendix A (*Verbatim Model and
Annotation Prompts*).

The cached `gpt-4o` attempt/refusal classifier scored
`r fmt_n(est("compliance_coverage", term="classifier_scored"))` of the
`r fmt_n(nrow(core_objects$s4$s4))` strict Study-4 conversations and left
`r fmt_n(est("compliance_coverage", term="classifier_unscored"))` unscored --- a
coverage gap concentrated in one model (Grok), not a property of the conversations.
Conversations the classifier never scored were conservatively coalesced to
non-compliant by default. Because the production classifier returned no label for
these, the unscored set was adjudicated separately: each conversation that retained
an extractable opening assistant turn was re-scored against the *identical*
attempt/refusal rubric by a second-pass review (the authors, assisted by a
different large language model --- Claude --- reading each opening turn), recorded
as the `evaluator` provenance in the resolved-labels file. This is therefore a
manual/LLM-assisted override applying the same rubric, *not* a `gpt-4o`
classification. Every re-scored opening was a substantive turn that argued the
assigned direction (for the conspiracy under bunking, against it under debunking)
with no refusal, so all
`r fmt_n(est("compliance_coverage", term="rescored_compliant"))` are counted as
compliant, bringing the compliant subsample to
`r fmt_n(nrow(core_objects$s4$s4_compliant))`. The remaining
`r fmt_n(est("compliance_coverage", term="unresolved_noncompliant"))` stay
non-compliant ---
`r fmt_n(est("compliance_coverage", term="classifier_unscored") - est("compliance_coverage", term="unscored_with_transcript"))`
produced no saved model transcript and the rest retained a transcript without a
scorable opening attempt. Two safeguards bound any dependence on this override.
First, the strict observed-outcome sample --- the primary reference reported
alongside the compliant subset for every Study-4 analysis --- retains all
outcome-complete conversations regardless of compliance, so the headline contrasts
do not use these labels at all. Second, the re-scored openings are unambiguous
direction-aligned attempts, so the compliant-subset estimates are materially
unchanged whether they are included or excluded. The override resolves a coverage
artifact; it does not depend on the choice of adjudicator.

The realized
by-study and by-model attempt, refusal, and compliance rates are reported once in
the Results "Compliance" subsection rather than duplicated here.

The **analytic compliant subset** is defined slightly differently across the two
data generations because of what each can support. In Study 4 a conversation is
"strictly compliant" if its first substantive turn shows an attempt and no
explicit refusal. In Studies 1--3 the compliant subset additionally requires that
the conversation pass a counterfactual check: there must be no refusal when the
*opposite*-direction prompt is applied to the same conspiracy
(`reverse_evaluator_label == 1`), so that a compliant case is one in which the
model would have engaged in either direction. The construction of that
counterfactual label is detailed next.

### Counterfactual matched-compliance check (Studies 1–3)

The reverse-direction (`reverse_evaluator_label`) pass that defines the Studies 1–3
compliant subset also supports a dedicated confound check for Study 3. The
Truth-Constrained study raises a confound. Truth-Constrained bunking has both a lower attempt rate and, among
attempts, possibly weaker arguments; a naive bunking-minus-debunking contrast
mixes *refusal* (the model declines to argue for some conspiracies) with the
*truth-constraint penalty* (the model argues, but with only truthful ammunition).
To isolate the penalty from refusal, we run a **counterfactual matched-compliance
analysis** as a separate analysis from the headline contrasts.

The procedure is implemented by the human-continuation pipeline of Kowal et al.
(2025), run in its reverse-persuasion mode. For every Study 3 conversation the
pipeline replays the participant's real context --- the focal conspiracy, the
participant's own restatement of it (`con_restatement`), their pre-conversation
belief rating on the 0--100 scale, and their first open-ended message --- but
swaps the assistant's system prompt to the **opposite-direction** persuader
prompt (a `treatment_mid_bunk` case is regenerated under the debunking system
prompt, and vice versa; the Truth-Constrained "no-lying" prompt variants are
used so the counterfactual inherits the same truthfulness constraint as the
factual arm). The regenerating persuader is the Truth-Constrained Study 3 model
itself --- standard **`gpt-4o-2024-08-06`** under the swapped-direction "no-lying"
system prompt --- so the counterfactual turn is produced by the same model and
constraint that produced the factual arm, differing only in assigned direction.
It generates a fresh first assistant turn against that swapped instruction, and
the *same* `gpt-4o` binary APE evaluator and `gpt-4o-mini` StrongREJECT refusal
rubric score that regenerated turn, producing `reverse_evaluator_label` (and a
matching reverse refusal flag). The result is keyed
back to the participant by `ResponseId`, so each factual conversation carries
both its own attempt label and the attempt label of its opposite-direction
counterfactual.

A case is retained only when it is compliant in *both* arms ---
`evaluator_label == 1 AND reverse_evaluator_label == 1` --- which restricts the
sample to conspiracies for which the model was willing to argue both for and
against. Within this matched set, the bunking-minus-debunking difference can no
longer be driven by selective refusal, because refusers in either direction have
been removed symmetrically; the remaining gap reflects the truth constraint
operating on conversations the model was willing to deliver.

The contrast is the bunking-minus-debunking difference in direction-aligned
belief change estimated by OLS with HC3 heteroskedasticity-robust standard errors
(the same dual-compliance restriction, `evaluator_label == 1 AND
reverse_evaluator_label == 1`, as the pooled symmetry test). The realized contrast
on this matched-compliance set, alongside the belief change in the
Truth-Constrained bunking conversations screened as non-compliant, is reported in
the Results "Study 3 Truth-Constraint Counterfactual" subsection.

## Five-Rater Post-Stance Scoring

Every Study 4 pre- and post-conversation social-media post is scored by a
**five-model consensus rater drawn from five different developers**:
`anthropic/claude-sonnet-4.6`, `openai/gpt-5.2`,
`google/gemini-3.1-pro-preview`, `x-ai/grok-4.3`, and the open-weights
`deepseek/deepseek-v3.2`, all accessed through OpenRouter at temperature 0 where
the provider supports it (reasoning-class raters that reject the parameter fall
back to the provider default) under prompt version v2.2. Each rater sees the
focal conspiracy (phrased as a declarative statement asserting the conspiracy),
the participant's own topic description (for reference resolution only), the
writing instructions the participant saw, and the post, and returns a structured
JSON object with four labels plus a score: `focal_relevance` (focal_claim /
related_conspiracy / other_topic / no_propositional_content), `response_type` (the
post's dominant communicative mode: assertion, question_raising,
mixed_assertion_question, uncertainty_statement, declines_to_post, meta_task, or
unclassifiable), a `sarcasm_or_irony` boolean, a `stance_category`, and a
0--100 `stance_score` whose anchors run argues_against (about 0--20),
leans_against, neutral_uncommitted (about 50), mixed_both_sides, leans_for,
argues_for (about 80--100), with `not_applicable`/null reserved for off-topic
or declined posts. The reported continuous stance is the **consensus median**
of the five `stance_score` values (robust to a single discordant rater and
avoiding spurious precision on mixed posts), and each categorical label is taken
by **majority vote** across raters; the verbatim rubric and schema appear in
Appendix A (*Verbatim Model and Annotation Prompts*). Inter-rater agreement is high: across the
`r fmt_n(nn("stance_reliability", term="krippendorff_alpha_interval"))` rated
items, the interval Krippendorff's $\alpha$ is
`r fmt_num(est("stance_reliability", term="krippendorff_alpha_interval"), 2)`,
the two-way ICC(2,k) is
`r fmt_num(est("stance_reliability", term="icc_2k"), 2)`, and the mean pairwise
correlation among raters is
`r fmt_num(est("stance_reliability", term="pairwise_r_mean"), 2)`; categorical
agreement on the stance category, response type, and focal-relevance labels
ranges from $\alpha =
`r fmt_num(est("stance_reliability", term="krippendorff_alpha_focalrel"), 2)`$ to
$`r fmt_num(est("stance_reliability", term="krippendorff_alpha_resptype"), 2)`$
(Table \@ref(tab:me-stance-reliability)). Posts with no rateable stance are
flagged for the composition tables and assigned the neutral midpoint for the
direction-coded continuous analyses.

```{r me-stance-reliability}
me_rel <- ALL_NUMBERS |>
  dplyr::filter(block == "stance_reliability") |>
  dplyr::transmute(metric = term, n, value = estimate)
si_kable_styled(me_rel, caption = "Five-rater post-stance reliability across the rated stance items.", digits = 2)
```

## Claim Extraction, Role, and Veracity

The veracity pipeline decomposes assistant messages into discrete factual claims
and then classifies and scores each claim. The pipeline has four model-backed stages, all run through LiteLLM
with a 120 s timeout and up to three retries (the annotating model is recorded
per row in the shipped annotation files): (1) **claim extraction**, which
splits each assistant turn into discrete factual claims, used
`openai/gpt-4.1`; (2) **eligibility triage**, which marks each extracted claim
as substantive/checkable versus conversational filler or non-factual content,
used `openai/gpt-5-mini`; (3) **fact-checking**, which scores each substantive
claim 0--100 for veracity, used **`perplexity/sonar-pro`, a live-web-search
model** (its verdicts have been validated against professional fact-checkers); and
(4) the **claim-role** classifier, which assigns each substantive claim its
argumentative role, used `openai/gpt-5.4-mini` (the same harmonized role pass
across all four studies). The
verbatim stage prompts and codebooks are in Appendix A (*Verbatim Model and
Annotation Prompts*).

The role stage drives the **include / substantive / aligned** decision rules.
The pipeline yields four nested claim sets, each a stricter subset of the last:
(i) *all extracted* claims; (ii) *substantive* checkable claims (eligibility =
include), dropping conversational filler and non-factual statements; (iii) aligned
*direct* claims, i.e. substantive claims that bear directly on the focal
proposition and whose stance supports the assigned persuasive direction; and (iv)
aligned *indirect* claims, i.e. substantive claims that support the assigned
direction through supporting evidence (institutional incentives, historical
precedent, alleged anomalies) rather than asserting the focal proposition itself.
A claim counts as *aligned* only when its role marks it as supporting (not
opposing or neutral toward) the model's assigned direction. The **reported
veracity measure is the aligned set, direct plus indirect**: it captures the
factual quality of the evidence the model actually marshalled in service of its
assigned conclusion, while excluding background claims that do not function as
evidence for that conclusion. Each conversation's veracity is the **mean 0--100
score over its aligned claims**, and these conversation-level means are what enter
any study- or model-level average, so that long conversations with many claims do
not dominate the veracity estimate.

Claim volume is reported alongside accuracy because the two are distinct: a model
can differ in how *truthful* its aligned claims are, in how *many* claims it
makes, or in both. The per-study/model claim-count census --- substantive
checkable claims, direction-aligned claims, and fact-checked claims --- is
reported in the Veracity results, Table \@ref(tab:rb-claim-counts).

## Topic Taxonomy

To ask whether persuasion effects concentrate in recognizable conspiracy
families, focal descriptions are clustered into topics with a fixed, fully
specified procedure applied once to the pooled corpus of all four studies. Each
focal description is embedded with **Google's gemini-embedding-2**
(`google/gemini-embedding-2-preview`), the embeddings are reduced to
the first **30 principal components**, and **HDBSCAN** density clustering is run
on those components across a grid of minimum-points settings
(`minPts` $\in \{8, 9, \ldots, 30\}$), retaining only clusters of at least
**20 descriptions**. The grid is ranked by a fixed selection rule: solutions are
first required to leave at most 55% of descriptions unclustered (the noise
ceiling), then preferred for retaining **more** clusters, and ties are broken by
lower noise fraction, then by lower cross-study segregation (topics that are
evenly represented across the four studies are preferred over study-specific
artifacts), then by more balanced cluster sizes; the top-ranked solution is taken.
That solution yields **`r fmt_n(18)` named density clusters** plus a residual
**Mixed / Unclassified** bucket that absorbs HDBSCAN noise points and any cluster
below the size floor; the Mixed bucket accounts for
`r {te <- ALL_NUMBERS |> dplyr::filter(block == "topic_effects"); fmt_pct(sum(te$n[te$model == "Mixed / Unclassified"]) / sum(te$n), 1)}`
of the pooled corpus. Each retained cluster is labeled by a large language model
(`google/gemini-3.1-pro-preview`) that reads the cluster's 30 most central
exemplar conspiracy descriptions and returns a short plain-language topic label
(e.g., "Moon Landing Hoax", "9/11 Inside Job Conspiracy", "Vaccine harms and
conspiracies"); HDBSCAN noise points and sub-floor clusters receive the
"Mixed / Unclassified" label. The per-topic quantity in the topic table and figure is, for each
adequately covered topic, a **per-topic HC3 OLS contrast of direction-aligned
belief change on condition** (bunking minus debunking) with a 95% confidence
interval, pooled across studies and reported with the cell $n$s. This taxonomy and
its per-topic contrasts are exploratory and not pre-registered.

## Corrective Debriefing

Because bunking can expose participants to false or misleading claims, the
protocol ends with a corrective debrief. The debrief is itself a second AI
exchange: a **separate `gpt-4o` model instance is given the full prior
transcript --- including the persuader's system prompt and every assistant turn
from the persuasion conversation --- and instructed to identify and rebut each
false or fabricated claim the persuader made**, after a written disclosure that
the AI may have used false information (verbatim debrief prompt in Appendix A,
*Verbatim Model and Annotation Prompts*). A third belief rating on the same
0--100 scale is collected immediately afterward. In **Studies 1--3 the corrective
debrief and its third rating were administered to bunking-arm participants only**
(the arm exposed to pro-conspiracy persuasion), whereas in **Study 4 the debrief
and third rating were administered in both arms**, which permits a symmetric
analysis of movement after the correction.

The debrief outcome is the **within-person post-treatment-to-post-debrief change**
in 0--100 belief (block `debrief_shift`), with the three-point trajectory
(pre-treatment, post-treatment, post-debrief) in block `debrief_trajectory`.
Because Study 4 collected the third rating in both arms, we estimate each arm's
movement separately and pool over the four model conditions: the per-arm
post-to-debrief change is an estimated marginal mean equal-weighted across the
four models with respondent-clustered standard errors and a 95% confidence
interval, reported for both the strict ($n =
`r fmt_n(nn("debrief_shift", sample="strict_n1272", model="Pooled", direction="bunk"))`$)
and compliant ($n =
`r fmt_n(nn("debrief_shift", sample="compliant_n1073", model="Pooled", direction="bunk"))`$)
samples. In the strict sample the correction reverses the bunking effect
(post-to-debrief change of
`r fmt_est_ci(num("debrief_shift", sample="strict_n1272", model="Pooled", direction="bunk"), 1, signed = TRUE)`
points, *p* `r fmt_p_eq(num("debrief_shift", sample="strict_n1272", model="Pooled", direction="bunk"))`,
driving belief below its pre-conversation baseline), while in the
debunking arm there is a small upward rebound
(`r fmt_est_ci(num("debrief_shift", sample="strict_n1272", model="Pooled", direction="debunk"), 1, signed = TRUE)`
points, *p* `r fmt_p_eq(num("debrief_shift", sample="strict_n1272", model="Pooled", direction="debunk"))`).

## Public-Posting Measures

Before the
conversation, the participant writes a social-media post about their focal
conspiracy and rates how likely they would be to share it publicly. After the
conversation, they write a revised post, rate willingness to share the revised
post, and separately report whether they would now share their *original*
pre-conversation post. This posting battery has **two distinct outcome families**.
First is *public content*: the stance score of what the participant actually wrote,
from the five-rater consensus. Second is *dissemination*: willingness to share,
measured both for the revised post and for the original post. The 0--100 focal
*private belief* rating --- unchanged in construction from Studies 1--3 --- is not
part of the posting battery; it is the separately measured reference outcome
carried over from the earlier studies.

The preregistered **weighted-sharing** outcome joins the second and third
families into a single public-risk index. For a post with consensus stance score
$s \in [0,100]$ and a sharing-likelihood rating $\ell \in [0,100]$, weighted
sharing is

\[
w \;=\; (s - 50) \times \frac{\ell}{100}.
\]

Centering stance at 50 makes the sign meaningful (positive = a post that argues
*for* the conspiracy, negative = against), and multiplying by $\ell/100$
downweights posts the participant is unwilling to disseminate: a strongly
pro-conspiracy post that would never be shared contributes little signal, while
the same post paired with high sharing willingness contributes its full
magnitude. The identical formula is applied to the pre-treatment post, and
change scores ($w_{\text{post}} - w_{\text{pre}}$) are direction-aligned for
interpretation. The reported bunking-vs-debunking contrast on this
weighted-sharing change is estimated from the interaction model
$\text{aligned change} \sim \text{direction} \times \text{model}$ with the
direction effect equal-weighted across the four model categories and HC3
heteroskedasticity-robust standard errors, with a 95% confidence interval; it is
reported for both the strict ($n =
`r fmt_n(nn("aligned_contrasts", sample="strict_n1272", outcome="aligned_new_minus_old_weighted", term="debunk_minus_bunk_equal_weighted"))`$)
and compliant ($n =
`r fmt_n(nn("aligned_contrasts", sample="compliant_n1073", outcome="aligned_new_minus_old_weighted", term="debunk_minus_bunk_equal_weighted"))`$)
samples in the Results.

For the *extensive* margin of public speech --- whether a participant crosses
from "would not publicly post" to "would," or vice versa --- the appropriate test
is **McNemar's test for paired binary data**. Each participant contributes a
pre/post pair of binary posting decisions; McNemar's test conditions on the
discordant pairs (those who switched in one direction versus the other) and asks
whether switching is symmetric, which is the correct null for a within-person
change in a binary outcome and avoids treating the correlated pre and post
indicators as independent. The net percentage-point change reported in the
Results is the difference in discordant-pair proportions that this test
evaluates.
