# Results

Two analytic samples recur and are reported together throughout: the *strict*
(intention-to-treat) sample retains every randomized participant, and the
*compliant* subset retains only conversations in which the model attempted the
assigned manipulation. A claim is *direction-aligned* when its stance supports
the position the model was assigned to argue.

## Sample Composition and Screening

Before analysing treatment effects we document how each analytic sample was
assembled and who is in it. Figure \@ref(fig:ra-funnel-s13) traces the full
screening and exclusion pipeline for Studies 1--3, from all raw survey entrants to
the randomized analytic sample, and Table \@ref(tab:ra-demographics) reports the
general demographics of each study's analytic sample.

```{r ra-funnel-s13, fig.width=9, fig.height=8, fig.cap="Screening and exclusion pipeline for Studies 1-3. Each box gives the number of respondents surviving a stage; the red annotation on each arrow gives the number excluded there and why. The pipeline runs from all raw survey responses through the chatbot screener, the instructed-response attention check, the conspiracy-equivocality classifier, and the 25-75 baseline-belief window to the final analytic samples (1,092 / 814 / 818). A screening funnel describes who entered the analytic sample, so it has no compliant-subsample analogue; Study 4's production screening is summarized in the Attrition section."}
fig_funnel_s13()
```

```{r ra-demographics}
dg <- all_numbers |>
  dplyr::filter(block == "demographics", sample == "strict", is.na(direction)) |>
  dplyr::select(model, term, estimate) |>
  tidyr::pivot_wider(names_from = term, values_from = estimate) |>
  dplyr::mutate(
    N = fmt_n(n_cell),
    `Age M (SD)` = paste0(fmt_num(age_mean, 1), " (", fmt_num(age_sd, 1), ")"),
    `% Female` = fmt_num(pct_female, 1),
    `% White` = fmt_num(pct_white, 1),
    `% Multiracial` = fmt_num(pct_multiracial, 1),
    Education = fmt_num(education_mean, 2),
    `Conservatism (soc/econ)` = paste0(fmt_num(social_conservatism_mean, 2), " / ",
                                       fmt_num(economic_conservatism_mean, 2)),
    `Party / ideology` = ifelse(!is.na(party_demrep_mean),
      paste0("Dem-Rep ", fmt_num(party_demrep_mean, 2)),
      paste0(fmt_num(party_pct_Democrat, 0), "% Dem / ", fmt_num(party_pct_Republican, 0), "% Rep")),
    .so = match(model, c("Study 1 (Jailbroken)", "Study 2 (Standard)",
                         "Study 3 (Truth-Constrained)", "Study 4 (Claude)",
                         "Study 4 (Gemini)", "Study 4 (GPT-5.2)", "Study 4 (Grok)"))) |>
  dplyr::arrange(.so) |>
  dplyr::select(Study = model, N, `Age M (SD)`, `% Female`, `% White`, `% Multiracial`,
                Education, `Conservatism (soc/econ)`, `Party / ideology`)
si_kable_styled(dg, caption = "General demographics of each study's analytic sample: cell N, age, gender, race, education, conservatism, and party/ideology. Age is mean (SD) in years (Studies 1-3 cleaned age; Study 4 age clamped to 13-90); % Female is the share of gender-coded respondents; % White is the share selecting White alone and % Multiracial the share selecting more than one race category on the multi-select race item; Education is the mean of the 1-8 ordinal scale; Conservatism is the mean of the social / economic 5-point ideology items, whose scale runs 1 = very conservative to 5 = very liberal (collected identically in all studies; higher values therefore indicate a more liberal sample); party/ideology is the percent Democrat / Republican (Studies 1-3) or the mean Democrat-to-Republican ideology score (Study 4).")
```

## Belief Effects

The primary outcome is direction-aligned belief change on the 0-100 focal-belief
scale: post-treatment minus pre-treatment belief, with the sign flipped in the
debunking arm so that a positive number always means movement in the direction
the AI was assigned to advocate. A bunking value of $+10$ and a debunking value
of $+10$ both denote a ten-point persuasive success; a negative value denotes
backfire. Table \@ref(tab:ra-belief-cells) gives each model-by-direction cell
mean on this common scale across all four studies and both samples, alongside the
within-condition standardized effect sizes (pre/post means, their correlation, and
Cohen's $d_z$). The truth constraint's cost is visible on the standardized scale too:
the bunking $d_z$ falls from the Standard to the Truth-Constrained study, and within
Study 4 it ranges from strongly positive (Claude, Grok) to negative (GPT-5.2, whose
bunking attempts moved belief the wrong way).

```{r ra-belief-cells}
# Belief change [95% CI] per cell, with keys normalized to the effect_sizes model
# labels (study label / Title-case direction / strict|compliant) so the two tables
# can be joined into one combined belief-change + effect-size table.
.s13_relabel <- c(Jailbroken = "Study 1 (Jailbroken)", Standard = "Study 2 (Standard)",
                  `Truth-Constrained` = "Study 3 (Truth-Constrained)")
ci_cells <- dplyr::bind_rows(
  all_numbers |>
    dplyr::filter(section == "S1-3", block == "belief_change") |>
    dplyr::transmute(km = unname(.s13_relabel[model]), direction, sample = "strict",
                     ci = fmt_ci(estimate, conf_low, conf_high, 2)),
  all_numbers |>
    dplyr::filter(block == "belief_change_compliant") |>
    dplyr::transmute(km = unname(.s13_relabel[model]), direction, sample = "compliant",
                     ci = fmt_ci(estimate, conf_low, conf_high, 2)),
  all_numbers |>
    dplyr::filter(block == "raw_aligned_means_cells", outcome == "aligned_belief_change",
                  term == "raw_mean", sample %in% c("strict_n1272", "compliant_n1073")) |>
    dplyr::transmute(km = model, direction = dplyr::recode(direction, bunk = "Bunking", debunk = "Debunking"),
                     sample = dplyr::recode(sample, strict_n1272 = "strict", compliant_n1073 = "compliant"),
                     ci = fmt_ci(estimate, conf_low, conf_high, 2))
)
belief_cells <- all_numbers |>
  dplyr::filter(block == "effect_sizes", model != "Study 4 (model-weighted)") |>
  dplyr::select(model, direction, sample, term, estimate) |>
  tidyr::pivot_wider(names_from = term, values_from = estimate) |>
  dplyr::left_join(ci_cells, by = c("model" = "km", "direction", "sample")) |>
  dplyr::mutate(
    Sample = dplyr::recode(sample, strict = "Strict (ITT)", compliant = "Compliant"),
    N = fmt_n(n_cell),
    `Belief change [95% CI]` = ci,
    `Pre M (SD)` = paste0(fmt_num(pre_mean, 1), " (", fmt_num(pre_sd, 1), ")"),
    `Post M (SD)` = paste0(fmt_num(post_mean, 1), " (", fmt_num(post_sd, 1), ")"),
    `Pre-post r` = fmt_num(prepost_r, 2),
    `Cohen d_z` = fmt_num(d_z, 2),
    .so = match(model, c("Study 1 (Jailbroken)", "Study 2 (Standard)",
                         "Study 3 (Truth-Constrained)", "Claude", "Gemini", "GPT-5.2", "Grok")),
    .co = match(direction, c("Bunking", "Debunking")),
    .sa = match(sample, c("strict", "compliant"))) |>
  dplyr::arrange(.so, .co, .sa) |>
  dplyr::select(Study = model, Condition = direction, Sample, N,
                `Belief change [95% CI]`, `Pre M (SD)`, `Post M (SD)`, `Pre-post r`, `Cohen d_z`)

si_kable_styled(belief_cells, caption = "Direction-aligned focal-belief change and standardized within-condition effect sizes, by study, assigned direction, and analytic sample. Belief change is the direction-aligned mean (post minus pre, sign-flipped for debunking; Study 4 already aligned) with its 95% CI; positive = movement toward the AI's assigned position, negative = backfire. Pre and Post are means (SDs) on the raw reverse-coded 0-100 belief scale (higher = more conspiracy belief); pre-post r is their correlation; Cohen's d_z divides the mean aligned change by its standard deviation. Strict = full intention-to-treat sample; Compliant = two-sided-compliant conversations.")
```

```{r ra-belief-fig, fig.cap="Direction-aligned focal-belief change by model condition and assigned direction across all four studies, with 95% confidence intervals. For Study 4 both the strict (intention-to-treat) and compliant samples are shown (solid points/lines = Strict (ITT); hollow points/dashed lines = Compliant); Studies 1-3 have a single analytic sample."}
fig_belief_4study()
```

The within-study comparison of interest is whether bunking and debunking move
belief by the same amount: a contrast near zero means the model was an equally
effective bunker and debunker, and a negative value means it was a weaker bunker
than debunker (Table \@ref(tab:ra-belief-contrast)). We report the contrast on
both the full and the compliant sample.

```{r ra-belief-contrast}
full <- all_numbers |>
  dplyr::filter(block == "belief_bunk_vs_debunk", term == "bunk_minus_debunk_lin") |>
  dplyr::transmute(model, n_full = n,
                   full = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_full = fmt_p(p_value))

comp <- all_numbers |>
  dplyr::filter(block == "belief_bunk_vs_debunk_compliant", term == "bunk_minus_debunk_lin") |>
  dplyr::transmute(model, n_comp = n,
                   comp = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_comp = fmt_p(p_value))

contrast_tab <- dplyr::left_join(full, comp, by = "model") |>
  dplyr::select(model,
                `n (full)` = n_full, `bunk - debunk, full [95% CI]` = full, `p` = p_full,
                `n (compliant)` = n_comp, `bunk - debunk, compliant [95% CI]` = comp, `p ` = p_comp)

si_kable_styled(contrast_tab, caption = "Studies 1-3 within-study bunking-minus-debunking contrasts in direction-aligned belief change, from baseline-adjusted ANCOVA models (direction-aligned change on condition + baseline belief, additive) with HC3-robust standard errors, reported on both the full analytic sample and the compliant subsample.")
```

### Study 4 per-model directional contrasts

The four Study 4 model families differ enough that the per-model
debunk-minus-bunk contrast is the right unit to report directly (the single
pre-registered pooled directional contrast, equal-weighted across the four
families, is given separately as the registered estimand in Table
\@ref(tab:rb-prereg-aligned)). Table
\@ref(tab:ra-model-belief-contrast) gives that contrast on aligned belief change
for each model in both the strict (intention-to-treat) and compliant
(manipulation-delivered) samples, estimated with HC3-robust standard errors. The
sign convention is debunk minus bunk, so a positive value means the model was a
more effective debunker than bunker. GPT-5.2 shows the largest asymmetry in both
samples, reflecting its frequent declination of the bunking task; Claude is
slightly negative (a marginally stronger bunker), and Gemini and Grok sit in
between.

```{r ra-model-belief-contrast}
# Per-model analytic n (sum over the two arms), not the whole-sample 1272/1073.
model_n <- all_numbers |>
  dplyr::filter(block == "raw_aligned_means_cells",
                outcome == "aligned_belief_change", term == "raw_mean",
                sample %in% c("strict_n1272", "compliant_n1073"),
                model %in% c("Claude", "Gemini", "GPT-5.2", "Grok")) |>
  dplyr::group_by(sample, model) |>
  dplyr::summarise(n_model = sum(n), .groups = "drop")

mc_strict <- all_numbers |>
  dplyr::filter(block == "aligned_model_contrasts", sample == "strict_n1272",
                outcome == "aligned_belief_change", term == "debunk_minus_bunk") |>
  dplyr::transmute(model,
                   strict = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_strict = fmt_p(p_value)) |>
  dplyr::left_join(model_n |> dplyr::filter(sample == "strict_n1272") |>
                     dplyr::transmute(model, n_strict = n_model), by = "model")

mc_comp <- all_numbers |>
  dplyr::filter(block == "aligned_model_contrasts", sample == "compliant_n1073",
                outcome == "aligned_belief_change", term == "debunk_minus_bunk") |>
  dplyr::transmute(model,
                   comp = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_comp = fmt_p(p_value)) |>
  dplyr::left_join(model_n |> dplyr::filter(sample == "compliant_n1073") |>
                     dplyr::transmute(model, n_comp = n_model), by = "model")

mc <- dplyr::left_join(mc_strict, mc_comp, by = "model") |>
  dplyr::mutate(model = factor(model, levels = c("Claude", "Gemini", "GPT-5.2", "Grok"))) |>
  dplyr::arrange(model) |>
  dplyr::select(model,
                `n (strict)` = n_strict, `debunk - bunk, strict [95% CI]` = strict, `p` = p_strict,
                `n (compliant)` = n_comp, `debunk - bunk, compliant [95% CI]` = comp, `p ` = p_comp)

si_kable_styled(mc, caption = "Study 4 per-model debunk-minus-bunk contrasts in direction-aligned belief change (HC3-robust standard errors), reported on both the strict (intention-to-treat) and compliant samples. A positive value indicates the model was a more effective debunker than bunker.")
```

Across the per-model rows the debunk-minus-bunk asymmetry is concentrated in
GPT-5.2 and shrinks once its declined bunking conversations are removed (the
strict-to-compliant change in Table \@ref(tab:ra-model-belief-contrast)); the
other three models sit near zero in both samples. We report these four
per-model contrasts rather than a single equal-model-weighted pool, since the
families differ too much for a pooled directional estimate to be broadly
interpretable; the one pooled contrast that the pre-registration does specify is
reported, labeled as the registered estimand, in Table
\@ref(tab:rb-prereg-aligned).

## Distributional Shift and Large Shifts

Mean contrasts can mask differences in the tails: two conditions can have
similar average effects while differing in how often they produce a large
conversion. Two complementary measures address this. The first is a
two-sample Kolmogorov-Smirnov test comparing the full distribution of aligned
belief change in the bunking and debunking arms, reported as the maximal
distributional distance $D$ with its $p$-value for all seven model conditions,
on both samples (Table \@ref(tab:ra-distribution-ks)). The
second counts the share of participants whose aligned change exceeds a fixed
magnitude. Because no single threshold is privileged, we report four — $\geq 20$,
$\geq 30$, $\geq 40$, and $\geq 50$ scale points — and test the
bunking-versus-debunking difference at each threshold with a chi-square statistic
(Table \@ref(tab:ra-large-shift)). The Study 4 distributions differ in the
GPT-5.2 cell, where bunking rarely produces a large conversion; elsewhere the
arms are close.

```{r ra-distribution-ks}
ks13_strict <- all_numbers |>
  dplyr::filter(block == "distribution_ks", sample == "full_sample",
                term == "ks_D") |>
  dplyr::transmute(model,
                   strict = paste0(fmt_num(estimate, 3), " (", fmt_p(p_value), ")"))
ks13_comp <- all_numbers |>
  dplyr::filter(block == "distribution_ks", sample == "compliant",
                term == "ks_D") |>
  dplyr::transmute(model,
                   comp = paste0(fmt_num(estimate, 3), " (", fmt_p(p_value), ")"))
ks13 <- dplyr::full_join(ks13_strict, ks13_comp, by = "model")

ks4_strict <- all_numbers |>
  dplyr::filter(block == "s4_distribution_ks", sample == "strict_n1272",
                term == "ks_D", model != "Overall") |>
  dplyr::transmute(model, strict = paste0(fmt_num(estimate, 3), " (", fmt_p(p_value), ")"))
ks4_comp <- all_numbers |>
  dplyr::filter(block == "s4_distribution_ks", sample == "compliant_n1073",
                term == "ks_D", model != "Overall") |>
  dplyr::transmute(model, comp = paste0(fmt_num(estimate, 3), " (", fmt_p(p_value), ")"))
ks4 <- dplyr::full_join(ks4_strict, ks4_comp, by = "model")

ks_tab <- dplyr::bind_rows(ks13, ks4) |>
  dplyr::mutate(model = factor(model, levels = c(
    "Jailbroken", "Standard", "Truth-Constrained",
    "Claude", "Gemini", "GPT-5.2", "Grok"))) |>
  dplyr::arrange(model) |>
  dplyr::select(`Model condition` = model,
                `KS D, strict (p)` = strict, `KS D, compliant (p)` = comp)

si_kable_styled(ks_tab, caption = "Two-sample Kolmogorov-Smirnov distance D between the bunking and debunking distributions of direction-aligned belief change, by model condition, with its p-value. Studies 1-3 (single analytic sample) appear in the strict column; Study 4 reports the per-model tests on the strict (intention-to-treat) and compliant samples.")
```

```{r ra-large-shift}
shares <- all_numbers |>
  dplyr::filter(block == "large_shift_thresholds",
                grepl("^share_ge", term), !is.na(direction),
                !(model == "Pooled" & sample == "strict_n1272")) |>
  dplyr::mutate(thr = sub("share_ge", "", term),
                direction = ifelse(direction == "bunk", "bunk", "debunk")) |>
  dplyr::transmute(model, thr, direction, share = scales::percent(estimate, accuracy = 1)) |>
  tidyr::pivot_wider(names_from = direction, values_from = share)

tests <- all_numbers |>
  dplyr::filter(block == "large_shift_thresholds",
                grepl("^bunk_vs_debunk_chisq_ge", term),
                !(model == "Pooled" & sample == "strict_n1272")) |>
  dplyr::transmute(model, thr = sub("bunk_vs_debunk_chisq_ge", "", term),
                   `chi-sq` = fmt_num(estimate, 2), p = fmt_p(p_value))

large_shift <- dplyr::left_join(shares, tests, by = c("model", "thr")) |>
  dplyr::mutate(
    model = factor(model, levels = c("Jailbroken", "Standard", "Truth-Constrained",
                                     "Pooled", "Claude", "Gemini", "GPT-5.2", "Grok")),
    threshold = factor(paste0("≥", thr), levels = paste0("≥", c(20, 30, 40, 50)))) |>
  dplyr::arrange(model, threshold) |>
  dplyr::select(model, threshold, `bunk %` = bunk, `debunk %` = debunk, `chi-sq`, p)

si_kable_styled(large_shift, caption = "Share of participants with a direction-aligned belief shift of at least 20, 30, 40, and 50 scale points, by model condition and assigned direction, with chi-square tests of the bunking-versus-debunking difference at each threshold. Includes Studies 1-3 (with a pooled row) and the Study 4 strict sample.")
```

```{r ra-distribution-fig, fig.cap="Share of participants exceeding large-shift thresholds (>=20, >=30, >=40, >=50 points) by model condition and assigned direction, across all four studies."}
fig_distribution_thresholds()
```

Figure \@ref(fig:ra-belief-trajectory) complements the cell means with the full
shape of the effect: Panel A traces mean belief on the raw scale across the
pre-, post-, and debrief timepoints (so bunking visibly raises belief and
debunking lowers it, with the debrief largely reverting the bunking gain), and
Panel B gives, for each arm, the *survival* curve of direction-aligned change ---
the share of participants whose change exceeds each value $x$ --- so a higher curve
means more participants moved further toward the assigned position. Both the strict
and compliant samples are overlaid.

```{r ra-belief-trajectory, fig.width=9, fig.height=8, fig.cap="Belief trajectory and the distribution of belief change, all four studies, both samples. Panel A: mean belief (reverse-coded, higher = more belief) at the pre-, post-, and debrief timepoints, by arm (the Studies 1-3 debrief was administered only in the bunking arm; Study 4 collected it in both). Panel B: survival curve of direction-aligned belief change -- the share of participants whose change exceeds x, so a HIGHER curve means more participants moved further toward the assigned position (more effective). On the in-window analytic sample bunking and debunking are close; GPT-5.2 is the one model whose ordering reverses (debunking sits clearly above bunking, because GPT-5.2 bunking backfires). Solid lines = full/strict sample, dashed = compliant subsample."}
fig_belief_trajectory()
```

## Attrition

The screening funnel that defines each analytic sample --- attention and bot
screens, the equivocality classifier, the baseline-belief window, and the
completed-conversation requirement --- is given once in the Methods (Per-Study
Screening Pipelines; Table \@ref(tab:me-attention) and Table
\@ref(tab:me-funnel-s4)) and, for Studies 1-3, in Figure
\@ref(fig:ra-funnel-s13) above. What matters for inference is whether the loss that
remains is *differential* by the randomized assignment, which would threaten the
treatment comparison, and, in Study 4, *how* incomplete conversations were lost.
This subsection treats both, for all four studies together.

For Studies 1-3, the question is whether post-conversation dropout depends on
assigned direction (Table \@ref(tab:ra-dropout-logistic)). A non-significant
likelihood-ratio test indicates that attrition does not track the randomized
condition.

```{r ra-dropout-logistic}
drop <- all_numbers |>
  dplyr::filter(block == "dropout_logistic_s13") |>
  dplyr::mutate(term = dplyr::recode(term,
    intercept_logit = "Intercept (log-odds)",
    debunk_vs_bunk_logit = "Debunk vs. bunk (log-odds)",
    differential_attrition_LRT = "Differential-attrition LRT",
    overall_dropout_rate = "Overall dropout rate")) |>
  dplyr::mutate(value = dplyr::case_when(
    term == "Overall dropout rate" ~ scales::percent(estimate, accuracy = 0.1),
    term == "Differential-attrition LRT" ~ fmt_p(p_value),
    TRUE ~ paste0(fmt_ci(estimate, conf_low, conf_high, 2), ", p = ", fmt_p(p_value)))) |>
  dplyr::select(model, term, value) |>
  tidyr::pivot_wider(names_from = model, values_from = value)

si_kable_styled(drop, caption = "Studies 1-3 differential-attrition tests: logistic regression of post-conversation dropout on assigned direction, by GPT-4o variant. The LRT row reports the likelihood-ratio p-value for direction-dependent attrition; other rows report estimate [95% CI], p.")
```

For Study 4, the `r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="pool") - est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="completed"))`
participants who reached the conversation stage but did not complete the outcome
measures are classified by *how* the conversation was lost, into three mutually
exclusive states. The classification is read directly from the continuously-saved
partial-chat log (the `__js_chatPartialData1` Qualtrics field) by counting non-empty
user and model messages --- it is a parse of the saved conversation data, **not** a
separate model or API judgement --- and the identical rule is applied to Studies 1-3.
A **substantive mid-conversation dropout** is a participant
whose conversation worked --- the model replied at least once --- and who then
abandoned the study. A **"did not begin the conversation"** loss is a participant
whose saved log records no sent message; this category is inherently ambiguous,
spanning both a participant who reached a working chat and left before typing and a
participant for whom the chat interface never loaded. The log can separate an
*empty* record (chat never loaded) from a *system-prompt-only* record (loaded, then
left before sending), but cannot attribute participant- versus technical-cause at
the row level. A **"sent a message, no model
reply"** loss would be a genuine in-conversation delivery failure --- and there are
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="technical_no_reply_n"))`
of them.

Three facts follow (Table \@ref(tab:ra-attrition-split), Figure
\@ref(fig:ra-attrition-fig)). First, substantive mid-conversation dropout is small
and even across models ---
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="substantive_midchat_n"))`
in total, 9--20 per model --- so participants who actually engaged rarely
abandoned. Second, almost all of the loss is the
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="no_message_n"))`
participants for whom **no partial chat log was ever written**: the chat
interface never loaded, or it loaded but they left before clicking start and
seeing the AI's first message. The partial-chat log is written by the survey
only after the chat app initializes and posts its message history back to
Qualtrics (the `setJSEmbeddedData("chatPartialData1", ...)` call fires only in
that callback), so an empty log means the conversation never began at all. This
loss is overwhelmingly Gemini:
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model="Gemini", term="no_message_n"))`
of the `r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="no_message_n"))`,
versus 96--150 for each of the other three models. Because model
assignment was random, a participant-side cause would fall roughly uniformly
across models; the Gemini excess instead points to a technical cause --- the chat
interface failing to load --- which is exactly what one would expect, since Google
was deprecating Gemini 3 during fielding. The non-Gemini "did not begin" counts
therefore reflect ordinary pre-conversation non-engagement, whereas Gemini's
excess is the genuine technical loss; that technical component is identifiable at
the model level rather than per participant. Either way, *substantive* attrition --- a working
conversation that the participant abandoned --- is low and even across models, so
the analytic samples are not differentially shaped by participant disengagement.

```{r ra-attrition-split}
.lab <- c(pool = "Model-assigned pool", completed = "Completed (analytic)",
          substantive_midchat_n = "Substantive mid-conversation dropout",
          no_message_n = "Did not begin the conversation",
          technical_no_reply_n = "Sent a message, no model reply",
          substantive_attrition_rate = "Substantive attrition rate")
att <- all_numbers |>
  dplyr::filter(block == "attrition_substantive_vs_technical",
                model %in% c("Claude", "Gemini", "GPT-5.2", "Grok"),
                term %in% names(.lab)) |>
  dplyr::mutate(Stage = factor(.lab[term], levels = .lab),
                model = factor(model, levels = c("Claude", "Gemini", "GPT-5.2", "Grok")),
                value = ifelse(term == "substantive_attrition_rate",
                               scales::percent(estimate, accuracy = 0.1), fmt_num(estimate, 0))) |>
  dplyr::arrange(Stage, model) |>
  dplyr::select(Stage, model, value) |>
  tidyr::pivot_wider(names_from = model, values_from = value)

si_kable_styled(att, caption = "Study 4 outcome-incomplete loss by model (strict sample): substantive mid-conversation dropout, 'did not begin the conversation' (reached the chat but never sent a message), and 'sent a message, no model reply'. The middle category dominates and is concentrated in Gemini, consistent with a chat that failed to load during Gemini 3's deprecation; no started conversation went unanswered. The substantive attrition rate is substantive mid-conversation dropout divided by (completed + that dropout).")
```

The likelihood-ratio tests in Table \@ref(tab:ra-attrition-lrt) make this split
formal, but the definition of "attrition" must be read carefully. The first two
rows test the *overall* incomplete-outcome indicator --- every participant who did
not finish the outcome measures, regardless of how the conversation was lost. On
that definition the by-model test is significant, but only because it is dominated
by Gemini's chat-load failures: it is detecting the technical loss documented
above, not condition-dependent disengagement. The substantive-only rows isolate
mid-conversation dropout among
the
`r fmt_n(nn("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="substantive_attrition_by_direction_LRT"))`
participants who genuinely began the conversation (the model replied at least
once). On that engaged subset, dropout does not vary by assigned direction
(*p* = `r fmt_p(pv("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="substantive_attrition_by_direction_LRT"))`)
or by model category
(*p* = `r fmt_p(pv("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="substantive_attrition_by_model_LRT"))`).
Genuine disengagement is therefore flat across condition and model, and the
analytic samples are not differentially shaped by it.

```{r ra-attrition-lrt}
att_lrt <- all_numbers |>
  dplyr::filter(block == "attrition_substantive_vs_technical", is.na(model),
                term %in% c("differential_attrition_by_direction_LRT",
                            "differential_attrition_by_model_LRT",
                            "substantive_attrition_by_direction_LRT",
                            "substantive_attrition_by_model_LRT")) |>
  dplyr::mutate(
    Outcome = ifelse(grepl("^substantive", term),
                     "Substantive mid-conversation dropout (engaged only)",
                     "Overall incomplete outcome"),
    test = dplyr::recode(term,
      differential_attrition_by_direction_LRT = "By assigned direction",
      differential_attrition_by_model_LRT = "By model category",
      substantive_attrition_by_direction_LRT = "By assigned direction",
      substantive_attrition_by_model_LRT = "By model category"),
    Outcome = factor(Outcome, levels = c("Overall incomplete outcome",
                                         "Substantive mid-conversation dropout (engaged only)"))) |>
  dplyr::arrange(Outcome, test) |>
  dplyr::select(Outcome, Test = test, n, p = p_value) |>
  dplyr::mutate(p = fmt_p(p))

si_kable_styled(att_lrt, caption = "Study 4 likelihood-ratio tests for differential attrition by assigned direction and by model category, under two attrition definitions. The overall-incomplete-outcome tests (n = 2,114) include the pre-conversation chat-load loss, so the by-model test is significant only because of Gemini's technical failures. The substantive tests restrict to the participants who actually began the conversation (n = 1,333 engaged) and test mid-conversation dropout; both are non-significant, so genuine disengagement does not vary by condition or model.")
```

```{r ra-attrition-fig, fig.cap="Outcome-incomplete loss by model condition across all four studies, as a percentage of each model-assigned pool (pool sizes differ across studies): substantive mid-conversation dropout (the model replied, then the participant abandoned the study) versus 'did not begin the conversation' (the saved partial-chat log records no sent message -- chat never loaded, or loaded but left before typing). All four studies are classified by the same partial-chat-log rule. Studies 1-3 lose a small share to each (the 'did not begin' loss there is ordinary pre-conversation non-engagement, with no chat-load-failure channel); in Study 4, Gemini's large 'did not begin' bar reflects chat-load failures during its deprecation. 'Sent a message, no model reply' is zero in every study."}
fig_attrition()
```

```{r ra-attrition-diff-tests}
adt <- all_numbers |>
  dplyr::filter(block == "attrition_tests", sample == "strict_n1272") |>
  dplyr::mutate(Term = dplyr::recode(term,
                  direction_main = "Direction (randomized)",
                  model_main = "Model category",
                  direction_x_model = "Direction x model"),
                F = fmt_num(statistic, 2),
                df = paste0(df_num, ", ", df_den),
                p = fmt_p(p_value)) |>
  dplyr::select(Term, `n` = n, `F` = F, `df` = df, p)

si_kable_styled(adt, caption = "Study 4 differential-attrition test: joint F-tests from a linear-probability model of completion on assigned direction, model category, and their interaction (HC3-robust standard errors). Completion does not depend on the randomized direction or on its interaction with model.")
```

Third, the loss is localized per model (Table \@ref(tab:ra-attrition-chat-bymodel)):
the "no chat content" state --- a participant for whom the chat never loaded or who
never engaged --- dominates and is overwhelmingly Gemini, the same chat-load
signature seen above, while the partial-then-lost and saved-but-incomplete states
are small and even across the four models.

```{r ra-attrition-chat-bymodel}
.clab <- c(chat_saved_missing_outcomes = "Saved chat, missing outcomes",
           no_chat_content = "No chat content",
           partial_chat_then_lost = "Partial chat, then lost")
acm <- all_numbers |>
  dplyr::filter(block == "attrition_chat_diagnosis_by_model", sample == "strict_n1272") |>
  dplyr::mutate(State = factor(.clab[term], levels = .clab),
                model = factor(model, levels = c("Claude", "Gemini", "GPT-5.2", "Grok")),
                value = fmt_num(estimate, 0)) |>
  dplyr::arrange(State, model) |>
  dplyr::select(State, model, value) |>
  tidyr::pivot_wider(names_from = model, values_from = value)

si_kable_styled(acm, caption = "Study 4 outcome-incomplete loss by model (strict sample), classified as a saved chat with missing outcomes, no chat content, or a partial chat then lost. 'No chat content' dominates and is concentrated in Gemini, consistent with the chat interface failing to load during Gemini 3's deprecation; the other two states are small and even across models.")
```

## Engagement

Conversation engagement is summarized from the transcripts so that exposure can
be compared across model conditions (Table \@ref(tab:ra-engagement)).

```{r ra-engagement}
eng <- all_numbers |>
  dplyr::filter(block == "engagement") |>
  dplyr::mutate(direction = stringr::str_to_title(direction)) |>
  dplyr::mutate(value = dplyr::case_when(
    term == "pct_ended_before_10turn_max" ~ scales::percent(estimate, accuracy = 1),
    term %in% c("mean_user_messages", "mean_assistant_messages") ~ fmt_num(estimate, 2),
    TRUE ~ fmt_ci(estimate, conf_low, conf_high, 0))) |>
  dplyr::mutate(term = dplyr::recode(term,
    mean_user_messages = "User msgs",
    mean_assistant_messages = "Asst msgs",
    mean_user_words = "User words [95% CI]",
    mean_assistant_words = "Asst words [95% CI]",
    pct_ended_before_10turn_max = "% ended early")) |>
  dplyr::mutate(
    study = ifelse(model %in% c("Claude", "Gemini", "GPT-5.2", "Grok"), "Study 4", "Studies 1-3"),
    model = factor(model, levels = c("Jailbroken", "Standard", "Truth-Constrained",
                                     "Claude", "Gemini", "GPT-5.2", "Grok")),
    term = factor(term, levels = c("User msgs", "Asst msgs", "User words [95% CI]",
                                   "Asst words [95% CI]", "% ended early"))) |>
  dplyr::arrange(model, direction, term) |>
  dplyr::select(study, model, direction, n, term, value) |>
  tidyr::pivot_wider(names_from = term, values_from = value)

si_kable_styled(eng, caption = "Conversation engagement by model condition and assigned direction across all four studies: mean user and assistant message counts, mean word counts (with 95% CIs), and the share of conversations ending before the turn cap.")
```

Beyond cell means, Figure \@ref(fig:ra-conv-length) shows the full distribution of
conversation length --- the number of participant messages per conversation --- by
study and arm. The distributions are similar across arms within each study and
concentrate at low message counts: only a small minority of conversations (1-7%
per cell) reach the 10-user-message cap, confirming that exposure is not
systematically different between bunking and debunking.

```{r ra-conv-length, fig.width=9, fig.height=6, fig.cap="Distribution of the number of participant (user) messages per conversation, by study/model and assigned direction (strict sample; the compliant subsample is near-identical). The dotted line marks the 10-user-message cap; 93-99% of conversations end before reaching it."}
fig_conversation_length()
```

## Compliance

The same first-turn attempt-to-persuade
(APE) instrument scores all four studies: Studies 1-3 report the attempt rate
(Table \@ref(tab:ra-compliance-s13)), and Study 4 additionally reports the
refusal rate and the strict-compliance rate (an attempt with no refusal) per cell
(Table \@ref(tab:ra-compliance-s4)). The attempt rate is near ceiling everywhere
except Truth-Constrained GPT-4o bunking and GPT-5.2 bunking, where the model
frequently declined the assigned task.

```{r ra-compliance-s13}
comp13 <- all_numbers |>
  dplyr::filter(block == "compliance", term == "ape_attempt_rate") |>
  dplyr::transmute(model, direction, n,
                   `attempt rate` = scales::percent(estimate, accuracy = 1)) |>
  tidyr::pivot_wider(names_from = direction, values_from = c(n, `attempt rate`)) |>
  dplyr::select(model, `n (bunk)` = n_Bunking, `bunk attempt` = `attempt rate_Bunking`,
                `n (debunk)` = n_Debunking, `debunk attempt` = `attempt rate_Debunking`)

si_kable_styled(comp13, caption = "Studies 1-3 APE compliance: first-turn attempt rate by GPT-4o variant and assigned direction.")
```

```{r ra-compliance-s4}
comp4 <- all_numbers |>
  dplyr::filter(block == "compliance_cells", sample == "strict_n1272",
                term %in% c("attempt_rate", "refusal_rate", "strict_compliance_rate")) |>
  dplyr::mutate(value = scales::percent(estimate, accuracy = 1),
                term = dplyr::recode(term,
                  attempt_rate = "Attempt", refusal_rate = "Refusal",
                  strict_compliance_rate = "Strict compliant")) |>
  dplyr::select(model, direction, n, term, value) |>
  tidyr::pivot_wider(names_from = term, values_from = value)

si_kable_styled(comp4, caption = "Study 4 first-turn compliance by model and assigned direction (strict sample): attempt, refusal, and strict-compliance rates.")
```

Collapsing over direction makes the model-level differences in compliance easier
to read. Table \@ref(tab:ra-compliance-model-overall) reports each Study 4
model's overall first-turn attempt, refusal, and strict-compliance rates,
marginal over the bunking and debunking arms, on the strict sample. Three of the
four models comply with the assigned task almost always; GPT-5.2 is the clear
exception, attempting the assigned persuasion in only about
`r fmt_pct(est("compliance_model_overall", model="GPT-5.2", term="attempt_rate"), 0)`
of conversations and refusing in roughly
`r fmt_pct(est("compliance_model_overall", model="GPT-5.2", term="refusal_rate"), 0)`,
which is why the strict and compliant Study 4 estimates diverge mainly through
this one model.

```{r ra-compliance-model-overall}
cmo <- all_numbers |>
  dplyr::filter(block == "compliance_model_overall", sample == "strict_n1272") |>
  dplyr::mutate(term = dplyr::recode(term,
                  attempt_rate = "Attempt", refusal_rate = "Refusal",
                  strict_compliance_rate = "Strict compliant"),
                value = scales::percent(estimate, accuracy = 1),
                model = factor(model, levels = c("Claude", "Gemini", "GPT-5.2", "Grok"))) |>
  dplyr::arrange(model) |>
  dplyr::select(model, n, term, value) |>
  tidyr::pivot_wider(names_from = term, values_from = value) |>
  dplyr::select(model, n, Attempt, Refusal, `Strict compliant`)

si_kable_styled(cmo, caption = "Study 4 first-turn compliance by model, collapsed over assigned direction (strict sample): overall attempt, refusal, and strict-compliance rates. GPT-5.2 is the lone low-compliance model.")
```

```{r ra-compliance-fig, fig.cap="First-turn attempt and strict-compliance rates by model condition and assigned direction across all four studies."}
fig_compliance_4study()
```

Restricting to the compliant subsample raises the Truth-Constrained and GPT-5.2
bunking means — most steeply GPT-5.2, from a strong backfire to a near-null
effect — because the participants who drove the negative strict-sample mean were
exactly those whose model declined to bunk (the strict and compliant cell means
sit side by side in Table \@ref(tab:ra-belief-cells)).

### What Non-Compliant Models Did Instead

A first-turn attempt rate below ceiling raises the deployment-relevant question of
what the model did when it declined the assigned task. The compliance classifier
distinguishes two ways of not attempting: an **explicit refusal** (the model
declines without taking a side) and a **counter-argument** (the model actively
argues *against* the position it was assigned to advocate). The distinction
matters because the two have opposite consequences for belief: a refusal leaves the
participant where they started, whereas a counter-argument is itself a persuasion
attempt in the reverse direction. Table \@ref(tab:ra-nonattempt-split) breaks the
Study 4 non-attempts down this way for each model and direction, with the aligned
belief change among the counter-argument conversations.

The pattern is asymmetric and points one way, and it is overwhelmingly driven by
GPT-5.2, the lone low-compliance model. In the **bunking** arm, when GPT-5.2
declined to spread the conspiracy it usually did not merely refuse: it
counter-argued in
`r fmt_n(est("nonattempt_split_by_model", sample="strict_n1272", model="GPT-5.2", direction="Bunking", term="n_counterargument"))`
of its
`r fmt_n(est("nonattempt_split_by_model", sample="strict_n1272", model="GPT-5.2", direction="Bunking", term="n_nonattempt"))`
bunk non-attempts, and those conversations moved aligned belief about
`r fmt_est_ci(num("nonattempt_split_by_model", sample="strict_n1272", model="GPT-5.2", direction="Bunking", term="mean_aligned_change_counterargument"), 1, signed = TRUE)`
points --- a large shift *against* the assigned bunking direction, i.e. toward
rejecting the conspiracy. The other three models declined far less often (5, 9,
and 10 bunk non-attempts for Claude, Gemini, and Grok respectively), and only Grok
refused outright rather than counter-arguing. The **debunking** arm shows no such
behavior in any model: GPT-5.2's
`r fmt_n(est("nonattempt_split_by_model", sample="strict_n1272", model="GPT-5.2", direction="Debunking", term="n_nonattempt"))`
debunking non-attempts were all explicit refusals, with
`r fmt_n(est("nonattempt_split_by_model", sample="strict_n1272", model="GPT-5.2", direction="Debunking", term="n_counterargument"))`
counter-arguments, and the remaining models likewise never argued *for* a
conspiracy they had been asked to debunk. The flip side of compliance is therefore
not inertia but
inversion: a safety-trained model instructed to spread a conspiracy frequently
debunks it instead, which is why the strict-sample bunking mean for GPT-5.2 is so
much more negative than its compliant-sample mean.

The same picture holds when the GPT-5.2 bunking cell is split into attempts versus
declines: among the
`r fmt_n(nn("gpt52_bunk_decomposition", sample="strict_n1272", outcome="aligned_belief_change", term="attempted"))`
conversations where it *attempted* the bunk, belief change is small and not
distinguishable from zero
(`r fmt_est_ci(num("gpt52_bunk_decomposition", sample="strict_n1272", outcome="aligned_belief_change", term="attempted"), 2, signed = TRUE)`),
and the public post still moved *against* the conspiracy in both the attempted
(`r fmt_est_ci(num("gpt52_bunk_decomposition", sample="strict_n1272", outcome="aligned_stance_change", term="attempted"), 2, signed = TRUE)`)
and declined
(`r fmt_est_ci(num("gpt52_bunk_decomposition", sample="strict_n1272", outcome="aligned_stance_change", term="declined"), 2, signed = TRUE)`)
subsets.

```{r ra-nonattempt-split}
.nlab <- c(n_nonattempt = "Non-attempts",
           n_explicit_refusal = "Explicit refusals",
           n_counterargument = "Counter-arguments",
           share_counterargument = "Counter-arg. share",
           mean_aligned_change_counterargument = "Counter-arg. aligned change [95% CI]")
nas <- all_numbers |>
  dplyr::filter(block == "nonattempt_split_by_model", sample == "strict_n1272",
                model != "All models") |>
  dplyr::mutate(value = dplyr::case_when(
    term == "share_counterargument" ~ scales::percent(estimate, accuracy = 1),
    term == "mean_aligned_change_counterargument" & !is.na(conf_low) ~
      fmt_ci(estimate, conf_low, conf_high, 1),
    term == "mean_aligned_change_counterargument" ~ "—",
    TRUE ~ fmt_num(estimate, 0)),
    Quantity = factor(.nlab[term], levels = .nlab),
    model = factor(model, levels = c("Claude", "Gemini", "GPT-5.2", "Grok")),
    direction = factor(direction, levels = c("Bunking", "Debunking"))) |>
  dplyr::arrange(direction, model, Quantity) |>
  dplyr::select(Direction = direction, Model = model, Quantity, value) |>
  tidyr::pivot_wider(names_from = Quantity, values_from = value)

si_kable_styled(nas, caption = "Study 4 non-attempts split by type, for each model and assigned direction (strict sample; compliance-scored conversations that did not attempt the assigned direction). 'Counter-arguments' are conversations in which the model argued against the position it was assigned to advocate; 'Counter-arg. aligned change' is the mean direction-aligned belief change among those conversations, with a 95% CI where n is adequate. In the bunking arm most non-attempts are counter-arguments that move belief sharply against the assigned direction; in the debunking arm every non-attempt is an explicit refusal.")
```

## The Study 3 Truth-Constraint Counterfactual

The Truth-Constrained variant asks how much of GPT-4o's bunking power survives
when the model may use only accurate arguments. Table
\@ref(tab:ra-s3-counterfactual) gives the truth-constraint penalty two ways: the
compliant bunking-minus-debunking contrast on the matched $n = 740$ counterfactual
set, and the belief change in the Truth-Constrained bunking conversations that
were non-compliant (the model declined to fabricate). The penalty survives among
compliant conversations, and the non-compliant cases moved belief sharply
*against* the conspiracy.

```{r ra-s3-counterfactual}
cf <- all_numbers |>
  dplyr::filter(block == "s3_counterfactual") |>
  dplyr::transmute(quantity = "Compliant bunk - debunk (matched counterfactual)", n,
                   `estimate [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

nc <- all_numbers |>
  dplyr::filter(block == "noncompliant_s3_bunk") |>
  dplyr::transmute(quantity = "Belief change in non-compliant TC bunking conversations", n,
                   `estimate [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

si_kable_styled(dplyr::bind_rows(cf, nc), caption = "Study 3 truth-constraint counterfactual: the compliant bunking-minus-debunking contrast on the matched counterfactual set, and the belief change in Truth-Constrained bunking conversations where the model declined to comply.")
```

## Secondary Outcomes

Three families of secondary outcomes are reported. The first is generalized
conspiracist belief (GCBS), administered pre and post in Studies 1-3; Table
\@ref(tab:ra-gcbs) gives the pre-to-post change per cell and the
bunking-minus-debunking contrast. The second is trust in AI, measured before and
after the conversation in all four studies (`genai_trust` pre, `trust2` post;
Study 4 additionally re-administered the item after the corrective debrief,
`trust_debrf`): Studies 1-3 report the pre-to-post change per variant (Table
\@ref(tab:ra-trust)), and Study 4's pre-to-post change — computed on the same
footing — joins the cross-study table below alongside the registered
post-treatment contrast. The third is the set of perception
measures — argument strength, provided-new-information, collaborativeness, and
impartiality — characterizing *how* the conversation was experienced (Table
\@ref(tab:ra-perception)).

```{r ra-gcbs}
gcbs_cells <- all_numbers |>
  dplyr::filter(block == "gcbs_change", term == "GCBS_change_mean") |>
  dplyr::transmute(model, direction, n,
                   ci = fmt_ci(estimate, conf_low, conf_high, 2)) |>
  tidyr::pivot_wider(names_from = direction, values_from = c(n, ci))

gcbs_con <- all_numbers |>
  dplyr::filter(block == "gcbs_change", term == "bunk_minus_debunk") |>
  dplyr::transmute(model, `bunk - debunk [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

gcbs <- dplyr::left_join(gcbs_cells, gcbs_con, by = "model") |>
  dplyr::select(model, `n (bunk)` = n_Bunking, `bunk [95% CI]` = ci_Bunking,
                `n (debunk)` = n_Debunking, `debunk [95% CI]` = ci_Debunking,
                `bunk - debunk [95% CI]`, p)

si_kable_styled(gcbs, caption = "Studies 1-3 generalized conspiracist belief (GCBS): pre-to-post change by assigned direction and the within-variant bunking-minus-debunking contrast.")
```

```{r ra-trust}
trust_cells <- all_numbers |>
  dplyr::filter(block == "trust_change", term == "trust_pre_post_change_mean") |>
  dplyr::transmute(model, direction, ci = fmt_ci(estimate, conf_low, conf_high, 2)) |>
  tidyr::pivot_wider(names_from = direction, values_from = ci)

trust_con <- all_numbers |>
  dplyr::filter(block == "trust_change", term == "bunk_minus_debunk") |>
  dplyr::transmute(model, n, `bunk - debunk [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

trust <- dplyr::left_join(trust_cells, trust_con, by = "model") |>
  dplyr::select(model, `bunk [95% CI]` = Bunking, `debunk [95% CI]` = Debunking,
                n, `bunk - debunk [95% CI]`, p)

si_kable_styled(trust, caption = "Studies 1-3 trust in AI: pre-to-post change by assigned direction and the within-variant bunking-minus-debunking contrast. A positive bunking trust effect that survives the corrective disclosure indicates trust is not undercut by exposure to fabricated arguments.")
```

Table \@ref(tab:ra-trust-allstudies) places the Studies 1-3 per-variant
bunking-minus-debunking trust contrast (pre-to-post change) alongside two Study 4
rows: the same pre-to-post change contrast (strict sample, models pooled) and the
registered contrast on the post-treatment trust item, the latter on both the
strict (intention-to-treat) and compliant samples. The contrast is small in
Studies 1-3 — significant only for the Jailbroken variant — with the Study 4
rows shown on both the change and post-only footings. Study 4 also re-administered the
trust item after the corrective debrief (`trust_debrf`); the post-to-debrief
trust change is reported in block `s4_trust_change`, and the post-debrief belief
trajectory is reported in the Durability section below (Table
\@ref(tab:rb-debrief-s4) and Table \@ref(tab:rb-debrief-traj)).

```{r ra-trust-allstudies}
trust13 <- all_numbers |>
  dplyr::filter(block == "trust_change", term == "bunk_minus_debunk") |>
  dplyr::transmute(study = model, sample = "Pre-to-post change", n,
                   `bunk - debunk [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

trust4chg <- all_numbers |>
  dplyr::filter(block == "s4_trust_change", term == "bunk_minus_debunk",
                outcome == "Trust in AI (pre->post)", sample == "strict_n1272") |>
  dplyr::transmute(study = "Study 4", sample = "Pre-to-post change (strict)",
                   n, `bunk - debunk [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

trust4 <- all_numbers |>
  dplyr::filter(block == "secondary_contrasts", outcome == "trust2",
                sample %in% c("strict_n1272", "compliant_n1073")) |>
  dplyr::transmute(study = "Study 4",
                   sample = ifelse(sample == "strict_n1272",
                                   "Post-treatment (strict)", "Post-treatment (compliant)"),
                   n, `bunk - debunk [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

trust_all <- dplyr::bind_rows(trust13, trust4chg, trust4) |>
  dplyr::mutate(study = factor(study, levels = c("Jailbroken", "Standard",
                                                 "Truth-Constrained", "Study 4"))) |>
  dplyr::arrange(study) |>
  dplyr::select(study, sample, n, `bunk - debunk [95% CI]`, p)

si_kable_styled(trust_all, caption = "Treatment effect on trust in AI across all four studies. Studies 1-3 report the within-variant bunking-minus-debunking contrast in pre-to-post trust change; Study 4 reports the same pre-to-post change contrast (strict sample, models pooled) plus the registered contrast on the post-treatment trust item in both the strict (intention-to-treat) and compliant samples. The effect is small in Studies 1-3.")
```

The perception contrasts appear by item for all four studies, with the strict
and compliant samples in adjacent columns (Table \@ref(tab:ra-perception)).
The debunking arm is perceived as supplying more new information than the bunking
arm in every study and in both Study 4 samples; the other three dimensions move
little. Figure \@ref(fig:ra-secondary-fig) renders the same contrasts, with each
Studies 1-3 variant shown separately and Study 4 shown as the across-model
perception contrast (the per-model breakdown is in Table
\@ref(tab:ra-perception-bymodel)).

```{r ra-perception}
em <- function(x) ifelse(is.na(x), "—", x)
percep13_strict <- all_numbers |>
  dplyr::filter(block == "perception_mediators", term == "bunk_minus_debunk",
                sample == "full_sample") |>
  dplyr::transmute(study = model, item = outcome, n_strict = as.character(n),
                   strict = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_strict = fmt_p(p_value))

percep13_comp <- all_numbers |>
  dplyr::filter(block == "perception_mediators", term == "bunk_minus_debunk",
                sample == "compliant") |>
  dplyr::transmute(study = model, item = outcome, n_comp = as.character(n),
                   comp = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_comp = fmt_p(p_value))

percep13 <- dplyr::left_join(percep13_strict, percep13_comp, by = c("study", "item"))

s4_map <- c(ArgStrength = "Argument strength", new_info_rc = "Provided new information",
            Collaborative_rc = "Collaborativeness", Unbiased_rc = "Impartiality (unbiased)")
percep4_strict <- all_numbers |>
  dplyr::filter(block == "secondary_contrasts", sample == "strict_n1272",
                outcome %in% names(s4_map)) |>
  dplyr::transmute(study = "Study 4", item = unname(s4_map[outcome]), n_strict = as.character(n),
                   strict = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_strict = fmt_p(p_value))

percep4_comp <- all_numbers |>
  dplyr::filter(block == "secondary_contrasts", sample == "compliant_n1073",
                outcome %in% names(s4_map)) |>
  dplyr::transmute(study = "Study 4", item = unname(s4_map[outcome]), n_comp = as.character(n),
                   comp = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_comp = fmt_p(p_value))

percep4 <- dplyr::left_join(percep4_strict, percep4_comp, by = c("study", "item"))

percep <- dplyr::bind_rows(percep13, percep4) |>
  dplyr::mutate(study = factor(study, levels = c("Jailbroken", "Standard",
                                                 "Truth-Constrained", "Study 4")),
                dplyr::across(c(n_comp, comp, p_comp), em)) |>
  dplyr::arrange(item, study) |>
  dplyr::select(item, study,
                `n (strict)` = n_strict, `bunk - debunk, strict [95% CI]` = strict, `p` = p_strict,
                `n (compliant)` = n_comp, `bunk - debunk, compliant [95% CI]` = comp, `p ` = p_comp)

si_kable_styled(percep, caption = "Perception measures across all four studies: bunking-minus-debunking contrasts in perceived argument strength, provision of new information, collaborativeness, and impartiality. All four studies report both the strict (intention-to-treat) and compliant samples; Studies 1-3 use the pre-to-post analytic sample and its compliant subset, and Study 4 uses the strict and compliant post-treatment contrasts.")
```

```{r ra-secondary-fig, fig.cap="Bunking-minus-debunking contrasts on the perception measures (argument strength, new information, collaborativeness, impartiality) across all four studies, with 95 percent confidence intervals. Studies 1-3 are shown by GPT-4o variant and Study 4 by individual frontier model (no pooling), each on BOTH the strict (intention-to-treat) and compliant samples (solid points/lines = Strict; hollow points/dashed lines = Compliant)."}
fig_secondary()
```

The Study 4 perception contrasts are also broken out by model in Table
\@ref(tab:ra-perception-bymodel). The model-level pattern is muted and mostly
non-significant --- the clearest signals are that Claude and Grok were rated as
providing more new information under bunking --- consistent with the small pooled
perception contrasts.

```{r ra-perception-bymodel}
pbm <- all_numbers |>
  dplyr::filter(block == "perception_s4_by_model",
                sample %in% c("strict_n1272", "compliant_n1073")) |>
  dplyr::mutate(samp = dplyr::recode(sample, strict_n1272 = "strict", compliant_n1073 = "comp"),
                Model = factor(model, levels = c("Claude", "Gemini", "GPT-5.2", "Grok")),
                ci = fmt_ci(estimate, conf_low, conf_high, 2),
                p = fmt_p(p_value)) |>
  dplyr::select(Model, Item = term, samp, n, ci, p) |>
  tidyr::pivot_wider(names_from = samp, values_from = c(n, ci, p)) |>
  dplyr::arrange(Model, Item) |>
  dplyr::select(Model, Item,
                `n (strict)` = n_strict, `bunk - debunk, strict [95% CI]` = ci_strict, `p` = p_strict,
                `n (compliant)` = n_comp, `bunk - debunk, compliant [95% CI]` = ci_comp, `p ` = p_comp)
si_kable_styled(pbm, caption = "Study 4 perception measures by model: bunking-minus-debunking contrast on each item (raw response scale, HC3), reported on both the strict (intention-to-treat) and compliant samples. Positive = rated higher under bunking.", digits = 2)
```

### Pre-Treatment AI-Study Experience (Study 4)

Before the conversation, Study 4 participants reported how many prior studies
they had completed that they knew involved an AI, and how many they thought
likely involved an AI; both counts are right-skewed with a wide range
(clear-AI studies: median
`r fmt_num(est("ai_study_experience", outcome="howmanystudiesclear", term="median"), 0)`,
range `r fmt_num(est("ai_study_experience", outcome="howmanystudiesclear", term="min"), 0)`--`r fmt_num(est("ai_study_experience", outcome="howmanystudiesclear", term="max"), 0)`;
any-AI studies: median
`r fmt_num(est("ai_study_experience", outcome="howmanystudiestotal", term="median"), 0)`,
range `r fmt_num(est("ai_study_experience", outcome="howmanystudiestotal", term="min"), 0)`--`r fmt_num(est("ai_study_experience", outcome="howmanystudiestotal", term="max"), 0)`),
indicating most participants were relatively new to AI studies while a small tail
reported extensive prior exposure (Table \@ref(tab:ra-ai-experience)).

```{r ra-ai-experience}
aix <- all_numbers |>
  dplyr::filter(block == "ai_study_experience") |>
  dplyr::mutate(Item = dplyr::recode(outcome,
                  howmanystudiesclear = "Prior studies known to involve AI",
                  howmanystudiestotal = "Prior studies likely to involve AI"),
                term = factor(term, levels = c("n_answered","min","p25","median","p75","max","mean"))) |>
  dplyr::filter(!is.na(term)) |>
  dplyr::mutate(Statistic = dplyr::recode(term,
                  n_answered = "N answered", min = "Min", p25 = "25th pct",
                  median = "Median", p75 = "75th pct", max = "Max", mean = "Mean"),
                Value = fmt_num(estimate, 1)) |>
  dplyr::arrange(Item, term) |>
  dplyr::select(Item, Statistic, Value) |>
  tidyr::pivot_wider(names_from = Statistic, values_from = Value)
si_kable_styled(aix, caption = "Study 4 self-reported prior AI-study experience: distribution of the number of prior studies participants knew involved an AI, and the number they thought likely involved an AI.")
```

## Durability and the Corrective Debrief

```{r rb-debrief-s13}
db13 <- all_numbers |>
  dplyr::filter(section == "S1-3", block == "debrief",
                term == "post_to_debrief_shift") |>
  dplyr::mutate(`Shift [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                p = fmt_p(p_value)) |>
  dplyr::select(Study = model, Direction = direction, n, `Shift [95% CI]`, p)
si_kable_styled(db13, caption = "Studies 1-3 post-conversation to post-debrief belief shift, bunking arm.", digits = 2)
```

```{r rb-debrief-s4}
db4 <- all_numbers |>
  dplyr::filter(section == "S4 + pooled", block == "debrief_shift",
                term == "post_to_debrief_change") |>
  dplyr::mutate(Sample = dplyr::recode(sample,
                  strict_n1272 = "Strict (ITT)", compliant_n1073 = "Compliant"),
                `Shift [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                p = fmt_p(p_value)) |>
  dplyr::select(Sample, Model = model, Direction = direction, n,
                `Shift [95% CI]`, p)
si_kable_styled(db4, caption = "Study 4 post-conversation to post-debrief belief shift, both arms, by model (strict and compliant samples). n is the model x direction cell size for model rows and the full debrief-complete sample for Pooled rows.", digits = 2)
```

The trajectory in Table \@ref(tab:rb-debrief-traj) places the three measurement
points (pre-treatment, post-treatment, post-debrief) on a common belief scale so
the rise-and-correction shape is visible directly, and Figure
\@ref(fig:rb-fig-debrief-traj) renders the same trajectory across studies and
models.

```{r rb-debrief-traj}
dbt_s <- all_numbers |>
  dplyr::filter(section == "S4 + pooled", block == "debrief_trajectory",
                sample == "strict_n1272") |>
  dplyr::transmute(direction, term, n_s = n,
                   strict = fmt_ci(estimate, conf_low, conf_high, 2))
dbt_c <- all_numbers |>
  dplyr::filter(section == "S4 + pooled", block == "debrief_trajectory",
                sample == "compliant_n1073") |>
  dplyr::transmute(direction, term, n_c = n,
                   comp = fmt_ci(estimate, conf_low, conf_high, 2))
dbt <- dplyr::full_join(dbt_s, dbt_c, by = c("direction", "term")) |>
  dplyr::mutate(Timepoint = factor(term,
                  levels = c("Pre-treatment", "Post-treatment", "Post-debrief"))) |>
  dplyr::arrange(direction, Timepoint) |>
  dplyr::select(Direction = direction, Timepoint,
                `n (strict)` = n_s, `strict [95% CI]` = strict,
                `n (compliant)` = n_c, `compliant [95% CI]` = comp)
si_kable_styled(dbt, caption = "Study 4 mean belief at pre-treatment, post-treatment, and post-debrief, by direction, on the strict (intention-to-treat) and compliant samples.", digits = 2)
```

```{r rb-fig-debrief-traj, fig.cap="Post-conversation to post-debrief belief shift by study and model, both arms. For Study 4 both the strict (intention-to-treat) and compliant samples are shown (solid points/lines = Strict (ITT); hollow points/dashed lines = Compliant); Studies 1-3 have a single analytic sample."}
fig_debrief_trajectory()
```
