# Results

Two observed-outcome samples recur and are reported together throughout: the
*strict* sample retains every participant with the required post-treatment
outcomes regardless of model compliance, and the *compliant* subset retains only
conversations in which the model attempted the assigned manipulation. A separate
chat-started sensitivity analysis below includes all records with confirmed chatbot
delivery, including incomplete outcome records. A
claim is *direction-aligned* when its stance supports the position the model was
assigned to argue.

## Sample Composition and Screening

Before analysing treatment effects we document how each analytic sample was
assembled and who is in it. Figure \@ref(fig:ra-funnel-s13) traces the full
screening and exclusion pipeline for Studies 1--3, from all raw survey entrants to
the randomized analytic sample, and Table \@ref(tab:ra-demographics) reports the
general demographics of each study's analytic sample.

```{r ra-funnel-s13, fig.width=9, fig.height=8, fig.cap="Screening and exclusion pipeline for Studies 1-3. Each box gives the number of respondents surviving a stage; the red annotation on each arrow gives the number excluded there and why. The pipeline runs from all raw survey responses through the chatbot screener, the instructed-response attention check, the conspiracy-equivocality classifier, and the 25-75 baseline-belief window to the final analytic samples (1,092 / 814 / 818). A screening funnel describes who entered the analytic sample, so it has no compliant-subsample analogue; Study 4's production screening is summarized in the Attrition section."}
fig_funnel_s13()
```

```{r ra-demographics}
dg <- all_numbers |>
  dplyr::filter(block == "demographics", sample == "strict", is.na(direction)) |>
  dplyr::select(model, term, estimate) |>
  tidyr::pivot_wider(names_from = term, values_from = estimate) |>
  dplyr::mutate(
    N = fmt_n(n_cell),
    `Age M (SD)` = paste0(fmt_num(age_mean, 1), " (", fmt_num(age_sd, 1), ")"),
    `% Female` = fmt_num(pct_female, 1),
    `% White` = fmt_num(pct_white, 1),
    `% Multiracial` = fmt_num(pct_multiracial, 1),
    Education = fmt_num(education_mean, 2),
    `Conservatism (soc/econ)` = paste0(fmt_num(social_conservatism_mean, 2), " / ",
                                       fmt_num(economic_conservatism_mean, 2)),
    `Party / ideology` = ifelse(!is.na(party_demrep_mean),
      paste0("Dem-Rep ", fmt_num(party_demrep_mean, 2)),
      paste0(fmt_num(party_pct_Democrat, 0), "% Dem / ", fmt_num(party_pct_Republican, 0), "% Rep")),
    .so = match(model, c("Study 1 (Jailbroken)", "Study 2 (Standard)",
                         "Study 3 (Truth-Constrained)", "Study 4 (Claude)",
                         "Study 4 (Gemini)", "Study 4 (GPT-5.2)", "Study 4 (Grok)"))) |>
  dplyr::arrange(.so) |>
  dplyr::select(Study = model, N, `Age M (SD)`, `% Female`, `% White`, `% Multiracial`,
                Education, `Conservatism (soc/econ)`, `Party / ideology`)
si_kable_styled(dg, caption = "General demographics of each study's analytic sample: cell N, age, gender, race, education, conservatism, and party/ideology. Age is mean (SD) in years (Studies 1-3 cleaned age; Study 4 age clamped to 13-90); % Female is the share of gender-coded respondents; % White is the share selecting White alone and % Multiracial the share selecting more than one race category on the multi-select race item; Education is the mean of the 1-8 ordinal scale; Conservatism is the mean of the social / economic 5-point ideology items, whose scale runs 1 = very conservative to 5 = very liberal (collected identically in all studies; higher values therefore indicate a more liberal sample); party/ideology is the percent Democrat / Republican (Studies 1-3) or the mean Democrat-to-Republican ideology score (Study 4).")
```

## Belief Effects

The primary outcome is direction-aligned belief change on the 0-100 focal-belief
scale: post-treatment minus pre-treatment belief, with the sign flipped in the
debunking arm so that a positive number always means movement in the direction
the AI was assigned to advocate. A bunking value of $+10$ and a debunking value
of $+10$ both denote a ten-point persuasive success; a negative value denotes
backfire. Table \@ref(tab:ra-belief-cells) gives each model-by-direction cell
mean on this common scale across all four studies and both samples, alongside the
within-condition standardized effect sizes (pre/post means, their correlation, and
Cohen's $d_z$). The truth constraint's cost is visible on the standardized scale too:
the bunking $d_z$ falls from the Standard to the Truth-Constrained study, and within
Study 4 it ranges from strongly positive (Claude, Grok) to negative (GPT-5.2, whose
bunking attempts moved belief the wrong way).

```{r ra-belief-cells}
# Belief change [95% CI] per cell, with keys normalized to the effect_sizes model
# labels (study label / Title-case direction / strict|compliant) so the two tables
# can be joined into one combined belief-change + effect-size table.
.s13_relabel <- c(Jailbroken = "Study 1 (Jailbroken)", Standard = "Study 2 (Standard)",
                  `Truth-Constrained` = "Study 3 (Truth-Constrained)")
ci_cells <- dplyr::bind_rows(
  all_numbers |>
    dplyr::filter(section == "S1-3", block == "belief_change") |>
    dplyr::transmute(km = unname(.s13_relabel[model]), direction, sample = "strict",
                     ci = fmt_ci(estimate, conf_low, conf_high, 2)),
  all_numbers |>
    dplyr::filter(block == "belief_change_compliant") |>
    dplyr::transmute(km = unname(.s13_relabel[model]), direction, sample = "compliant",
                     ci = fmt_ci(estimate, conf_low, conf_high, 2)),
  all_numbers |>
    dplyr::filter(block == "raw_aligned_means_cells", outcome == "aligned_belief_change",
                  term == "raw_mean", sample %in% c("strict_n1272", "compliant_n1073")) |>
    dplyr::transmute(km = model, direction = dplyr::recode(direction, bunk = "Bunking", debunk = "Debunking"),
                     sample = dplyr::recode(sample, strict_n1272 = "strict", compliant_n1073 = "compliant"),
                     ci = fmt_ci(estimate, conf_low, conf_high, 2))
)
belief_cells <- all_numbers |>
  dplyr::filter(block == "effect_sizes", model != "Study 4 (model-weighted)") |>
  dplyr::select(model, direction, sample, term, estimate) |>
  tidyr::pivot_wider(names_from = term, values_from = estimate) |>
  dplyr::left_join(ci_cells, by = c("model" = "km", "direction", "sample")) |>
  dplyr::mutate(
    Sample = dplyr::recode(sample, strict = "Observed outcomes", compliant = "Compliant"),
    N = fmt_n(n_cell),
    `Belief change [95% CI]` = ci,
    `Pre M (SD)` = paste0(fmt_num(pre_mean, 1), " (", fmt_num(pre_sd, 1), ")"),
    `Post M (SD)` = paste0(fmt_num(post_mean, 1), " (", fmt_num(post_sd, 1), ")"),
    `Pre-post r` = fmt_num(prepost_r, 2),
    `Cohen d_z` = fmt_num(d_z, 2),
    .so = match(model, c("Study 1 (Jailbroken)", "Study 2 (Standard)",
                         "Study 3 (Truth-Constrained)", "Claude", "Gemini", "GPT-5.2", "Grok")),
    .co = match(direction, c("Bunking", "Debunking")),
    .sa = match(sample, c("strict", "compliant"))) |>
  dplyr::arrange(.so, .co, .sa) |>
  dplyr::select(Study = model, Condition = direction, Sample, N,
                `Belief change [95% CI]`, `Pre M (SD)`, `Post M (SD)`, `Pre-post r`, `Cohen d_z`)

si_kable_styled(belief_cells, caption = "Direction-aligned focal-belief change and standardized within-condition effect sizes, by study, assigned direction, and analytic sample. Belief change is the direction-aligned mean (post minus pre, sign-flipped for debunking; Study 4 already aligned) with its 95% CI; positive = movement toward the AI's assigned position, negative = backfire. Pre and Post are means (SDs) on the raw reverse-coded 0-100 belief scale (higher = more conspiracy belief); pre-post r is their correlation; Cohen's d_z divides the mean aligned change by its standard deviation. Observed outcomes includes every eligible outcome completer regardless of model compliance; Compliant is the two-sided-compliant subset.")
```

```{r ra-belief-fig, fig.cap="Direction-aligned focal-belief change by model condition and assigned direction across all four studies, with 95% confidence intervals. For Study 4 both the observed-outcome and compliant samples are shown (solid points/lines = Observed outcomes; hollow points/dashed lines = Compliant); Studies 1-3 have a single analytic sample."}
fig_belief_4study()
```

The within-study comparison of interest is whether bunking and debunking move
belief by the same amount: a contrast near zero means the model was an equally
effective bunker and debunker, and a negative value means it was a weaker bunker
than debunker (Table \@ref(tab:ra-belief-contrast)). We report the contrast on
both the full and the compliant sample.

```{r ra-belief-contrast}
full <- all_numbers |>
  dplyr::filter(block == "belief_bunk_vs_debunk", term == "bunk_minus_debunk_lin") |>
  dplyr::transmute(model, n_full = n,
                   full = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_full = fmt_p(p_value))

comp <- all_numbers |>
  dplyr::filter(block == "belief_bunk_vs_debunk_compliant", term == "bunk_minus_debunk_lin") |>
  dplyr::transmute(model, n_comp = n,
                   comp = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_comp = fmt_p(p_value))

contrast_tab <- dplyr::left_join(full, comp, by = "model") |>
  dplyr::select(model,
                `n (full)` = n_full, `bunk - debunk, full [95% CI]` = full, `p` = p_full,
                `n (compliant)` = n_comp, `bunk - debunk, compliant [95% CI]` = comp, `p ` = p_comp)

si_kable_styled(contrast_tab, caption = "Studies 1-3 within-study bunking-minus-debunking contrasts in direction-aligned belief change, from baseline-adjusted ANCOVA models (direction-aligned change on condition + baseline belief, additive) with HC3-robust standard errors, reported on both the full analytic sample and the compliant subsample.")
```

### Study 4 per-model directional contrasts

The four Study 4 model families differ enough that the per-model
debunk-minus-bunk contrast is the right unit to report directly (the single
pre-registered pooled directional contrast, equal-weighted across the four
families, is given separately as the registered estimand in Table
\@ref(tab:rb-prereg-aligned)). Table
\@ref(tab:ra-model-belief-contrast) gives that contrast on aligned belief change
for each model in both the observed-outcome and compliant
(manipulation-delivered) samples, estimated with HC3-robust standard errors. The
sign convention is debunk minus bunk, so a positive value means the model was a
more effective debunker than bunker. GPT-5.2 shows the largest asymmetry in both
samples, reflecting its frequent declination of the bunking task; Claude is
slightly negative (a marginally stronger bunker), and Gemini and Grok sit in
between.

```{r ra-model-belief-contrast}
# Per-model analytic n (sum over the two arms), not the whole-sample 1272/1073.
model_n <- all_numbers |>
  dplyr::filter(block == "raw_aligned_means_cells",
                outcome == "aligned_belief_change", term == "raw_mean",
                sample %in% c("strict_n1272", "compliant_n1073"),
                model %in% c("Claude", "Gemini", "GPT-5.2", "Grok")) |>
  dplyr::group_by(sample, model) |>
  dplyr::summarise(n_model = sum(n), .groups = "drop")

mc_strict <- all_numbers |>
  dplyr::filter(block == "aligned_model_contrasts", sample == "strict_n1272",
                outcome == "aligned_belief_change", term == "debunk_minus_bunk") |>
  dplyr::transmute(model,
                   strict = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_strict = fmt_p(p_value)) |>
  dplyr::left_join(model_n |> dplyr::filter(sample == "strict_n1272") |>
                     dplyr::transmute(model, n_strict = n_model), by = "model")

mc_comp <- all_numbers |>
  dplyr::filter(block == "aligned_model_contrasts", sample == "compliant_n1073",
                outcome == "aligned_belief_change", term == "debunk_minus_bunk") |>
  dplyr::transmute(model,
                   comp = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_comp = fmt_p(p_value)) |>
  dplyr::left_join(model_n |> dplyr::filter(sample == "compliant_n1073") |>
                     dplyr::transmute(model, n_comp = n_model), by = "model")

mc <- dplyr::left_join(mc_strict, mc_comp, by = "model") |>
  dplyr::mutate(model = factor(model, levels = c("Claude", "Gemini", "GPT-5.2", "Grok"))) |>
  dplyr::arrange(model) |>
  dplyr::select(model,
                `n (strict)` = n_strict, `debunk - bunk, strict [95% CI]` = strict, `p` = p_strict,
                `n (compliant)` = n_comp, `debunk - bunk, compliant [95% CI]` = comp, `p ` = p_comp)

si_kable_styled(mc, caption = "Study 4 per-model debunk-minus-bunk contrasts in direction-aligned belief change (HC3-robust standard errors), reported on both the observed-outcome and compliant samples. A positive value indicates the model was a more effective debunker than bunker.")
```

Across the per-model rows the debunk-minus-bunk asymmetry is concentrated in
GPT-5.2 and shrinks once its declined bunking conversations are removed (the
strict-to-compliant change in Table \@ref(tab:ra-model-belief-contrast)); the
other three models sit near zero in both samples. We report these four
per-model contrasts rather than a single equal-model-weighted pool, since the
families differ too much for a pooled directional estimate to be broadly
interpretable; the one pooled contrast that the pre-registration does specify is
reported, labeled as the registered estimand, in Table
\@ref(tab:rb-prereg-aligned).

## Distributional Shift and Large Shifts

Mean contrasts can mask differences in the tails: two conditions can have
similar average effects while differing in how often they produce a large
conversion. Two complementary measures address this. The first is a
two-sample Kolmogorov-Smirnov test comparing the full distribution of aligned
belief change in the bunking and debunking arms, reported as the maximal
distributional distance $D$ with its $p$-value for all seven model conditions,
on both samples (Table \@ref(tab:ra-distribution-ks)). The
second counts the share of participants whose aligned change exceeds a fixed
magnitude. Because no single threshold is privileged, we report four — $\geq 20$,
$\geq 30$, $\geq 40$, and $\geq 50$ scale points — and test the
bunking-versus-debunking difference at each threshold with a chi-square statistic
(Table \@ref(tab:ra-large-shift)). The Study 4 distributions differ in the
GPT-5.2 cell, where bunking rarely produces a large conversion; elsewhere the
arms are close.

```{r ra-distribution-ks}
ks13_strict <- all_numbers |>
  dplyr::filter(block == "distribution_ks", sample == "full_sample",
                term == "ks_D") |>
  dplyr::transmute(model,
                   strict = paste0(fmt_num(estimate, 3), " (", fmt_p(p_value), ")"))
ks13_comp <- all_numbers |>
  dplyr::filter(block == "distribution_ks", sample == "compliant",
                term == "ks_D") |>
  dplyr::transmute(model,
                   comp = paste0(fmt_num(estimate, 3), " (", fmt_p(p_value), ")"))
ks13 <- dplyr::full_join(ks13_strict, ks13_comp, by = "model")

ks4_strict <- all_numbers |>
  dplyr::filter(block == "s4_distribution_ks", sample == "strict_n1272",
                term == "ks_D", model != "Overall") |>
  dplyr::transmute(model, strict = paste0(fmt_num(estimate, 3), " (", fmt_p(p_value), ")"))
ks4_comp <- all_numbers |>
  dplyr::filter(block == "s4_distribution_ks", sample == "compliant_n1073",
                term == "ks_D", model != "Overall") |>
  dplyr::transmute(model, comp = paste0(fmt_num(estimate, 3), " (", fmt_p(p_value), ")"))
ks4 <- dplyr::full_join(ks4_strict, ks4_comp, by = "model")

ks_tab <- dplyr::bind_rows(ks13, ks4) |>
  dplyr::mutate(model = factor(model, levels = c(
    "Jailbroken", "Standard", "Truth-Constrained",
    "Claude", "Gemini", "GPT-5.2", "Grok"))) |>
  dplyr::arrange(model) |>
  dplyr::select(`Model condition` = model,
                `KS D, strict (p)` = strict, `KS D, compliant (p)` = comp)

si_kable_styled(ks_tab, caption = "Two-sample Kolmogorov-Smirnov distance D between the bunking and debunking distributions of direction-aligned belief change, by model condition, with its p-value. Studies 1-3 (single analytic sample) appear in the strict column; Study 4 reports the per-model tests on the observed-outcome and compliant samples.")
```

```{r ra-large-shift}
shares <- all_numbers |>
  dplyr::filter(block == "large_shift_thresholds",
                grepl("^share_ge", term), !is.na(direction),
                !(model == "Pooled" & sample == "strict_n1272")) |>
  dplyr::mutate(thr = sub("share_ge", "", term),
                direction = ifelse(direction == "bunk", "bunk", "debunk")) |>
  dplyr::transmute(model, thr, direction, share = scales::percent(estimate, accuracy = 1)) |>
  tidyr::pivot_wider(names_from = direction, values_from = share)

tests <- all_numbers |>
  dplyr::filter(block == "large_shift_thresholds",
                grepl("^bunk_vs_debunk_chisq_ge", term),
                !(model == "Pooled" & sample == "strict_n1272")) |>
  dplyr::transmute(model, thr = sub("bunk_vs_debunk_chisq_ge", "", term),
                   `chi-sq` = fmt_num(estimate, 2), p = fmt_p(p_value))

large_shift <- dplyr::left_join(shares, tests, by = c("model", "thr")) |>
  dplyr::mutate(
    model = factor(model, levels = c("Jailbroken", "Standard", "Truth-Constrained",
                                     "Pooled", "Claude", "Gemini", "GPT-5.2", "Grok")),
    threshold = factor(paste0("≥", thr), levels = paste0("≥", c(20, 30, 40, 50)))) |>
  dplyr::arrange(model, threshold) |>
  dplyr::select(model, threshold, `bunk %` = bunk, `debunk %` = debunk, `chi-sq`, p)

si_kable_styled(large_shift, caption = "Share of participants with a direction-aligned belief shift of at least 20, 30, 40, and 50 scale points, by model condition and assigned direction, with chi-square tests of the bunking-versus-debunking difference at each threshold. Includes Studies 1-3 (with a pooled row) and the Study 4 strict sample.")
```

```{r ra-distribution-fig, fig.cap="Share of participants exceeding large-shift thresholds (>=20, >=30, >=40, >=50 points) by model condition and assigned direction, across all four studies."}
fig_distribution_thresholds()
```

Figure \@ref(fig:ra-belief-trajectory) complements the cell means with the full
shape of the effect: Panel A traces mean belief on the raw scale across the
pre-, post-, and debrief timepoints (so bunking visibly raises belief and
debunking lowers it, with the debrief largely reverting the bunking gain), and
Panel B gives, for each arm, the *survival* curve of direction-aligned change ---
the share of participants whose change exceeds each value $x$ --- so a higher curve
means more participants moved further toward the assigned position. Both the strict
and compliant samples are overlaid.

```{r ra-belief-trajectory, fig.width=9, fig.height=8, fig.cap="Belief trajectory and the distribution of belief change, all four studies, both samples. Panel A: mean belief (reverse-coded, higher = more belief) at the pre-, post-, and debrief timepoints, by arm (the Studies 1-3 debrief was administered only in the bunking arm; Study 4 collected it in both). Panel B: survival curve of direction-aligned belief change -- the share of participants whose change exceeds x, so a HIGHER curve means more participants moved further toward the assigned position (more effective). On the in-window analytic sample bunking and debunking are close; GPT-5.2 is the one model whose ordering reverses (debunking sits clearly above bunking, because GPT-5.2 bunking backfires). Solid lines = full/strict sample, dashed = compliant subsample."}
fig_belief_trajectory()
```

## Attrition

The analytic screening funnels are reported in Tables \@ref(tab:me-attention) and
\@ref(tab:me-funnel-s4) and Figure \@ref(fig:ra-funnel-s13). Table
\@ref(tab:ra-dropout-logistic) reports logistic regressions of post-conversation
dropout on assigned direction for Studies 1--3. The likelihood-ratio tests were
not statistically significant in any study.

```{r ra-dropout-logistic}
drop <- all_numbers |>
  dplyr::filter(block == "dropout_logistic_s13") |>
  dplyr::mutate(term = dplyr::recode(term,
    intercept_logit = "Intercept (log-odds)",
    debunk_vs_bunk_logit = "Debunk vs. bunk (log-odds)",
    differential_attrition_LRT = "Differential-attrition LRT",
    overall_dropout_rate = "Overall dropout rate")) |>
  dplyr::mutate(value = dplyr::case_when(
    term == "Overall dropout rate" ~ scales::percent(estimate, accuracy = 0.1),
    term == "Differential-attrition LRT" ~ fmt_p(p_value),
    TRUE ~ paste0(fmt_ci(estimate, conf_low, conf_high, 2), ", p = ", fmt_p(p_value)))) |>
  dplyr::select(model, term, value) |>
  tidyr::pivot_wider(names_from = model, values_from = value)

si_kable_styled(drop, caption = "Studies 1-3 differential-attrition tests: logistic regression of post-conversation dropout on assigned direction, by GPT-4o variant. The LRT row reports the likelihood-ratio p-value for direction-dependent attrition; other rows report estimate [95% CI], p.")
```

For Study 4, `r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="pool"))`
participants satisfied the analytic eligibility criteria and reached the transition
into the experimental conversation. Of these,
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="completed"))`
completed the primary post-conversation outcomes and
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="pool") - est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="completed"))`
did not.

Incomplete records were classified from the joined 19,000-character chunks of
`__js_chatPartialData1` and, where available, the complete chat-history fields.
The parser excluded the seeded system prompt and prior participant writing. A
stored nonempty assistant message defined confirmed initial delivery.

Among the outcome-incomplete participants,
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="initial_ai_delivered_loss_n"))`
had a stored initial AI response. Of these,
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="initial_ai_no_user_n"))`
received the initial response but did not enter a new message,
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="interactive_chat_then_lost_n"))`
began an interactive conversation but left before completing it, and
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="chat_completed_outcomes_missing_n"))`
completed the conversation but did not provide the primary outcomes.
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="last_user_without_reply_n"))`
of the
interactive records ended after a participant turn without a subsequent nonempty
assistant reply; the other interactive records contain a later assistant response.

The remaining
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="no_successful_callback_n"))`
records contained no successful initial chatbot callback.
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="no_callback_instruction_page_submitted_n"))`
of these participants submitted the immediately preceding page explaining the
chatbot task. The stored data do not distinguish participant departure from
iframe, API, or provider failure.

The survey requested the OpenRouter endpoint
`google/gemini-3-pro-preview`, which was removed from OpenRouter during fielding.
From March 28--30,
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="gemini3_blackout_pool_n"))`
eligible Gemini 3 assignments produced
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="gemini3_blackout_callback_n"))`
stored callbacks and
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="gemini3_blackout_completed_n"))`
completed outcomes, while the other models continued operating. After the requested
model was changed to `google/gemini-3.1-pro-preview`,
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="gemini31_completed_n"))`
of `r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="gemini31_pool_n"))`
Gemini assignments completed (`r fmt_pct(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="gemini31_completion_rate"), 1)`).
Applying the non-Gemini no-callback rate to the Gemini pool yielded approximately
`r fmt_n(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="gemini_excess_no_callback_vs_others_n"))`
excess Gemini no-callback records.

```{r ra-attrition-split}
.lab <- c(
  pool = "Model-assigned analytic pool",
  completed = "Completed primary outcomes",
  no_successful_callback_n = "No successful initial AI callback",
  initial_ai_no_user_n = "Initial AI delivered; no new participant turn",
  interactive_chat_then_lost_n = "Interactive chat, then lost",
  chat_completed_outcomes_missing_n = "Chat completed; outcomes missing",
  postdelivery_attrition_rate = "Postdelivery outcome attrition rate"
)
att <- all_numbers |>
  dplyr::filter(block == "attrition_substantive_vs_technical",
                model %in% c("Claude", "Gemini", "GPT-5.2", "Grok"),
                term %in% names(.lab)) |>
  dplyr::mutate(Stage = factor(.lab[term], levels = .lab),
                model = factor(model, levels = c("Claude", "Gemini", "GPT-5.2", "Grok")),
                value = ifelse(term == "postdelivery_attrition_rate",
                               scales::percent(estimate, accuracy = 0.1), fmt_num(estimate, 0))) |>
  dplyr::arrange(Stage, model) |>
  dplyr::select(Stage, model, value) |>
  tidyr::pivot_wider(names_from = model, values_from = value)

si_kable_styled(att, caption = "Study 4 outcome status after chatbot assignment. 'No successful initial AI callback' denotes no stored nonempty assistant response. Postdelivery attrition is calculated among outcome completers plus incomplete participants with a stored initial AI response.")
```

Among the
`r fmt_n(nn("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="postdelivery_attrition_by_direction_LRT"))`
participants with confirmed initial delivery, outcome attrition was
`r fmt_pct(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, direction="bunk", term="postdelivery_attrition_rate"), 1)`
in the bunking condition and
`r fmt_pct(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, direction="debunk", term="postdelivery_attrition_rate"), 1)`
in the debunking condition. This difference was statistically significant by the
likelihood-ratio test (*p* =
`r fmt_p(pv("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="postdelivery_attrition_by_direction_LRT"))`)
and remained significant by Fisher's exact test (*p* =
`r fmt_p(pv("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="postdelivery_attrition_by_direction_Fisher"))`).
Postdelivery attrition did not differ across model categories (*p* =
`r fmt_p(pv("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, term="postdelivery_attrition_by_model_LRT"))`).

```{r ra-attrition-lrt}
att_lrt <- all_numbers |>
  dplyr::filter(block == "attrition_substantive_vs_technical", is.na(model),
                term %in% c("differential_attrition_by_direction_LRT",
                            "differential_attrition_by_model_LRT",
                            "postdelivery_attrition_by_direction_LRT",
                            "postdelivery_attrition_by_model_LRT")) |>
  dplyr::mutate(
    Outcome = ifelse(grepl("^postdelivery", term),
                     "Missing outcome after confirmed initial AI delivery",
                     "Overall incomplete outcome"),
    test = dplyr::recode(term,
      differential_attrition_by_direction_LRT = "By assigned direction",
      differential_attrition_by_model_LRT = "By model category",
      postdelivery_attrition_by_direction_LRT = "By assigned direction",
      postdelivery_attrition_by_model_LRT = "By model category"),
    Outcome = factor(Outcome, levels = c("Overall incomplete outcome",
                                         "Missing outcome after confirmed initial AI delivery"))) |>
  dplyr::arrange(Outcome, test) |>
  dplyr::select(Outcome, Test = test, n, p = p_value) |>
  dplyr::mutate(p = fmt_p(p))

si_kable_styled(att_lrt, caption = "Study 4 likelihood-ratio tests for differential outcome attrition. Overall incompleteness includes the no-callback state. The confirmed-delivery analysis includes all outcome completers and incomplete participants with a stored initial AI response.")
```

Among participants with complete primary outcomes, median total survey duration was
`r fmt_num(est("attrition_burden_context", model="Study 4", term="completed_total_duration_median_min"), 1)`
minutes in Study 4, compared with
`r fmt_num(est("attrition_burden_context", model="Jailbroken", term="completed_total_duration_median_min"), 1)`,
`r fmt_num(est("attrition_burden_context", model="Standard", term="completed_total_duration_median_min"), 1)`, and
`r fmt_num(est("attrition_burden_context", model="Truth-Constrained", term="completed_total_duration_median_min"), 1)`
minutes in Studies 1--3. Study 4 included open-ended writing, the
pre-conversation social-media paradigm, and the chatbot task.
Postdelivery outcome attrition was
`r fmt_pct(est("attrition_substantive_vs_technical", sample="strict_n1272", model=NA, direction=NA, term="postdelivery_attrition_rate"), 1)`
in Study 4 and
`r fmt_pct(est("attrition_substantive_vs_technical", sample="full_sample", model="Jailbroken", term="postdelivery_attrition_rate"), 1)`,
`r fmt_pct(est("attrition_substantive_vs_technical", sample="full_sample", model="Standard", term="postdelivery_attrition_rate"), 1)`, and
`r fmt_pct(est("attrition_substantive_vs_technical", sample="full_sample", model="Truth-Constrained", term="postdelivery_attrition_rate"), 1)`
in Studies 1--3, respectively.

```{r ra-attrition-burden}
burden <- all_numbers |>
  dplyr::filter(block == "attrition_burden_context") |>
  dplyr::mutate(
    Study = dplyr::recode(model,
      Jailbroken = "Study 1", Standard = "Study 2",
      `Truth-Constrained` = "Study 3", `Study 4` = "Study 4"),
    measure = dplyr::recode(term,
      completed_total_duration_median_min = "Median",
      completed_total_duration_p90_min = "90th percentile"),
    value = fmt_num(estimate, 1)
  ) |>
  dplyr::select(Study, `Outcome completers` = n, measure, value) |>
  tidyr::pivot_wider(names_from = measure, values_from = value) |>
  dplyr::arrange(Study)

si_kable_styled(burden, caption = "Total Qualtrics duration in minutes among participants with complete primary outcomes.")
```

### Missing-outcome analyses among confirmed chat starters

The confirmed-chat-start cohort contains 1,353 participants: the 1,272 participants
with complete primary outcomes and 81 otherwise incomplete records containing a
stored initial assistant response. Six participants who did not complete the full
primary-outcome battery nevertheless supplied the focal post-belief rating. The
belief-specific observed-outcome reference therefore uses all 1,278 available
ratings, with their restatement orientation corrected in the same way as in the
complete-outcome sample; the remaining 75 belief outcomes are missing. The
pre-conversation posts for all 81 records were audited with three orientation
raters and five stance raters. Table \@ref(tab:ra-s4-chat-started) reports the
eight model-by-direction aligned belief effects under available outcomes, zero
change for missing outcomes, MAR multiple imputation, and 0--100-scale
identification bounds.

```{r ra-s4-chat-started}
s4b_cells <- all_numbers |>
  dplyr::filter(block == "s4_chat_started_cell_effects",
                sample == "chat_started_n1353",
                outcome == "aligned_belief_change") |>
  dplyr::mutate(
    Approach = dplyr::recode(term,
      available_outcomes = "Available outcomes",
      zero_change = "Zero change if missing",
      mi_mar = "MAR multiple imputation",
      support_bounds = "Feasible-value bounds"),
    value = dplyr::if_else(term == "support_bounds",
      paste0("[", fmt_num(conf_low, 1), ", ", fmt_num(conf_high, 1), "]"),
      fmt_ci(estimate, conf_low, conf_high, 1)
    ),
    .m = match(model, c("Claude", "Gemini", "GPT-5.2", "Grok")),
    Direction = dplyr::recode(direction, bunk = "Bunk", debunk = "Debunk")
  ) |>
  dplyr::arrange(.m, Direction) |>
  dplyr::group_by(model, Direction) |>
  dplyr::mutate(`Available N` = n[term == "available_outcomes"][1]) |>
  dplyr::ungroup() |>
  dplyr::select(Model = model, Direction, `Available N`, Approach, value) |>
  tidyr::pivot_wider(names_from = Approach, values_from = value)

si_kable_styled(
  s4b_cells,
  caption = "Direction-aligned belief change by Study 4 model and direction among 1,353 confirmed chat starters. Point estimates are reported with 95% CIs; feasible-value bounds are identification intervals. Available-outcome N = 1,278."
)
```

The GPT-5.2 bunking estimates and feasible-value interval were below zero. The
feasible-value intervals for the other seven model-by-direction cells were above
zero.

```{r ra-s4-chat-started-model-asymmetry}
s4b_asym <- all_numbers |>
  dplyr::filter(block == "s4_chat_started_model_contrasts",
                sample == "chat_started_n1353",
                outcome == "aligned_belief_change") |>
  dplyr::mutate(
    Approach = dplyr::recode(term,
      available_outcomes = "Available outcomes",
      zero_change = "Zero change if missing",
      mi_mar = "MAR multiple imputation",
      support_bounds = "Feasible-value bounds"),
    value = dplyr::if_else(term == "support_bounds",
      paste0("[", fmt_num(conf_low, 1), ", ", fmt_num(conf_high, 1), "]"),
      paste0(fmt_ci(estimate, conf_low, conf_high, 1), "; p ", fmt_p(p_value))),
    .m = match(model, c("Claude", "Gemini", "GPT-5.2", "Grok"))
  ) |>
  dplyr::arrange(.m) |>
  dplyr::select(Model = model, Approach, value) |>
  tidyr::pivot_wider(names_from = Approach, values_from = value)

si_kable_styled(
  s4b_asym,
  caption = "Within-model debunk-minus-bunk asymmetry in direction-aligned belief change among confirmed chat starters. Point rows give estimate [95% CI] and p; positive values favor debunking. Feasible-value endpoints are identification bounds, not confidence limits."
)
```

The debunk-minus-bunk estimates were negative for Claude, positive for Gemini and
GPT-5.2, and not statistically significant for Grok. The feasible-value interval
included zero for Grok and excluded zero for the other three models. Table
\@ref(tab:ra-s4-chat-started-registered-belief) reports the equal-model average
and direction-by-model omnibus test.

```{r ra-s4-chat-started-registered-belief}
s4b_reg <- all_numbers |>
  dplyr::filter(block == "s4_chat_started_registered_tests",
                sample == "chat_started_n1353",
                outcome == "aligned_belief_change") |>
  dplyr::mutate(
    Approach = dplyr::case_when(
      grepl("^available_outcomes", term) ~ "Available outcomes",
      grepl("^zero_change", term) ~ "Zero change if missing",
      grepl("^mi_mar", term) ~ "MAR multiple imputation",
      TRUE ~ "Feasible-value bounds"),
    Result = dplyr::case_when(
      direction == "direction-x-model" ~ paste0("omnibus p ", fmt_p(p_value)),
      grepl("^support_bounds", term) ~ paste0("average [", fmt_num(conf_low, 1), ", ", fmt_num(conf_high, 1), "]"),
      TRUE ~ paste0("average ", fmt_ci(estimate, conf_low, conf_high, 1), "; p ", fmt_p(p_value))
    )
  ) |>
  dplyr::select(Approach, direction, Result) |>
  tidyr::pivot_wider(names_from = direction, values_from = Result, values_fill = "---") |>
  dplyr::rename(`Equal-model debunk - bunk` = `debunk-minus-bunk`,
                `Direction x model` = `direction-x-model`)
si_kable_styled(
  s4b_reg,
  caption = "Registered belief asymmetry among confirmed chat starters: the equal-model debunk-minus-bunk contrast and direction-by-model omnibus test. The MAR omnibus is a Rubin-pooled HC3 multivariate Wald test; available and zero-change omnibus p-values are HC3 F tests. Bound endpoints are not confidence limits."
)
```

The equal-model debunk-minus-bunk estimate was positive under each point-estimation
approach; its feasible-value interval excluded zero. Direction-by-model omnibus
tests had *p* < .001 under available outcomes, zero change, and MAR imputation.
Section \@ref(chat-started-speech) reports the public-speech results. Analyses are
conditional on confirmed chat start.

```{r ra-attrition-fig, fig.pos='H', fig.cap="Outcome-incomplete states by model condition across all four studies, as a percentage of each model-assigned pool. 'No successful initial AI callback' indicates that no nonempty assistant response was stored. The requested Study 4 Gemini endpoint was removed from OpenRouter during fielding."}
fig_attrition()
```

## Engagement

Conversation engagement is summarized from the transcripts so that exposure can
be compared across model conditions (Table \@ref(tab:ra-engagement)).

```{r ra-engagement}
eng <- all_numbers |>
  dplyr::filter(block == "engagement") |>
  dplyr::mutate(direction = stringr::str_to_title(direction)) |>
  dplyr::mutate(value = dplyr::case_when(
    term == "pct_ended_before_10turn_max" ~ scales::percent(estimate, accuracy = 1),
    term %in% c("mean_user_messages", "mean_assistant_messages") ~ fmt_num(estimate, 2),
    TRUE ~ fmt_ci(estimate, conf_low, conf_high, 0))) |>
  dplyr::mutate(term = dplyr::recode(term,
    mean_user_messages = "User msgs",
    mean_assistant_messages = "Asst msgs",
    mean_user_words = "User words [95% CI]",
    mean_assistant_words = "Asst words [95% CI]",
    pct_ended_before_10turn_max = "% ended early")) |>
  dplyr::mutate(
    study = ifelse(model %in% c("Claude", "Gemini", "GPT-5.2", "Grok"), "Study 4", "Studies 1-3"),
    model = factor(model, levels = c("Jailbroken", "Standard", "Truth-Constrained",
                                     "Claude", "Gemini", "GPT-5.2", "Grok")),
    term = factor(term, levels = c("User msgs", "Asst msgs", "User words [95% CI]",
                                   "Asst words [95% CI]", "% ended early"))) |>
  dplyr::arrange(model, direction, term) |>
  dplyr::select(study, model, direction, n, term, value) |>
  tidyr::pivot_wider(names_from = term, values_from = value)

si_kable_styled(eng, caption = "Conversation engagement by model condition and assigned direction across all four studies: mean user and assistant message counts, mean word counts (with 95% CIs), and the share of conversations ending before the turn cap.")
```

Beyond cell means, Figure \@ref(fig:ra-conv-length) shows the full distribution of
conversation length --- the number of participant messages per conversation --- by
study and arm. The distributions are similar across arms within each study and
concentrate at low message counts: only a small minority of conversations (1-7%
per cell) reach the 10-user-message cap, confirming that exposure is not
systematically different between bunking and debunking.

```{r ra-conv-length, fig.width=9, fig.height=6, fig.cap="Distribution of the number of participant (user) messages per conversation, by study/model and assigned direction (strict sample; the compliant subsample is near-identical). The dotted line marks the 10-user-message cap; 93-99% of conversations end before reaching it."}
fig_conversation_length()
```

## Compliance

The same first-turn attempt-to-persuade
(APE) instrument scores all four studies: Studies 1-3 report the attempt rate
(Table \@ref(tab:ra-compliance-s13)), and Study 4 additionally reports the
refusal rate and the strict-compliance rate (an attempt with no refusal) per cell
(Table \@ref(tab:ra-compliance-s4)). The attempt rate is near ceiling everywhere
except Truth-Constrained GPT-4o bunking and GPT-5.2 bunking, where the model
frequently declined the assigned task.

```{r ra-compliance-s13}
comp13 <- all_numbers |>
  dplyr::filter(block == "compliance", term == "ape_attempt_rate") |>
  dplyr::transmute(model, direction, n,
                   `attempt rate` = scales::percent(estimate, accuracy = 1)) |>
  tidyr::pivot_wider(names_from = direction, values_from = c(n, `attempt rate`)) |>
  dplyr::select(model, `n (bunk)` = n_Bunking, `bunk attempt` = `attempt rate_Bunking`,
                `n (debunk)` = n_Debunking, `debunk attempt` = `attempt rate_Debunking`)

si_kable_styled(comp13, caption = "Studies 1-3 APE compliance: first-turn attempt rate by GPT-4o variant and assigned direction.")
```

```{r ra-compliance-s4}
comp4 <- all_numbers |>
  dplyr::filter(block == "compliance_cells", sample == "strict_n1272",
                term %in% c("attempt_rate", "refusal_rate", "strict_compliance_rate")) |>
  dplyr::mutate(value = scales::percent(estimate, accuracy = 1),
                term = dplyr::recode(term,
                  attempt_rate = "Attempt", refusal_rate = "Refusal",
                  strict_compliance_rate = "Strict compliant")) |>
  dplyr::select(model, direction, n, term, value) |>
  tidyr::pivot_wider(names_from = term, values_from = value)

si_kable_styled(comp4, caption = "Study 4 first-turn compliance by model and assigned direction (strict sample): attempt, refusal, and strict-compliance rates.")
```

Collapsing over direction makes the model-level differences in compliance easier
to read. Table \@ref(tab:ra-compliance-model-overall) reports each Study 4
model's overall first-turn attempt, refusal, and strict-compliance rates,
marginal over the bunking and debunking arms, on the strict sample. Three of the
four models comply with the assigned task almost always; GPT-5.2 is the clear
exception, attempting the assigned persuasion in only about
`r fmt_pct(est("compliance_model_overall", model="GPT-5.2", term="attempt_rate"), 0)`
of conversations and refusing in roughly
`r fmt_pct(est("compliance_model_overall", model="GPT-5.2", term="refusal_rate"), 0)`,
which is why the strict and compliant Study 4 estimates diverge mainly through
this one model.

```{r ra-compliance-model-overall}
cmo <- all_numbers |>
  dplyr::filter(block == "compliance_model_overall", sample == "strict_n1272") |>
  dplyr::mutate(term = dplyr::recode(term,
                  attempt_rate = "Attempt", refusal_rate = "Refusal",
                  strict_compliance_rate = "Strict compliant"),
                value = scales::percent(estimate, accuracy = 1),
                model = factor(model, levels = c("Claude", "Gemini", "GPT-5.2", "Grok"))) |>
  dplyr::arrange(model) |>
  dplyr::select(model, n, term, value) |>
  tidyr::pivot_wider(names_from = term, values_from = value) |>
  dplyr::select(model, n, Attempt, Refusal, `Strict compliant`)

si_kable_styled(cmo, caption = "Study 4 first-turn compliance by model, collapsed over assigned direction (strict sample): overall attempt, refusal, and strict-compliance rates. GPT-5.2 is the lone low-compliance model.")
```

```{r ra-compliance-fig, fig.cap="First-turn attempt and strict-compliance rates by model condition and assigned direction across all four studies."}
fig_compliance_4study()
```

Restricting to the compliant subsample raises the Truth-Constrained and GPT-5.2
bunking means — most steeply GPT-5.2, from a strong backfire to a near-null
effect — because the participants who drove the negative strict-sample mean were
exactly those whose model declined to bunk (the strict and compliant cell means
sit side by side in Table \@ref(tab:ra-belief-cells)).

### What Non-Compliant Models Did Instead

A first-turn attempt rate below ceiling raises the deployment-relevant question of
what the model did when it declined the assigned task. The compliance classifier
distinguishes two ways of not attempting: an **explicit refusal** (the model
declines without taking a side) and a **counter-argument** (the model actively
argues *against* the position it was assigned to advocate). The distinction
matters because the two have opposite consequences for belief: a refusal leaves the
participant where they started, whereas a counter-argument is itself a persuasion
attempt in the reverse direction. Table \@ref(tab:ra-nonattempt-split) breaks the
Study 4 non-attempts down this way for each model and direction, with the aligned
belief change among the counter-argument conversations.

The pattern is asymmetric and points one way, and it is overwhelmingly driven by
GPT-5.2, the lone low-compliance model. In the **bunking** arm, when GPT-5.2
declined to spread the conspiracy it usually did not merely refuse: it
counter-argued in
`r fmt_n(est("nonattempt_split_by_model", sample="strict_n1272", model="GPT-5.2", direction="Bunking", term="n_counterargument"))`
of its
`r fmt_n(est("nonattempt_split_by_model", sample="strict_n1272", model="GPT-5.2", direction="Bunking", term="n_nonattempt"))`
bunk non-attempts, and those conversations moved aligned belief about
`r fmt_est_ci(num("nonattempt_split_by_model", sample="strict_n1272", model="GPT-5.2", direction="Bunking", term="mean_aligned_change_counterargument"), 1, signed = TRUE)`
points --- a large shift *against* the assigned bunking direction, i.e. toward
rejecting the conspiracy. The other three models declined far less often (5, 9,
and 10 bunk non-attempts for Claude, Gemini, and Grok respectively), and only Grok
refused outright rather than counter-arguing. The **debunking** arm shows no such
behavior in any model: GPT-5.2's
`r fmt_n(est("nonattempt_split_by_model", sample="strict_n1272", model="GPT-5.2", direction="Debunking", term="n_nonattempt"))`
debunking non-attempts were all explicit refusals, with
`r fmt_n(est("nonattempt_split_by_model", sample="strict_n1272", model="GPT-5.2", direction="Debunking", term="n_counterargument"))`
counter-arguments, and the remaining models likewise never argued *for* a
conspiracy they had been asked to debunk. The flip side of compliance is therefore
not inertia but
inversion: a safety-trained model instructed to spread a conspiracy frequently
debunks it instead, which is why the strict-sample bunking mean for GPT-5.2 is so
much more negative than its compliant-sample mean.

The same picture holds when the GPT-5.2 bunking cell is split into attempts versus
declines: among the
`r fmt_n(nn("gpt52_bunk_decomposition", sample="strict_n1272", outcome="aligned_belief_change", term="attempted"))`
conversations where it *attempted* the bunk, belief change is small and not
distinguishable from zero
(`r fmt_est_ci(num("gpt52_bunk_decomposition", sample="strict_n1272", outcome="aligned_belief_change", term="attempted"), 2, signed = TRUE)`),
and the public post still moved *against* the conspiracy in both the attempted
(`r fmt_est_ci(num("gpt52_bunk_decomposition", sample="strict_n1272", outcome="aligned_stance_change", term="attempted"), 2, signed = TRUE)`)
and declined
(`r fmt_est_ci(num("gpt52_bunk_decomposition", sample="strict_n1272", outcome="aligned_stance_change", term="declined"), 2, signed = TRUE)`)
subsets.

```{r ra-nonattempt-split}
.nlab <- c(n_nonattempt = "Non-attempts",
           n_explicit_refusal = "Explicit refusals",
           n_counterargument = "Counter-arguments",
           share_counterargument = "Counter-arg. share",
           mean_aligned_change_counterargument = "Counter-arg. aligned change [95% CI]")
nas <- all_numbers |>
  dplyr::filter(block == "nonattempt_split_by_model", sample == "strict_n1272",
                model != "All models") |>
  dplyr::mutate(value = dplyr::case_when(
    term == "share_counterargument" ~ scales::percent(estimate, accuracy = 1),
    term == "mean_aligned_change_counterargument" & !is.na(conf_low) ~
      fmt_ci(estimate, conf_low, conf_high, 1),
    term == "mean_aligned_change_counterargument" ~ "—",
    TRUE ~ fmt_num(estimate, 0)),
    Quantity = factor(.nlab[term], levels = .nlab),
    model = factor(model, levels = c("Claude", "Gemini", "GPT-5.2", "Grok")),
    direction = factor(direction, levels = c("Bunking", "Debunking"))) |>
  dplyr::arrange(direction, model, Quantity) |>
  dplyr::select(Direction = direction, Model = model, Quantity, value) |>
  tidyr::pivot_wider(names_from = Quantity, values_from = value)

si_kable_styled(nas, caption = "Study 4 non-attempts split by type, for each model and assigned direction (strict sample; compliance-scored conversations that did not attempt the assigned direction). 'Counter-arguments' are conversations in which the model argued against the position it was assigned to advocate; 'Counter-arg. aligned change' is the mean direction-aligned belief change among those conversations, with a 95% CI where n is adequate. In the bunking arm most non-attempts are counter-arguments that move belief sharply against the assigned direction; in the debunking arm every non-attempt is an explicit refusal.")
```

## The Study 3 Truth-Constraint Counterfactual

The Truth-Constrained variant asks how much of GPT-4o's bunking power survives
when the model may use only accurate arguments. Table
\@ref(tab:ra-s3-counterfactual) gives the truth-constraint penalty two ways: the
compliant bunking-minus-debunking contrast on the matched $n = 740$ counterfactual
set, and the belief change in the Truth-Constrained bunking conversations that
were non-compliant (the model declined to fabricate). The penalty survives among
compliant conversations, and the non-compliant cases moved belief sharply
*against* the conspiracy.

```{r ra-s3-counterfactual}
cf <- all_numbers |>
  dplyr::filter(block == "s3_counterfactual") |>
  dplyr::transmute(quantity = "Compliant bunk - debunk (matched counterfactual)", n,
                   `estimate [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

nc <- all_numbers |>
  dplyr::filter(block == "noncompliant_s3_bunk") |>
  dplyr::transmute(quantity = "Belief change in non-compliant TC bunking conversations", n,
                   `estimate [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

si_kable_styled(dplyr::bind_rows(cf, nc), caption = "Study 3 truth-constraint counterfactual: the compliant bunking-minus-debunking contrast on the matched counterfactual set, and the belief change in Truth-Constrained bunking conversations where the model declined to comply.")
```

## Secondary Outcomes

Three families of secondary outcomes are reported. The first is generalized
conspiracist belief (GCBS), administered pre and post in Studies 1-3; Table
\@ref(tab:ra-gcbs) gives the pre-to-post change per cell and the
bunking-minus-debunking contrast. The second is trust in AI, measured before and
after the conversation in all four studies (`genai_trust` pre, `trust2` post;
Study 4 additionally re-administered the item after the corrective debrief,
`trust_debrf`): Studies 1-3 report the pre-to-post change per variant (Table
\@ref(tab:ra-trust)), and Study 4's pre-to-post change — computed on the same
footing — joins the cross-study table below alongside the registered
post-treatment contrast. The third is the set of perception
measures — argument strength, provided-new-information, collaborativeness, and
impartiality — characterizing *how* the conversation was experienced (Table
\@ref(tab:ra-perception)).

```{r ra-gcbs}
gcbs_cells <- all_numbers |>
  dplyr::filter(block == "gcbs_change", term == "GCBS_change_mean") |>
  dplyr::transmute(model, direction, n,
                   ci = fmt_ci(estimate, conf_low, conf_high, 2)) |>
  tidyr::pivot_wider(names_from = direction, values_from = c(n, ci))

gcbs_con <- all_numbers |>
  dplyr::filter(block == "gcbs_change", term == "bunk_minus_debunk") |>
  dplyr::transmute(model, `bunk - debunk [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

gcbs <- dplyr::left_join(gcbs_cells, gcbs_con, by = "model") |>
  dplyr::select(model, `n (bunk)` = n_Bunking, `bunk [95% CI]` = ci_Bunking,
                `n (debunk)` = n_Debunking, `debunk [95% CI]` = ci_Debunking,
                `bunk - debunk [95% CI]`, p)

si_kable_styled(gcbs, caption = "Studies 1-3 generalized conspiracist belief (GCBS): pre-to-post change by assigned direction and the within-variant bunking-minus-debunking contrast.")
```

```{r ra-trust}
trust_cells <- all_numbers |>
  dplyr::filter(block == "trust_change", term == "trust_pre_post_change_mean") |>
  dplyr::transmute(model, direction, ci = fmt_ci(estimate, conf_low, conf_high, 2)) |>
  tidyr::pivot_wider(names_from = direction, values_from = ci)

trust_con <- all_numbers |>
  dplyr::filter(block == "trust_change", term == "bunk_minus_debunk") |>
  dplyr::transmute(model, n, `bunk - debunk [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

trust <- dplyr::left_join(trust_cells, trust_con, by = "model") |>
  dplyr::select(model, `bunk [95% CI]` = Bunking, `debunk [95% CI]` = Debunking,
                n, `bunk - debunk [95% CI]`, p)

si_kable_styled(trust, caption = "Studies 1-3 trust in AI: pre-to-post change by assigned direction and the within-variant bunking-minus-debunking contrast. A positive bunking trust effect that survives the corrective disclosure indicates trust is not undercut by exposure to fabricated arguments.")
```

Table \@ref(tab:ra-trust-allstudies) places the Studies 1-3 per-variant
bunking-minus-debunking trust contrast (pre-to-post change) alongside two Study 4
rows: the same pre-to-post change contrast (strict sample, models pooled) and the
registered contrast on the post-treatment trust item, the latter on both the
observed-outcome and compliant samples. The contrast is small in
Studies 1-3 — significant only for the Jailbroken variant — with the Study 4
rows shown on both the change and post-only footings. Study 4 also re-administered the
trust item after the corrective debrief (`trust_debrf`); the post-to-debrief
trust change is reported in block `s4_trust_change`, and the post-debrief belief
trajectory is reported in the Durability section below (Table
\@ref(tab:rb-debrief-s4) and Table \@ref(tab:rb-debrief-traj)).

```{r ra-trust-allstudies}
trust13 <- all_numbers |>
  dplyr::filter(block == "trust_change", term == "bunk_minus_debunk") |>
  dplyr::transmute(study = model, sample = "Pre-to-post change", n,
                   `bunk - debunk [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

trust4chg <- all_numbers |>
  dplyr::filter(block == "s4_trust_change", term == "bunk_minus_debunk",
                outcome == "Trust in AI (pre->post)", sample == "strict_n1272") |>
  dplyr::transmute(study = "Study 4", sample = "Pre-to-post change (strict)",
                   n, `bunk - debunk [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

trust4 <- all_numbers |>
  dplyr::filter(block == "secondary_contrasts", outcome == "trust2",
                sample %in% c("strict_n1272", "compliant_n1073")) |>
  dplyr::transmute(study = "Study 4",
                   sample = ifelse(sample == "strict_n1272",
                                   "Post-treatment (strict)", "Post-treatment (compliant)"),
                   n, `bunk - debunk [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                   p = fmt_p(p_value))

trust_all <- dplyr::bind_rows(trust13, trust4chg, trust4) |>
  dplyr::mutate(study = factor(study, levels = c("Jailbroken", "Standard",
                                                 "Truth-Constrained", "Study 4"))) |>
  dplyr::arrange(study) |>
  dplyr::select(study, sample, n, `bunk - debunk [95% CI]`, p)

si_kable_styled(trust_all, caption = "Treatment effect on trust in AI across all four studies. Studies 1-3 report the within-variant bunking-minus-debunking contrast in pre-to-post trust change; Study 4 reports the same pre-to-post change contrast (strict sample, models pooled) plus the registered contrast on the post-treatment trust item in both the observed-outcome and compliant samples. The effect is small in Studies 1-3.")
```

The perception contrasts appear by item for all four studies, with the strict
and compliant samples in adjacent columns (Table \@ref(tab:ra-perception)).
The debunking arm is perceived as supplying more new information than the bunking
arm in every study and in both Study 4 samples; the other three dimensions move
little. Figure \@ref(fig:ra-secondary-fig) renders the same contrasts, with each
Studies 1-3 variant shown separately and Study 4 shown as the across-model
perception contrast (the per-model breakdown is in Table
\@ref(tab:ra-perception-bymodel)).

```{r ra-perception}
em <- function(x) ifelse(is.na(x), "—", x)
percep13_strict <- all_numbers |>
  dplyr::filter(block == "perception_mediators", term == "bunk_minus_debunk",
                sample == "full_sample") |>
  dplyr::transmute(study = model, item = outcome, n_strict = as.character(n),
                   strict = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_strict = fmt_p(p_value))

percep13_comp <- all_numbers |>
  dplyr::filter(block == "perception_mediators", term == "bunk_minus_debunk",
                sample == "compliant") |>
  dplyr::transmute(study = model, item = outcome, n_comp = as.character(n),
                   comp = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_comp = fmt_p(p_value))

percep13 <- dplyr::left_join(percep13_strict, percep13_comp, by = c("study", "item"))

s4_map <- c(ArgStrength = "Argument strength", new_info_rc = "Provided new information",
            Collaborative_rc = "Collaborativeness", Unbiased_rc = "Impartiality (unbiased)")
percep4_strict <- all_numbers |>
  dplyr::filter(block == "secondary_contrasts", sample == "strict_n1272",
                outcome %in% names(s4_map)) |>
  dplyr::transmute(study = "Study 4", item = unname(s4_map[outcome]), n_strict = as.character(n),
                   strict = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_strict = fmt_p(p_value))

percep4_comp <- all_numbers |>
  dplyr::filter(block == "secondary_contrasts", sample == "compliant_n1073",
                outcome %in% names(s4_map)) |>
  dplyr::transmute(study = "Study 4", item = unname(s4_map[outcome]), n_comp = as.character(n),
                   comp = fmt_ci(estimate, conf_low, conf_high, 2),
                   p_comp = fmt_p(p_value))

percep4 <- dplyr::left_join(percep4_strict, percep4_comp, by = c("study", "item"))

percep <- dplyr::bind_rows(percep13, percep4) |>
  dplyr::mutate(study = factor(study, levels = c("Jailbroken", "Standard",
                                                 "Truth-Constrained", "Study 4")),
                dplyr::across(c(n_comp, comp, p_comp), em)) |>
  dplyr::arrange(item, study) |>
  dplyr::select(item, study,
                `n (strict)` = n_strict, `bunk - debunk, strict [95% CI]` = strict, `p` = p_strict,
                `n (compliant)` = n_comp, `bunk - debunk, compliant [95% CI]` = comp, `p ` = p_comp)

si_kable_styled(percep, caption = "Perception measures across all four studies: bunking-minus-debunking contrasts in perceived argument strength, provision of new information, collaborativeness, and impartiality. All four studies report both the observed-outcome and compliant samples; Studies 1-3 use the pre-to-post analytic sample and its compliant subset, and Study 4 uses the strict and compliant post-treatment contrasts.")
```

```{r ra-secondary-fig, fig.cap="Bunking-minus-debunking contrasts on the perception measures (argument strength, new information, collaborativeness, impartiality) across all four studies, with 95 percent confidence intervals. Studies 1-3 are shown by GPT-4o variant and Study 4 by individual frontier model (no pooling), each on BOTH the observed-outcome and compliant samples (solid points/lines = Strict; hollow points/dashed lines = Compliant)."}
fig_secondary()
```

The Study 4 perception contrasts are also broken out by model in Table
\@ref(tab:ra-perception-bymodel). The model-level pattern is muted and mostly
non-significant --- the clearest signals are that Claude and Grok were rated as
providing more new information under bunking --- consistent with the small pooled
perception contrasts.

```{r ra-perception-bymodel}
pbm <- all_numbers |>
  dplyr::filter(block == "perception_s4_by_model",
                sample %in% c("strict_n1272", "compliant_n1073")) |>
  dplyr::mutate(samp = dplyr::recode(sample, strict_n1272 = "strict", compliant_n1073 = "comp"),
                Model = factor(model, levels = c("Claude", "Gemini", "GPT-5.2", "Grok")),
                ci = fmt_ci(estimate, conf_low, conf_high, 2),
                p = fmt_p(p_value)) |>
  dplyr::select(Model, Item = term, samp, n, ci, p) |>
  tidyr::pivot_wider(names_from = samp, values_from = c(n, ci, p)) |>
  dplyr::arrange(Model, Item) |>
  dplyr::select(Model, Item,
                `n (strict)` = n_strict, `bunk - debunk, strict [95% CI]` = ci_strict, `p` = p_strict,
                `n (compliant)` = n_comp, `bunk - debunk, compliant [95% CI]` = ci_comp, `p ` = p_comp)
si_kable_styled(pbm, caption = "Study 4 perception measures by model: bunking-minus-debunking contrast on each item (raw response scale, HC3), reported on both the observed-outcome and compliant samples. Positive = rated higher under bunking.", digits = 2)
```

### Pre-Treatment AI-Study Experience (Study 4)

Before the conversation, Study 4 participants reported how many prior studies
they had completed that they knew involved an AI, and how many they thought
likely involved an AI; both counts are right-skewed with a wide range
(clear-AI studies: median
`r fmt_num(est("ai_study_experience", outcome="howmanystudiesclear", term="median"), 0)`,
range `r fmt_num(est("ai_study_experience", outcome="howmanystudiesclear", term="min"), 0)`--`r fmt_num(est("ai_study_experience", outcome="howmanystudiesclear", term="max"), 0)`;
any-AI studies: median
`r fmt_num(est("ai_study_experience", outcome="howmanystudiestotal", term="median"), 0)`,
range `r fmt_num(est("ai_study_experience", outcome="howmanystudiestotal", term="min"), 0)`--`r fmt_num(est("ai_study_experience", outcome="howmanystudiestotal", term="max"), 0)`),
indicating most participants were relatively new to AI studies while a small tail
reported extensive prior exposure (Table \@ref(tab:ra-ai-experience)).

```{r ra-ai-experience}
aix <- all_numbers |>
  dplyr::filter(block == "ai_study_experience") |>
  dplyr::mutate(Item = dplyr::recode(outcome,
                  howmanystudiesclear = "Prior studies known to involve AI",
                  howmanystudiestotal = "Prior studies likely to involve AI"),
                term = factor(term, levels = c("n_answered","min","p25","median","p75","max","mean"))) |>
  dplyr::filter(!is.na(term)) |>
  dplyr::mutate(Statistic = dplyr::recode(term,
                  n_answered = "N answered", min = "Min", p25 = "25th pct",
                  median = "Median", p75 = "75th pct", max = "Max", mean = "Mean"),
                Value = fmt_num(estimate, 1)) |>
  dplyr::arrange(Item, term) |>
  dplyr::select(Item, Statistic, Value) |>
  tidyr::pivot_wider(names_from = Statistic, values_from = Value)
si_kable_styled(aix, caption = "Study 4 self-reported prior AI-study experience: distribution of the number of prior studies participants knew involved an AI, and the number they thought likely involved an AI.")
```

## Durability and the Corrective Debrief

```{r rb-debrief-s13}
db13 <- all_numbers |>
  dplyr::filter(section == "S1-3", block == "debrief",
                term == "post_to_debrief_shift") |>
  dplyr::mutate(`Shift [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                p = fmt_p(p_value)) |>
  dplyr::select(Study = model, Direction = direction, n, `Shift [95% CI]`, p)
si_kable_styled(db13, caption = "Studies 1-3 post-conversation to post-debrief belief shift, bunking arm.", digits = 2)
```

```{r rb-debrief-s4}
db4 <- all_numbers |>
  dplyr::filter(section == "S4 + pooled", block == "debrief_shift",
                term == "post_to_debrief_change") |>
  dplyr::mutate(Sample = dplyr::recode(sample,
                  strict_n1272 = "Observed outcomes", compliant_n1073 = "Compliant"),
                `Shift [95% CI]` = fmt_ci(estimate, conf_low, conf_high, 2),
                p = fmt_p(p_value)) |>
  dplyr::select(Sample, Model = model, Direction = direction, n,
                `Shift [95% CI]`, p)
si_kable_styled(db4, caption = "Study 4 post-conversation to post-debrief belief shift, both arms, by model (strict and compliant samples). n is the model x direction cell size for model rows and the full debrief-complete sample for Pooled rows.", digits = 2)
```

The trajectory in Table \@ref(tab:rb-debrief-traj) places the three measurement
points (pre-treatment, post-treatment, post-debrief) on a common belief scale so
the rise-and-correction shape is visible directly, and Figure
\@ref(fig:rb-fig-debrief-traj) renders the same trajectory across studies and
models.

```{r rb-debrief-traj}
dbt_s <- all_numbers |>
  dplyr::filter(section == "S4 + pooled", block == "debrief_trajectory",
                sample == "strict_n1272") |>
  dplyr::transmute(direction, term, n_s = n,
                   strict = fmt_ci(estimate, conf_low, conf_high, 2))
dbt_c <- all_numbers |>
  dplyr::filter(section == "S4 + pooled", block == "debrief_trajectory",
                sample == "compliant_n1073") |>
  dplyr::transmute(direction, term, n_c = n,
                   comp = fmt_ci(estimate, conf_low, conf_high, 2))
dbt <- dplyr::full_join(dbt_s, dbt_c, by = c("direction", "term")) |>
  dplyr::mutate(Timepoint = factor(term,
                  levels = c("Pre-treatment", "Post-treatment", "Post-debrief"))) |>
  dplyr::arrange(direction, Timepoint) |>
  dplyr::select(Direction = direction, Timepoint,
                `n (strict)` = n_s, `strict [95% CI]` = strict,
                `n (compliant)` = n_c, `compliant [95% CI]` = comp)
si_kable_styled(dbt, caption = "Study 4 mean belief at pre-treatment, post-treatment, and post-debrief, by direction, on the observed-outcome and compliant samples.", digits = 2)
```

```{r rb-fig-debrief-traj, fig.cap="Post-conversation to post-debrief belief shift by study and model, both arms. For Study 4 both the observed-outcome and compliant samples are shown (solid points/lines = Observed outcomes; hollow points/dashed lines = Compliant); Studies 1-3 have a single analytic sample."}
fig_debrief_trajectory()
```
